### 2018-04-20 direct_integrals.tex
creating repo with existing draft

\documentclass[reqno,T1]{amsproc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[canadian]{babel}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{xifthen} % provides \isempty
\usepackage{xparse} % new commands with more than one optional argument

% commutative diagrams
\usepackage{tikz}
\usetikzlibrary{arrows,cd,decorations.markings,shapes.geometric,shapes}

% color of links
\definecolor{myurlcolor}{rgb}{0,0,0.4}
\definecolor{mycitecolor}{rgb}{0,0.5,0}
\definecolor{myrefcolor}{rgb}{0.5,0,0}
\usepackage[pagebackref,draft=false]{hyperref}
\hypersetup{colorlinks,
linkcolor=myrefcolor,
citecolor=mycitecolor,
urlcolor=myurlcolor}
\renewcommand*{\backref}[1]{$\uparrow$\,#1}

% general mathematical macros
\renewcommand{\H}{\mathcal{H}}	% Hilbert space
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\eps}{\varepsilon}
\newcommand{\lin}[1]{\mathrm{lin}(#1)}
\newcommand{\dom}{\mathrm{dom}}
\newcommand{\supp}{\mathrm{supp}}

% categories
\newcommand{\cat}[1]{\mathsf{#1}}		% generic category
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Vect}[1]{\mathsf{Vect}_{#1}}	% vector spaces over a field
\newcommand{\Ban}{\mathsf{Ban}}			% (complex) Banach spaces
\newcommand{\Hilb}{\mathsf{Hilb}} 		% Hilb, cat of Hilbert spaces
\newcommand{\Rep}[1]{\mathsf{Rep}(#1)}		% cat normal Hilbert space reps of a C*-algebra
\newcommand{\NRep}[1]{\mathsf{NRep}(#1)}	% cat of normal Hilbert space reps of a W*-algebra
\newcommand{\id}[1]{1_{#1}}			% identity morphism

% Theorem Environments
\swapnumbers
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{qstn}[thm]{Question}
\newtheorem{defn}[thm]{Definition}
\newtheorem{lemdefn}[thm]{Lemma + Definition}
\newtheorem{prob}[thm]{Problem}
\theoremstyle{remark}
\newtheorem{ex}[thm]{Example}
\newtheorem{rem}[thm]{Remark}
\newtheorem{note}[thm]{Note}
\numberwithin{equation}{section}

%%%%%%%%%%%% Enumeration via lowercase letters
\renewcommand{\labelenumi}{(\alph{enumi})}
\renewcommand{\theenumi}{(\alph{enumi})}
\renewcommand{\labelitemi}{$\circ$}

%%% pagebreaks allowed for align environment
\allowdisplaybreaks

% string diagram stuff
\pgfdeclarelayer{edgelayer}
\pgfdeclarelayer{nodelayer}
\pgfsetlayers{edgelayer,nodelayer,main}
\tikzstyle{none}=[inner sep=0pt]
\tikzstyle{simple}=[-,draw=black,line width=1.000]
\tikzstyle{bn}=[circle,inner sep=2pt,fill=black,draw=black,line width=0.8 pt]

%%% draft stuff
%\usepackage[draft]{showkeys}
\usepackage{todonotes}
\newcommand{\tob}[1]{\todo[color=blue!40,inline]{#1}}

%-------------------------------------------------------------------

\begin{document}
\sloppy

% vertical spacing in multiline equations
\setlength{\jot}{6pt}

%-------------------------------------------------------------------

%%%%%%%%%%%% title page stuff %%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The universal property of direct integrals in W*-categories}

\author{Tobias Fritz}

\address{Max Planck Institute for Mathematics in the Sciences, Leipzig, Germany}
\email{fritz@mis.mpg.de}

\keywords{}

\subjclass[2010]{Primary: ; Secondary: }

\thanks{\textit{Acknowledgements.} }

\begin{abstract}
We formulate a universal property for infinite direct sums in W*-categories, and prove that this indeed characterizes the usual infinite direct sums in the W*-category of representations of a C*-algebra. We do the same for direct integrals, in which case we need to work with measurable W*-categories.
\end{abstract}

\maketitle

\section{Introduction}

\section{W*-categories}

Let $C$ be a W*-category~\cite{wstarcat}, such as $\Hilb$ or $\Rep{A}$, the category of Hilbert space representations of a C*-algebra $A$.

\begin{lem}
Let $u : A \to B$ be an invertible morphism with $\| u \| = \| u^{-1} \| = 1$. Then $u$ is unitary.
\label{unitaries}
\end{lem}

\todo[inline]{We could state this at the generality of C*-cats}

\begin{proof}
We have $\| u^* u \| = 1$ by the C*-identity, and likewise $\| u^{-1} (u^{-1})^* \| = 1$. Since $(u^* u)^{-1} = u^{-1} (u^{-1})^*$, we also have $\| (u^* u)^{-1} \| = 1$. This makes $u^* u$ into a positive element of unit norm whose inverse also has unit norm, and therefore $u^* u = \id{A}$. By the same token, we obtain $u u^* = \id{B}$.
\end{proof}

Now let $\Ban$ be the category of Banach spaces with bounded linear maps of norm $\leq 1$. Every W*-category is $\Ban$-enriched. Lemma~\ref{unitaries} and the Yoneda lemma imply that for any representable functor $\cat{C} \to \Ban$, the representing object is unique up to unique \emph{unitary} isomorphism. This is a desideratum of a notion of universal property for W*-categories, and more generally for all dagger categories~\cite{daglims}. The lemma tells us that the compatibility with the dagger structure is automatic thanks to the enrichment in $\Ban$, so that we do not need to formulate our universal properties in the form of dagger limits as in~\cite{daglims}.

\section{The universal property of infinite direct sums}

For $(A_i)_{i\in I}$ a family of objects in $\cat{C}$, we consider the functor
\[
	\bigoplus_i \cat{C}(A_i,-) \: : \: \cat{C} \to \Ban
\]
given by associating to every object $B$ the vector space
\[
	\bigoplus_i \cat{C}(A_i , B) := \left\{ (f_i : A_i \to B)_{i\in I} \Biggm\vert \sum_i f_i f_i^* < \infty \right\},
\]
considered as a Banach space under the norm
\[
	\| (f_i)_{i \in I} \| := \left\| \sum_i f_i f_i^* \right\|^{1/2}.
\]
This is functorial in $B$ in the obvious way.

\todo[inline]{Is it actually complete?}

\begin{defn}
\label{directsumdef}
An \emph{$I$-indexed direct sum} of the family of objects $(A_i)_{i\in I}$ is an object $\bigoplus_i A_i$ which represents the $\Ban$-enriched functor $\bigoplus_i \cat{C}(A_i,-)$.
\end{defn}

By Lemma~\ref{unitaries} and the $\Ban$-enriched Yoneda lemma, this characterizes infinite direct sums up to unique isomorphism. Since we are in a dagger category, the same holds dually, meaning that $\bigoplus_i A_i$ also represents the functor
\[
	\bigoplus_i \cat{C}(-,A_i) \: : \: \cat{C}^\op \to \Ban
\]
given by associating to every object $B$ the vector space
\[
	\bigoplus_i \cat{C}(B, A_i) := \left\{ (f_i : B \to A_i)_{i\in I} \Biggm\vert \sum_i f_i^* f_i < \infty \right\},
\]
with norm
\[
	\| (f_i)_{i \in I} \| := \left\| \sum_i f_i^* f_i \right\|^{1/2}.
\]
which is functorial in $B$ in the obvious way.

\newcommand{\alg}{N}	% a W*-algebra

For a W*-algebra $\alg$, we let $\NRep{\alg}$ be its category of normal representations on Hilbert spaces. For example, $\alg$ may be the double dual of a C*-algebra, in which case $\NRep{\alg}$ is equivalent to the category of representations of the C*-algebra.\footnote{Our choice of font $\alg$ is the same as that for categories, suggesting that $\alg$ may as well more generally be a small W*-category, and the same statements would hold. Since one can always take the convolution W*-algebra of any small W*-category, resulting in the same category of representations, there is no loss of generality in restricting ourselves to W*-algebras.}

\begin{thm}
In $\NRep{\alg}$, any $\ell^2$-direct sum
\[
	\bigoplus A_i := \left\{ (\xi_i)_{i\in I} \Biggm \vert \xi_i\in A_i,\: \sum_i \langle\xi_i,\xi_i\rangle < \infty \right\},
\]
equipped with the blockwise representation of $\alg$, is indeed an \emph{$I$-indexed direct sum} in the sense of Definition~\ref{directsumdef}. Every $I$-indexed direct sum in $\NRep{\alg}$ is of this form.
\end{thm}

\begin{proof}
\todo[inline]{do this}
\end{proof}

\begin{thm}
The following are equivalent:
\begin{enumerate}
\item $\bigoplus_i A_i$ is an $I$-indexed direct sum.
\item There is a family of morphisms $\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$ such that $\kappa_j^*\kappa_k = \delta_{jk}$, and
\[
	\sum_j \kappa_j \kappa_j^* = 1_{\bigoplus_i A_i}
\]
ultraweakly.
\end{enumerate}
\end{thm}

\begin{proof}
\todo[inline]{do this}
\end{proof}

\todo[inline]{Does this also result in a dagger limit in the sense of~\cite{daglims}?}

\section{Direct integrals}

\newcommand{\Obj}[1]{\mathrm{Obj}(#1)}
\newcommand{\Mor}[1]{\mathrm{Mor}(#1)}

\begin{defn}
A \emph{measurable category} $\cat{C}$ is a category internal to measurable spaces. A \emph{measured family of objects} indexed by a measure space $(X,\Sigma,\mu)$ is a measurable function $A : X\to\Obj{\cat{C}}$.
\end{defn}

For $(A_x)_{x\in X}$ a measured family of objects in $\cat{C}$, we now consider the functor
\[
	\int_X \cat{C}(A_x,-) \: : \: \cat{C}\to\Set
\]
given by associating to every object $B$ the set
\[
	\int_X \cat{C}(A_x , B) := \left\{ f : X\to\Mor{\cat{C}} \textrm{ measurable} \Biggm\vert f_x : A_x\to B,\: \int_X f_x f_x^* \,dx < \infty \right\},
\]
which is functorial in $B$ in the obvious way.

\todo[inline]{Is this the ``right'' definition? What is the relation to~\cite{measureablecats}?}

\begin{defn}
A W*-category \emph{has direct integrals} if it has direct integrals for $\sigma$-finite measure spaces and arbitrary direct sums.
\end{defn}

\todo[inline]{Is the restriction to $\sigma$-finite the ``right'' definition?}

Due to the existence of countable direct sums, we could as well restrict to direct integrals on finite measure spaces:

\begin{prop}
If countable direct sums and direct integrals for finite measure spaces and exist, then so do direct integrals for $\sigma$-finite measure spaces.
\end{prop}

If everything works out, this is due to a nice kind of compositionality of direct integrals (essentially Fubini's theorem).

\todo[inline]{Prove Fubini's theorem for direct integrals from the universal property?}

\bibliographystyle{plain}
\bibliography{direct_integrals}

\end{document}

### 2018-04-27 direct_integrals.tex
two lemmas

@@
 \usepackage{amsfonts}
+\usepackage{multicol}
 \usepackage[canadian]{babel}
@@
 \section{Introduction}
+
+\section{C*-categories}
+For the balance of this section, let~$C$ be a C$^*$-category. \cite{wstarcat}
+We start with a generalization of~\cite[Lemma 7]{westerbaan2016universal}
+    to C$^*$-categories.
+\begin{lem}
+For any~$a\colon A \to B$ with~$\|a\| \leq 1$
+    and projections~$s \colon A \to A$,
+        $t\colon B\to B$,
+    we have
+    \begin{equation*}
+     a^* t a \leq 1-s \quad\iff\quad
+    a s a^* \leq 1 - t \quad\iff\quad
+    tas = 0 \quad\iff\quad
+    sa^* t = 0.
+    \end{equation*}\label{contrapositionlemma}
+\end{lem}
+\begin{proof}
+We will first prove~$a^*t a \leq 1-s \Leftrightarrow tas=0$.
+    Assume~$a^* t a \leq 1-s$.
+    Then~$0 \leq s a^* t a s \leq s (1-s) s = 0$.
+    Thus~$s a^* t a s = 0$, which is to say~$tas = 0$.
+    Now, assume~$tas = 0$.
+    Then~$ta = ta(1-s)$ and~$a^*t = (1-s)a^* t$.
+    Hence~$a^* ta = (1-s)a^*ta(1-s) \leq (1-s)$.
+A similar argument proves~$as a^* \leq 1-t \Leftrightarrow sa^*t=0$.
+To finish the proof, it is sufficient to
+        show~$tas=0 \Leftrightarrow sa^*t=0$, which follows directly
+        by applying~$(\ )^*$.
+\end{proof}
+
+\begin{lem}
+Let~$a \colon A \leftrightarrows B \colon b$ be any two morphisms
+    such that~$\|a \| = \| b\| = 1$ and~$ab=1$.
+    Then~$b$ is an isometry with~$a=b^*$.
+\end{lem}
+\begin{proof}
+    To start, note~$\|a ^*a\| = \|a\|^2 = 1$, so~$a^*a \leq 1$.
+    Similarly~$b^*b \leq 1$.
+    Combined~$1 = b^*a^*ab \leq b^*b \leq 1$, hence~$b^*b=1$.
+    Similarly~$aa^*=1$.
+    Clearly~$bb^*bb^* = b 1 b^* = bb^*$, so~$b$ is an isometry.
+    Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0$
+    So~$a^* (1-0) a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
+    Similarly~$bb^* \leq a^*a$. Hence~$bb^*=a^*a$.
+    We are done: $a = aa^*a=abb^*=b^*$.
+\end{proof}
 
@@
 \end{document}
+
+% vim: ft=tex.latex

### 2018-04-27 direct_integrals.tex
added one direction to proof of main theorem

@@
 \newcommand{\Hilb}{\mathsf{Hilb}} 		% Hilb, cat of Hilbert spaces
-\newcommand{\Rep}[1]{\mathsf{Rep}(#1)}		% cat normal Hilbert space reps of a C*-algebra
-\newcommand{\NRep}[1]{\mathsf{NRep}(#1)}	% cat of normal Hilbert space reps of a W*-algebra
+\newcommand{\Rep}[1]{\mathsf{Rep}(#1)}		% cat normal Hilbert space reps of a C$^*$-algebra
+\newcommand{\NRep}[1]{\mathsf{NRep}(#1)}	% cat of normal Hilbert space reps of a W$^*$-algebra
 \newcommand{\id}[1]{1_{#1}}			% identity morphism
@@
 %%% draft stuff
-%\usepackage[draft]{showkeys}
+\usepackage[draft]{showkeys}
 \usepackage{todonotes}
@@
 
-\title{The universal property of direct integrals in W*-categories}
+\title{The universal property of direct integrals in W$^*$-categories}
 
-\author{Tobias Fritz}
+\author{Tobias Fritz, Robert Furber, Bas Westerbaan}
 
@@
 \begin{abstract}
-We formulate a universal property for infinite direct sums in W*-categories, and prove that this indeed characterizes the usual infinite direct sums in the W*-category of representations of a C*-algebra.
-We do the same for direct integrals, in which case we need to work with measurable W*-categories.
+Using enrichment in Banach spaces, we observe that one can formulate universal properties characterizing universal objects in C$^*$-categories in such a way that they guarantee uniqueness up to unique unitaries.
+We then formulate such a universal property for infinite direct sums in W$^*$-categories, and prove that this indeed characterizes the standard infinite direct sums in the W$^*$-category of normal representations of a W$^*$-algebra.
+We do the same for direct integrals, in which case we need to work with measurable W$^*$-categories.
 \end{abstract}
@@
 
-\section{C*-categories}
-For the balance of this section, let~$C$ be a C$^*$-category. \cite{wstarcat}
+\section{C$^*$-categories}
+
+For this section, we work in a C$^*$-category \cite{wstarcat} denoted by $\cat{C}$.
 We start with a generalization of~\cite[Lemma 7]{westerbaan2016universal}
     to C$^*$-categories.
+
 \begin{lem}
@@
         $t\colon B\to B$,
-    we have
-    \begin{equation*}
-     a^* t a \leq 1-s \quad\iff\quad
-    a s a^* \leq 1 - t \quad\iff\quad
-    tas = 0 \quad\iff\quad
-    sa^* t = 0.
-    \end{equation*}\label{contrapositionlemma}
+	the following are equivalent:
+\begin{enumerate}
+\item $a^* t a \leq 1-s$,
+\item $a s a^* \leq 1 - t$,
+\item $tas = 0$,
+\item $sa^* t = 0$.
+\end{enumerate}
+\label{contrapositionlemma}
 \end{lem}
@@
 Let~$a \colon A \leftrightarrows B \colon b$ be any two morphisms
-    such that~$\|a \| = \| b\| = 1$ and~$ab=1$.
+    such that~$\|a \| \leq 1$ and $\| b\| \leq 1$ and~$ab=1$.
     Then~$b$ is an isometry with~$a=b^*$.
+\label{isometrylemma}
 \end{lem}
 \begin{proof}
-    To start, note~$\|a ^*a\| = \|a\|^2 = 1$, so~$a^*a \leq 1$.
+    To start, note~$\|a ^*a\| = \|a\|^2 \leq 1$, so~$a^*a \leq 1$.
     Similarly~$b^*b \leq 1$.
@@
 
-\section{W*-categories}
+We will need the full power of this result only in the next section, but the following special case is generally useful for the definition of universal properties in C$^*$-categories:
 
-Let $C$ be a W*-category~\cite{wstarcat}, such as $\Hilb$ or $\Rep{A}$, the category of Hilbert space representations of a C*-algebra $A$.
-
-\begin{lem}
-Let $u : A \to B$ be an invertible morphism with $\| u \| = \| u^{-1} \| = 1$. Then $u$ is unitary.
+\begin{cor}
+Let $u : A \to B$ be an invertible morphism with $\| u \| \leq 1$ and $\| u^{-1} \| \leq 1$. Then $u$ is unitary.
 \label{unitaries}
-\end{lem}
-
-\todo[inline]{We could state this at the generality of C*-cats}
-
-\begin{proof}
-We have $\| u^* u \| = 1$ by the C*-identity, and likewise $\| u^{-1} (u^{-1})^* \| = 1$.
-Since $(u^* u)^{-1} = u^{-1} (u^{-1})^*$, we also have $\| (u^* u)^{-1} \| = 1$.
-This makes $u^* u$ into a positive element of unit norm whose inverse also has unit norm, and therefore $u^* u = \id{A}$.
-By the same token, we obtain $u u^* = \id{B}$.
-\end{proof}
+\end{cor}
 
 Now let $\Ban$ be the category of Banach spaces with bounded linear maps of norm $\leq 1$.
-Every W*-category is $\Ban$-enriched.
+Every C$^*$-category is $\Ban$-enriched.
 Lemma~\ref{unitaries} and the Yoneda lemma imply that for any representable functor $\cat{C} \to \Ban$, the representing object is unique up to unique \emph{unitary} isomorphism.
-This is a desideratum of a notion of universal property for W*-categories, and more generally for all dagger categories~\cite{daglims}.
+This is a desideratum of a notion of universal property for C$^*$-categories, and more generally for all dagger categories~\cite{daglims}.
 The lemma tells us that the compatibility with the dagger structure is automatic thanks to the enrichment in $\Ban$, so that we do not need to formulate our universal properties in the form of dagger limits as in~\cite{daglims}.
 
+In summary, our notion of universal property in a C$^*$-category is the standard one for enriched categories, namely with enrichment in $\Ban$:
+
+\begin{defn}
+A functor $F : \cat{C} \to \Ban$ is \emph{representable} if there is an object $A \in \cat{C}$ together with a natural isomorphism $\cat{C}(A,-)\cong F$ that is a componentwise isometric isomorphism of Banach spaces.
+\end{defn}
+
+Moreover, thanks to Corollary~\ref{unitaries}, the representing object is unique up to a unique unitary: if both $A$ and $B$ represent $F$ via $\cat{C}(A,-)\cong F\cong \cat{C}(B,-)$, then the Yoneda lemma gives us an isomorphism $u : A \to B$.
+Since the natural isomorphism between the hom-functors is required to be a componentwise isometry, we have $\| u \| = \| u^{-1} \| = 1$, making $u$ into a unitary.
+
 \section{The universal property of infinite direct sums}
+
+We now turn to a particular universal property and study infinite direct sums.
+(While our considerations also apply to finite direct sums, in this case our discussion provides nothing new, since then it reduces to the well-understood case of biproducts in an additive category.) From now on, we take $C$ to be a W$^*$-category~\cite{wstarcat}, such as $\Hilb$ or $\Rep{A}$, the category of Hilbert space representations of a C$^*$-algebra $A$.
 
@@
 
-\todo[inline]{Is it actually complete?}
+\begin{lem}
+$\bigoplus_i \cat{C}(A_i,-)$ is complete, and therefore a Banach space.
+\end{lem}
+
+The proof is analogous to the standard proof showing that $\ell^2(I)$ is a Hilbert space. \tob{I haven't done this yet, so for now this statement is wishful thinking}
+
+\begin{proof}
+Suppose that we are given a Cauchy sequence of families $( ( f_i^{(n)} )_{i\in I} )_{n\in\N}$.
+
+\todo{Complete this! Showing that $(f_i^{(n)})_{n\in\N}$ is Cauchy for each $i$ needs some form of Cauchy-Schwarz}
+\end{proof}
+
+We now get to our main definition:
 
@@
 
-By Lemma~\ref{unitaries} and the $\Ban$-enriched Yoneda lemma, this characterizes infinite direct sums up to unique isomorphism.
+By the discussion of the previous section, this characterizes infinite direct sums up to unique unitaries.
 Since we are in a dagger category, the same holds dually, meaning that $\bigoplus_i A_i$ also represents the functor
@@
 \[
-	\| (f_i)_{i \in I} \| := \left\| \sum_i f_i^* f_i \right\|^{1/2}.
+	\| (f_i)_{i \in I} \| := \left\| \sum_i f_i^* f_i \right\|^{1/2},
 \]
@@
 
-\newcommand{\alg}{N}	% a W*-algebra
+\begin{thm}
+The following are equivalent:
+\begin{enumerate}
+\item\label{universalprop} $\bigoplus_i A_i$ is an $I$-indexed direct sum.
+\item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$ such that $\kappa_j^*\kappa_k = \delta_{jk}$, and
+\beq
+\label{complete}
+	\sum_j \kappa_j \kappa_j^* = 1_{\bigoplus_i A_i}
+\eeq
+ultraweakly.
+\end{enumerate}
+\end{thm}
 
-For a W*-algebra $\alg$, we let $\NRep{\alg}$ be its category of normal representations on Hilbert spaces.
-For example, $\alg$ may be the double dual of a C*-algebra, in which case $\NRep{\alg}$ is equivalent to the category of representations of the C*-algebra.\footnote{Our choice of font $\alg$ is the same as that for categories, suggesting that $\alg$ may as well more generally be a small W*-category, and the same statements would hold.
-Since one can always take the convolution W*-algebra of any small W*-category, resulting in the same category of representations, there is no loss of generality in restricting ourselves to W*-algebras.}
+\begin{proof}
+We write $A$ as shorthand for $\bigoplus_i A_i$.
+
+Assuming \ref{universalprop}, we define $(\kappa_j : A_j \to A)$ to be the family that corresponds to $\id{A}$ under the given isomorphism $\bigoplus_i \cat{C}(A_i,A) \cong \cat{C}(A,A)$.
+For fixed $j\in J$, also the family $(\delta_{ij} : A_i \to A_j)$ corresponds to some morphism $\pi_j : A \to A_j$.
+Applying naturality with respect to $\pi_j$ gives the diagram
+\[\begin{tikzcd}
+	\bigoplus_i \cat{C}(A_i,A) \ar{r}{\cong} \ar{d}{(\pi_j\circ -)_i} & \cat{C}(A,A) \ar{d}{\pi_j\circ -} \\
+	\bigoplus_i \cat{C}(A_i,A_j) \ar{r}{\cong} & \cat{C}(A,A_j)
+\end{tikzcd}\]
+Starting with the family $(\kappa_i)$ on the upper left, we get $(\pi_j\circ\kappa_i)$ by going down, or alternatively $(\delta_{ij})$ by going around; therefore we have $\pi_j \kappa_i = \delta_{ij}$.
+By the isometric isomorphism assumption, we moreover have $\| \id{A} \| = \| \sum_i \kappa_i \kappa_i^* \|$, and therefore $\| \kappa_j \| \leq 1$; and similarly $\| \pi_j \| \leq 1$.
+By Lemma~\ref{isometrylemma}, this implies $\pi_j = \kappa_j^*$.
+We therefore have $\kappa_j^*\kappa_i = \delta_{ij}$, as desired.
+This also implies that the $\kappa_j \kappa_j^*$ are mutually orthogonal projections.
+In order to also prove the completeness relation~\eqref{complete}, let $p : A\to A$ be any other projection that is orthogonal to each of the $\kappa_j \kappa_j^*$; we need to show $p = 0$.
+Again by naturality, under the isomorphism $\bigoplus_i \cat{C}(A_i,A)$, the new $p$ corresponds to the family $(p\kappa_i) = (0)$.
+This implies indeed $p=0$.
+
+Assuming \ref{biprodsum}, \todo{write this}
+\end{proof}
+
+\begin{rem}
+For finite $I$, these direct sums are also dagger limits in the sense of~\cite{daglims}.
+For infinite $I$, this is no longer the case, since infinite direct sums are not even limits in the usual $\cat{Set}$-enriched sense; compare~\cite[Theorem~5.2]{daglims}.
+Hence our definition are better adapted to the context of W$^*$-categories, such as $\cat{Hilb}$, than those of~\cite{daglims}.
+\tob{We can tone down the last sentence a bit if it's too demeaning}
+\end{rem}
+
+\newcommand{\alg}{N}	% a W$^*$-algebra
+
+For a W$^*$-algebra $\alg$, we let $\NRep{\alg}$ be its category of normal representations on Hilbert spaces.
+For example, $\alg$ may be the double dual of a C$^*$-algebra, in which case $\NRep{\alg}$ is equivalent to the category of representations of the C$^*$-algebra.\footnote{Our choice of font $\alg$ is the same as that for categories, suggesting that $\alg$ may as well more generally be a small W$^*$-category, and the same statements would hold.
+Since one can always take the convolution W$^*$-algebra of any small W$^*$-category, resulting in the same category of representations, there is no loss of generality in restricting ourselves to W$^*$-algebras.}
 
@@
 \end{proof}
-
-\begin{thm}
-The following are equivalent:
-\begin{enumerate}
-\item $\bigoplus_i A_i$ is an $I$-indexed direct sum.
-\item There is a family of morphisms $\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$ such that $\kappa_j^*\kappa_k = \delta_{jk}$, and
-\[
-	\sum_j \kappa_j \kappa_j^* = 1_{\bigoplus_i A_i}
-\]
-ultraweakly.
-\end{enumerate}
-\end{thm}
-
-\begin{proof}
-\todo[inline]{do this}
-\end{proof}
-
-\todo[inline]{Does this also result in a dagger limit in the sense of~\cite{daglims}?}
 
@@
 \begin{defn}
-A W*-category \emph{has direct integrals} if it has direct integrals for $\sigma$-finite measure spaces and arbitrary direct sums.
+A W$^*$-category \emph{has direct integrals} if it has direct integrals for $\sigma$-finite measure spaces and arbitrary direct sums.
 \end{defn}

### 2018-04-27 direct_integrals.tex
small clarification

@@
 Starting with the family $(\kappa_i)$ on the upper left, we get $(\pi_j\circ\kappa_i)$ by going down, or alternatively $(\delta_{ij})$ by going around; therefore we have $\pi_j \kappa_i = \delta_{ij}$.
-By the isometric isomorphism assumption, we moreover have $\| \id{A} \| = \| \sum_i \kappa_i \kappa_i^* \|$, and therefore $\| \kappa_j \| \leq 1$; and similarly $\| \pi_j \| \leq 1$.
+By the isometric isomorphism assumption, we moreover have $\| \id{A} \| = \| \sum_i \kappa_i \kappa_i^* \|$, and therefore $\| \kappa_j \| \leq 1$; and similarly $\| \pi_j \| = \| \id{A} \| \leq 1$.
 By Lemma~\ref{isometrylemma}, this implies $\pi_j = \kappa_j^*$.

### 2018-04-30 direct_integrals.tex
proof

@@
 \usepackage{xparse} % new commands with more than one optional argument
+\newcommand{\unlim}{\qopname\relax m{unlim}}
 
@@
 \item\label{universalprop} $\bigoplus_i A_i$ is an $I$-indexed direct sum.
+\item\label{universalprop2} 
+There is a family of
+        morphisms~$\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$
+        with~$\sum_{j \in I} \kappa_j \kappa_j^* < \infty$
+    such that for any object~$B$
+        and family of
+        morphisms~$\left(T_j: A_j \to B\right)_{j\in I}$
+        with~$\sum_{j \in I}T_jT_j^* < \infty$,
+        there is a unique~$T\colon \bigoplus_i A_i \to B$
+        with~$T \kappa_j = T_j$ for all~$j \in I$.
 \item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$ such that $\kappa_j^*\kappa_k = \delta_{jk}$, and
@@
 \label{complete}
-	\sum_j \kappa_j \kappa_j^* = 1_{\bigoplus_i A_i}
+        \sum_{j\in I} \kappa_j \kappa_j^* \ =\  1_{\bigoplus_i A_i}.
 \eeq
-ultraweakly.
 \end{enumerate}
@@
 
-Assuming \ref{biprodsum}, \todo{write this}
+To finish, we will prove show \ref{biprodsum} $\Rightarrow$
+        \ref{universalprop2}.
+Let~$\left(T_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
+        with~$\sum_{j \in I}T_jT_j^* < \infty$.
+The homset~$C(B,A)$
+    is a self-dual Hilbert~$C(B,B)$-module
+    with~$C(B,B)$-valued inner
+    product~$\langle S, T\rangle \equiv S^*T$.\cite[2.15]{wstarcat}
+The self-duality is equivalent to the fact
+    that~$C(B,A)$ is complete in the \emph{ultranorm} topology;
+    that is, the uniformity
+    generated by the seminorms~$\| T \|_\omega \equiv \omega(\langle T,T\rangle)^\frac{1}{2}$
+    for normal states~$\omega$ on~$C(B,B)$.\cite[\S149\textsubscript{V}]{bas}
+    Pick any normal state~$\omega$ on~$C(B,B)$.
+        and finite subset~$S \subseteq I$.  We have
+\begin{alignat*}{2}
+    \bigl\| \sum_{j \in S} \kappa_j T_j^* \bigr\|^2_\omega
+    &\ \equiv \ 
+    \omega\Bigl(\bigl(\sum_{i \in S} \kappa_i T_i^*\bigr)^*
+            \bigl(\sum_{j \in S} \kappa_j T_j^*\bigr)\Bigr) \\
+    &\ =\ 
+    \omega\Bigl( \sum_{j \in S} T_j T_j^*
+    \Bigr) &\qquad&\text{as~$\kappa_j^*\kappa_i = \delta_{ij}$}.
+\end{alignat*}
+From this and~$\sum_{j \in I} T_jT_j^* < \infty$,
+    it follows that~$(\sum_{j \in S} \kappa_j T_j^*)_S$
+    is ultranorm Cauchy.
+    Define~$T \equiv ( \sum_{j \in I} \kappa_jT_j^* )^*$,
+        where the sum is understood to converge ultranorm.
+The assignment~$f \mapsto \kappa_j^*f$
+    yields a bounded~$C(B,B)$-linear map between
+    the Hilbert~$C(B,B)$-modules~$C(B,A)$ and~$C(B,A_j)$
+    and is therefore ultranorm continuous.\cite[\S148]{bas}
+    Hence
+        $\kappa_j^* T^*
+             =  \kappa_j^* \sum_{i} \kappa_i T_i^*
+             =  \sum_i \kappa_{j }^*\kappa_i T_i^*
+            =  T_j^*$
+            and so~$T \kappa_j = T_j$ as desired.
+Finally, to show uniqueness,
+    assume~$T'\colon A \to B$ is any (other)
+    morphism with~$T' \kappa_i = T_i$.
+    Note that~$\sum_j \kappa_j \kappa_j^*$
+       converges ultrastrongly
+       and so ultranorm as well.
+With similar reasoning as before,
+        the map~$f \mapsto T' f$ is ultranorm continuous,
+        hence
+\begin{equation*}
+    T'
+    \ =\ T' \sum_{j \in I} \kappa_j \kappa_j^*
+    \ =\ \sum_{j \in I} T' \kappa_j \kappa_j^*
+    \ =\ \sum_{j \in I} T_j \kappa_j^*
+    \ =\ \sum_{j \in I} T \kappa_j \kappa_j^*
+    \ =\ T,
+\end{equation*}
+    where the sums converge ultranorm.
 \end{proof}

### 2018-04-30 direct_integrals.tex
some uniformization

@@
 
-To finish, we will prove show \ref{biprodsum} $\Rightarrow$
-        \ref{universalprop2}.
-Let~$\left(T_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
+Assuming \ref{biprodsum},
+let~$\left(T_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
         with~$\sum_{j \in I}T_jT_j^* < \infty$.
 The homset~$C(B,A)$
-    is a self-dual Hilbert~$C(B,B)$-module
-    with~$C(B,B)$-valued inner
+    is a self-dual Hilbert~$\cat{C}(B,B)$-module
+    with~$\cat{C}(B,B)$-valued inner
     product~$\langle S, T\rangle \equiv S^*T$.\cite[2.15]{wstarcat}
 The self-duality is equivalent to the fact
-    that~$C(B,A)$ is complete in the \emph{ultranorm} topology;
+    that~$\cat{C}(B,A)$ is complete in the \emph{ultranorm} topology;
     that is, the uniformity
     generated by the seminorms~$\| T \|_\omega \equiv \omega(\langle T,T\rangle)^\frac{1}{2}$
-    for normal states~$\omega$ on~$C(B,B)$.\cite[\S149\textsubscript{V}]{bas}
-    Pick any normal state~$\omega$ on~$C(B,B)$.
+    for normal states~$\omega$ on~$\cat{C}(B,B)$~\cite[\S149\textsubscript{V}]{bas}.
+    Pick any normal state~$\omega$ on~$\cat{C}(B,B)$.
         and finite subset~$S \subseteq I$.  We have
@@
 The assignment~$f \mapsto \kappa_j^*f$
-    yields a bounded~$C(B,B)$-linear map between
-    the Hilbert~$C(B,B)$-modules~$C(B,A)$ and~$C(B,A_j)$
-    and is therefore ultranorm continuous.\cite[\S148]{bas}
+    yields a bounded~$\cat{C}(B,B)$-linear map between
+    the Hilbert~$\cat{C}(B,B)$-modules~$\cat{C}(B,A)$ and~$\cat{C}(B,A_j)$
+    and is therefore ultranorm continuous~\cite[\S148]{bas}.
     Hence

### 2018-04-30 direct_integrals.tex
correction and some more uniformization

--- 2018-04-27 direct_integrals.tex
@@
 \item\label{universalprop} $\bigoplus_i A_i$ is an $I$-indexed direct sum.
+\item\label{universalprop2} 
+There is a family of
+        morphisms~$\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$
+        with~$\sum_{j \in I} \kappa_j \kappa_j^* < \infty$
+    such that for any object~$B$
+        and family of
+        morphisms~$\left(f_j: A_j \to B\right)_{j\in I}$
+        with~$\sum_{j \in I}f_j f_j^* < \infty$,
+        there is a unique~$f\colon \bigoplus_i A_i \to B$
+        such that $f \kappa_j = f_j$ for all~$j \in I$, and $\| f \|^2 = \| \sum_j f_j f_j^* \|$.
 \item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$ such that $\kappa_j^*\kappa_k = \delta_{jk}$, and
@@
 
-Assuming \ref{biprodsum}, \todo{write this}
+Assuming \ref{biprodsum},
+let~$\left(f_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
+        with~$\sum_{j \in I}f_jf_j^* < \infty$.
+The homset~$C(B,A)$
+    is a self-dual Hilbert~$\cat{C}(B,B)$-module
+    with~$\cat{C}(B,B)$-valued inner
+    product~$\langle g, f\rangle \equiv g^*f$.\cite[2.15]{wstarcat}
+The self-duality is equivalent to the fact
+    that~$\cat{C}(B,A)$ is complete in the \emph{ultranorm} topology;
+    that is, the uniformity
+    generated by the seminorms~$\| f \|_\omega \equiv \omega(\langle f,f\rangle)^\frac{1}{2}$
+    for normal states~$\omega$ on~$\cat{C}(B,B)$~\cite[\S149\textsubscript{V}]{bas}.
+    Pick any normal state~$\omega$ on~$\cat{C}(B,B)$.
+        and finite subset~$S \subseteq I$.  We have
+\begin{alignat*}{2}
+    \bigl\| \sum_{j \in S} \kappa_j f_j^* \bigr\|^2_\omega
+    &\ \equiv \ 
+    \omega\Bigl(\bigl(\sum_{i \in S} \kappa_i f_i^*\bigr)^*
+            \bigl(\sum_{j \in S} \kappa_j f_j^*\bigr)\Bigr) \\
+    &\ =\ 
+    \omega\Bigl( \sum_{j \in S} f_j f_j^*
+    \Bigr) &\qquad&\text{as~$\kappa_j^*\kappa_i = \delta_{ij}$}.
+\end{alignat*}
+From this and~$\sum_{j \in I} f_jf_j^* < \infty$,
+    it follows that~$(\sum_{j \in S} \kappa_j f_j^*)_S$
+    is ultranorm Cauchy.
+    Define~$f \equiv ( \sum_{j \in I} \kappa_jf_j^* )^*$,
+        where the sum is understood to converge ultranorm.
+The assignment~$f \mapsto \kappa_j^*f$
+    yields a bounded~$\cat{C}(B,B)$-linear map between
+    the Hilbert~$\cat{C}(B,B)$-modules~$\cat{C}(B,A)$ and~$\cat{C}(B,A_j)$
+    and is therefore ultranorm continuous~\cite[\S148]{bas}.
+    Hence
+        $\kappa_j^* f^*
+             =  \kappa_j^* \sum_{i} \kappa_i f_i^*
+             =  \sum_i \kappa_{j }^*\kappa_i f_i^*
+            =  f_j^*$
+            and so~$f \kappa_j = f_j$ as desired.
+Finally, to show uniqueness,
+    assume~$T'\colon A \to B$ is any (other)
+    morphism with~$T' \kappa_i = T_i$.
+    Note that~$\sum_j \kappa_j \kappa_j^*$
+       converges ultrastrongly
+       and so ultranorm as well.
+With similar reasoning as before,
+        the map~$f \mapsto T' f$ is ultranorm continuous,
+        hence
+\begin{equation*}
+    T'
+    \ =\ T' \sum_{j \in I} \kappa_j \kappa_j^*
+    \ =\ \sum_{j \in I} T' \kappa_j \kappa_j^*
+    \ =\ \sum_{j \in I} T_j \kappa_j^*
+    \ =\ \sum_{j \in I} T \kappa_j \kappa_j^*
+    \ =\ T,
+\end{equation*}
+    where the sums converge ultranorm.
 \end{proof}

### 2018-04-30 direct_integrals.tex
completed missing adjustments

@@
 Finally, to show uniqueness,
-    assume~$T'\colon A \to B$ is any (other)
-    morphism with~$T' \kappa_i = T_i$.
+    assume~$f'\colon A \to B$ is any (other)
+    morphism with~$f' \kappa_i = f_i$.
     Note that~$\sum_j \kappa_j \kappa_j^*$
@@
 With similar reasoning as before,
-        the map~$f \mapsto T' f$ is ultranorm continuous,
+        the map~$f \mapsto f' f$ is ultranorm continuous,
         hence
 \begin{equation*}
-    T'
-    \ =\ T' \sum_{j \in I} \kappa_j \kappa_j^*
-    \ =\ \sum_{j \in I} T' \kappa_j \kappa_j^*
-    \ =\ \sum_{j \in I} T_j \kappa_j^*
-    \ =\ \sum_{j \in I} T \kappa_j \kappa_j^*
-    \ =\ T,
+    f'
+    \ =\ f' \sum_{j \in I} \kappa_j \kappa_j^*
+    \ =\ \sum_{j \in I} f' \kappa_j \kappa_j^*
+    \ =\ \sum_{j \in I} f_j \kappa_j^*
+    \ =\ \sum_{j \in I} f \kappa_j \kappa_j^*
+    \ =\ f,
 \end{equation*}

### 2018-05-01 direct_integrals.tex
some small additions and corrections

@@
 \usepackage{xparse} % new commands with more than one optional argument
+\usepackage{fixltx2e} % for \textsubscript
 \newcommand{\unlim}{\qopname\relax m{unlim}}
@@
 \newcommand{\Rep}[1]{\mathsf{Rep}(#1)}		% cat normal Hilbert space reps of a C$^*$-algebra
-\newcommand{\NRep}[1]{\mathsf{NRep}(#1)}	% cat of normal Hilbert space reps of a W$^*$-algebra
+\newcommand{\NRep}[1]{\mathsf{nRep}(#1)}	% cat of normal Hilbert space reps of a W$^*$-algebra
 \newcommand{\id}[1]{1_{#1}}			% identity morphism
@@
 \begin{thm}
-The following are equivalent:
+The following are equivalent for an object $A$:
 \begin{enumerate}
-\item\label{universalprop} $\bigoplus_i A_i$ is an $I$-indexed direct sum.
+\item\label{universalprop} $A$ is an $I$-indexed direct sum $\bigoplus_i A_i$.
 \item\label{universalprop2} 
 There is a family of
-        morphisms~$\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$
+        morphisms~$\left(\kappa_j : A_j \to A\right)_{j\in I}$
         with~$\sum_{j \in I} \kappa_j \kappa_j^* < \infty$
@@
         with~$\sum_{j \in I}f_j f_j^* < \infty$,
-        there is a unique~$f\colon \bigoplus_i A_i \to B$
+        there is a unique~$f\colon A \to B$
         such that $f \kappa_j = f_j$ for all~$j \in I$, and $\| f \|^2 = \| \sum_j f_j f_j^* \|$.
-\item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to \bigoplus_i A_i\right)_{j\in I}$ such that $\kappa_j^*\kappa_k = \delta_{jk}$, and
+\item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to A\right)_{j\in I}$ such that $\kappa_i^*\kappa_j = \delta_{ij}$, and
 \beq
 \label{complete}
-        \sum_{j\in I} \kappa_j \kappa_j^* \ =\  1_{\bigoplus_i A_i}.
+        \sum_{j\in I} \kappa_j \kappa_j^* \ =\  1_A.
 \eeq
@@
 
+One can also consider each of these three statements as a structure rather than a mere property, and then the proof establishes a bijection between these structures.
+
 \begin{proof}
-We write $A$ as shorthand for $\bigoplus_i A_i$.
+\ref{universalprop2} is a simple reformulation of the universal property \ref{universalprop} via the Yoneda lemma, as follows.
+Given a natural isometric isomorphism $\cat{C}(A,B) \cong \bigoplus_i \cat{C}(A_i,B)$, we define the family $(\kappa_i)$ to be the image of $\id{A}$.
+Then for any $f : A \to B$, the naturality diagram
+\[\begin{tikzcd}
+	\cat{C}(A,A) \ar{r}{\cong} \ar{d}{f\circ -} & \bigoplus_i \cat{C}(A_i,A) \ar{d}{(f\circ -)} \\
+	\cat{C}(A,B) \ar{r}{\cong} & \bigoplus_i \cat{C}(A_i,B)
+\end{tikzcd}\]
+shows that the universal property takes $f$ to the family $(f_j) := (f\kappa_j)$.
+By the definition of the norm on $\bigoplus_i \cat{C}(A_i,B)$ and the assumption that the universal property is an isometric isomorphism, we indeed have $\| f \|^2 = \left\| \sum_j f_j f_j^* \right\|$, which also gives $\sum_j \kappa_j \kappa_j^* < \infty$ in the case $f = \id{A}$.
+Conversely, it is clear that any family $(\kappa_j)$ as in~\ref{universalprop2} gives an isometric isomorphism $\cat{C}(A,B) \cong \bigoplus_i \cat{C}(A_i,B)$ that is natural in $B$.
 
-Assuming \ref{universalprop}, we define $(\kappa_j : A_j \to A)$ to be the family that corresponds to $\id{A}$ under the given isomorphism $\bigoplus_i \cat{C}(A_i,A) \cong \cat{C}(A,A)$.
-For fixed $j\in J$, also the family $(\delta_{ij} : A_i \to A_j)$ corresponds to some morphism $\pi_j : A \to A_j$.
-Applying naturality with respect to $\pi_j$ gives the diagram
-\[\begin{tikzcd}
-	\bigoplus_i \cat{C}(A_i,A) \ar{r}{\cong} \ar{d}{(\pi_j\circ -)_i} & \cat{C}(A,A) \ar{d}{\pi_j\circ -} \\
-	\bigoplus_i \cat{C}(A_i,A_j) \ar{r}{\cong} & \cat{C}(A,A_j)
-\end{tikzcd}\]
-Starting with the family $(\kappa_i)$ on the upper left, we get $(\pi_j\circ\kappa_i)$ by going down, or alternatively $(\delta_{ij})$ by going around; therefore we have $\pi_j \kappa_i = \delta_{ij}$.
-By the isometric isomorphism assumption, we moreover have $\| \id{A} \| = \| \sum_i \kappa_i \kappa_i^* \|$, and therefore $\| \kappa_j \| \leq 1$; and similarly $\| \pi_j \| = \| \id{A} \| \leq 1$.
+We now derive \ref{biprodsum} from \ref{universalprop2}.
+Applying the assumption on norms with $(f_j) = (\kappa_j)$ gives $\| \sum_i \kappa_i \kappa_i^* \| = \| \id{A} \|$, and therefore $\| \kappa_j \| \leq 1$.
+For fixed $j\in J$, the family $(\delta_{ij} : A_i \to A_j)$ corresponds to some morphism $\pi_j : A \to A_j$ such that $\pi_j\circ\kappa_i = \delta_{ij}$.
+Again by the isometry property, we have $\| \pi_j \| = \| \id{A} \| \leq 1$.
 By Lemma~\ref{isometrylemma}, this implies $\pi_j = \kappa_j^*$.
@@
 In order to also prove the completeness relation~\eqref{complete}, let $p : A\to A$ be any other projection that is orthogonal to each of the $\kappa_j \kappa_j^*$; we need to show $p = 0$.
-Again by naturality, under the isomorphism $\bigoplus_i \cat{C}(A_i,A)$, the new $p$ corresponds to the family $(p\kappa_i) = (0)$.
-This implies indeed $p=0$.
+Again by naturality, under the isomorphism $\cat{C}(A,A) \cong \bigoplus_i \cat{C}(A_i,A)$, the new $p$ corresponds to the family $(p\kappa_i) = (0)$, where $p\kappa_i = 0$ follows from the assumed orthogonality.
+This implies indeed $p=0$ by the uniqueness part of the universal property.
 
-Assuming \ref{biprodsum},
-let~$\left(f_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
+We finally show that \ref{biprodsum} implies \ref{universalprop2}. So let~$\left(f_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
         with~$\sum_{j \in I}f_jf_j^* < \infty$.
@@
     with~$\cat{C}(B,B)$-valued inner
-    product~$\langle g, f\rangle \equiv g^*f$.\cite[2.15]{wstarcat}
+    product~$\langle g, f\rangle \equiv g^*f$~\cite[2.15]{wstarcat}.
 The self-duality is equivalent to the fact
@@
     for normal states~$\omega$ on~$\cat{C}(B,B)$~\cite[\S149\textsubscript{V}]{bas}.
-    Pick any normal state~$\omega$ on~$\cat{C}(B,B)$.
+    Pick any normal state~$\omega$ on~$\cat{C}(B,B)$
         and finite subset~$S \subseteq I$.  We have
@@
 From this and~$\sum_{j \in I} f_jf_j^* < \infty$,
-    it follows that~$(\sum_{j \in S} \kappa_j f_j^*)_S$
-    is ultranorm Cauchy.
-    Define~$f \equiv ( \sum_{j \in I} \kappa_jf_j^* )^*$,
+    it follows that the net~$\left(\sum_{j \in S} \kappa_j f_j^*\right)_S$
+    is ultranorm Cauchy as $S$ ranges over all finite subsets of $I$.
+    Define~$f := \left( \sum_{j \in I} \kappa_jf_j^* \right)^*$,
         where the sum is understood to converge ultranorm.
@@
             and so~$f \kappa_j = f_j$ as desired.
+\todo[inline]{Show equality of norms}
 Finally, to show uniqueness,
@@
 With similar reasoning as before,
-        the map~$f \mapsto f' f$ is ultranorm continuous,
+        the composition map~$g \mapsto f' g$ is ultranorm continuous,
         hence
@@
 For finite $I$, these direct sums are also dagger limits in the sense of~\cite{daglims}.
-For infinite $I$, this is no longer the case, since infinite direct sums are not even limits in the usual $\cat{Set}$-enriched sense; compare~\cite[Theorem~5.2]{daglims}.
-Hence our definition are better adapted to the context of W$^*$-categories, such as $\cat{Hilb}$, than those of~\cite{daglims}.
+For infinite $I$, this is no longer the case, since infinite direct sums are not even limits; compare the degeneracy result~\cite[Theorem~5.2]{daglims}, which indicates that infinite dagger products cannot be expected to exist.
+Thus our Definition~\ref{directsumdef} seems better adapted to the context of Hilbert spaces or other W$^*$-categories than those of~\cite{daglims}.
 \tob{We can tone down the last sentence a bit if it's too demeaning}
@@
 
-\newcommand{\alg}{N}	% a W$^*$-algebra
+As a simple consequence, we have a generalization of~\cite[Proposition~4.9]{Rieffel1974}.
+
+\begin{cor}
+A normal $*$-functor automatically preserves direct sums.
+\end{cor}
+
+\newcommand{\alg}{\cat{N}}	% a W$^*$-algebra
 

### 2018-05-01 direct_integrals.tex
historical comment

@@
 
+As far as we know, the first definition of infinite direct sums has been given in~\cite[p.~100]{wstarcat} in the form~\ref{biprodsum}.
+From our perspective, it is preferable to replace this definition by a universal property as in Definition~\ref{directsumdef}, and regard the original definition as an alternative characterization instead.
+
 \begin{proof}

### 2018-05-03 direct_integrals.tex
comment on the choice of measures

@@
 
+\todo[inline]{Tob: I now think that the ``right'' definition may be in terms of \emph{semifinite} measures, by analogy with normal semifinite traces.
+My idea is that measurable families of objects are in some sense the objects of some ``$L^\infty(X,\cat{C})$'', which could perhaps be a W$^*$-category itself, along the lines of what Yetter~\cite{measureablecats} has proposed.
+This analogy with W$^*$-algebras makes me think that semifinite measures may be what we want}
+
 \begin{defn}

### 2018-06-04 direct_integrals.tex
add missing norm proof and some polish

@@
 \newcommand{\unlim}{\qopname\relax m{unlim}}
+\newcommand{\uwlim}{\qopname\relax m{uwlim}}
 
@@
     Then~$ta = ta(1-s)$ and~$a^*t = (1-s)a^* t$.
-    Hence~$a^* ta = (1-s)a^*ta(1-s) \leq (1-s)$.
+    Hence~$a^* ta = (1-s)a^*ta(1-s) \leq 1-s$, as desired.
 A similar argument proves~$as a^* \leq 1-t \Leftrightarrow sa^*t=0$.
@@
 Let~$a \colon A \leftrightarrows B \colon b$ be any two morphisms
-    such that~$\|a \| \leq 1$ and $\| b\| \leq 1$ and~$ab=1$.
+    such that~$\|a \| \leq 1$, $\| b\| \leq 1$ and~$ab=1$.
     Then~$b$ is an isometry with~$a=b^*$.
@@
     Clearly~$bb^*bb^* = b 1 b^* = bb^*$, so~$b$ is an isometry.
-    Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0$
+    Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0$.
     So~$a^* (1-0) a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
@@
 \[
-	\bigoplus_i \cat{C}(B, A_i) := \left\{ (f_i : B \to A_i)_{i\in I} \Biggm\vert \sum_i f_i^* f_i < \infty \right\},
+	\bigoplus_i \cat{C}(B, A_i) \ := \ \bigl\{\, (f_i : B \to A_i)_{i\in I} \bigm\vert \sum_i f_i^* f_i < \infty \,\bigr\},
 \]
-with norm
+where the sum is taken in the ultraweak topology\footnote{
+    For positive elements~$(x_i)_{i \in I}$ of a W$^*$-algebra,
+        convergence and value
+        of the sum~$\sum_{i \in I} x_i $
+        coincides in the ultraweak and ultrastrong topology
+        and, in fact, equals the supremum of the partial sums.}
+    and where we use the norm
 \[
-	\| (f_i)_{i \in I} \| := \left\| \sum_i f_i^* f_i \right\|^{1/2},
+	\| (f_i)_{i \in I} \|\  := \ \Bigl\| \,\sum_i f_i^* f_i \,\Bigr\|^{1/2},
 \]
@@
         there is a unique~$f\colon A \to B$
-        such that $f \kappa_j = f_j$ for all~$j \in I$, and $\| f \|^2 = \| \sum_j f_j f_j^* \|$.
+        such that $f \kappa_j = f_j$ for all~$j \in I$, and $\| f \|^2 = \| \sum_{j\in I} f_j f_j^* \|$.
 \item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to A\right)_{j\in I}$ such that $\kappa_i^*\kappa_j = \delta_{ij}$, and
@@
             and so~$f \kappa_j = f_j$ as desired.
-\todo[inline]{Show equality of norms}
+Next we will show that~$\|f \|^2 = \|\sum_j f_j f_j^*\|$.
+    By definition of~$f^*$ and \cite[\S148\textsubscript{V}]{bas}
+         we have~
+ \begin{equation*}
+         \langle f^*, f^*\rangle
+         \ =\  \uwlim_{S\subseteq I\text{ finite}} \, \Bigl\langle \sum_{j \in S} \kappa_j f^*_j,
+                        \sum_{j \in S}\kappa_j f^*_j \Bigr\rangle
+         \ =\  \uwlim_{S\subseteq I\text{ finite}} \sum_{j \in S} f_j f_j^*
+        \ = \ \sum_j f_j f_j^*
+ \end{equation*}
+and so~$\| f \|^2 = \|f^*\|^2 = \|\langle f,f \rangle \|
+                = \|\sum_j f_j f_j^* \|$ as desired.
 Finally, to show uniqueness,

### 2018-06-07 direct_integrals.tex
minor edits

@@
 %%% draft stuff
-\usepackage[draft]{showkeys}
+%\usepackage[draft]{showkeys}
 \usepackage{todonotes}
@@
 
-We will need the full power of this result only in the next section, but the following special case is generally useful for the definition of universal properties in C$^*$-categories:
+We will need the full power of this observation only in the next section. However, the following special case is generally useful for the definition of universal properties in C$^*$-categories:
 
@@
 The self-duality is equivalent to the fact
-    that~$\cat{C}(B,A)$ is complete in the \emph{ultranorm} topology;
-    that is, the uniformity
+    that~$\cat{C}(B,A)$ is complete in the \emph{ultranorm topology};
+    this is the topology
     generated by the seminorms~$\| f \|_\omega \equiv \omega(\langle f,f\rangle)^\frac{1}{2}$
-    for normal states~$\omega$ on~$\cat{C}(B,B)$~\cite[\S149\textsubscript{V}]{bas}.
+    indexed by normal states~$\omega$ on~$\cat{C}(B,B)$~\cite[\S149\textsubscript{V}]{bas}. We now turn to the construction of $f : A \to B$.
     Pick any normal state~$\omega$ on~$\cat{C}(B,B)$

### 2018-06-08 direct_integrals.tex
Lemma 3.1 (completeness)

@@
 
-The proof is analogous to the standard proof showing that $\ell^2(I)$ is a Hilbert space. \tob{I haven't done this yet, so for now this statement is wishful thinking}
+The proof is analogous to the standard proof showing that $\ell^2(I)$ is a Hilbert space (and generalizes it).
 
 \begin{proof}
-Suppose that we are given a Cauchy sequence of families $( ( f_i^{(n)} )_{i\in I} )_{n\in\N}$.
+Suppose that we are given a Cauchy sequence of families $(f^{(n)})_{n\in\N} = ( ( f_i^{(n)} )_{i\in I} )_{n\in\N}$.
+For given $i$, projecting to $\cat{C}(A_i,B)$ is bounded, and therefore $( f_i^{(n)} )_{n\in\N}$ is Cauchy as well; let us denote the limit by $f_i : A_i \to B$.
+We will now show that $f \in \oplus_i \cat{C}(A_i,B)$ for $f := (f_i)_{i\in I}$.
+The Cauchy assumption implies that the sequence $(f^{(n)})_{n\in\N}$ is uniformly bounded by some constant $C$,
+\[
+	\left\| \sum_i (f_i^{(n)})^* f_i^{(n)} \right\|^{1/2} \leq C.
+\]
+Restricting to a finite partial sum over $S\subseteq I$ and taking $n\to\infty$ shows that
+\[
+	\left| \sum_{i\in S} f_i^* f_i \right\|^{1/2} \leq C
+\]
+as well. Since $S$ was arbitrary, we conclude that $\| f \| \leq C$.
 
-\todo{Complete this! Showing that $(f_i^{(n)})_{n\in\N}$ is Cauchy for each $i$ needs some form of Cauchy-Schwarz}
+The convergence $f_i \to f$ works similarly.
+By assumption, for every $\eps > 0$ we have $\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for almost all $n$ and $m$.
+In particular, we know that for any finite $S \subseteq I$,
+\[
+	\left\| \sum_{i\in S} (f_i^{(n)} - f_i^{(m)})^* (f_i^{(n)} - f_i^{(m)}) \right\|^{1/2} \leq \eps
+\]
+for almost all $n$ and $m$. Taking $m\to\infty$ gives
+\[
+	\left\| \sum_{i\in S} (f_i^{(n)} - f_i)^* (f_i^{(n)} - f_i) \right\|^{1/2} \leq \eps.
+\]
+Since $S$ was arbitrary, we conclude $\| f^{(n)} - f \| \leq \eps$ for almost all $n$.
 \end{proof}

### 2018-06-20 direct_integrals.tex
switch to localizable measure spaces

@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\Rplusext}{\overline{\mathbb{R}}_+} % extended nonnegative reals, as where measures take values
 \newcommand{\C}{\mathbb{C}}
@@
 \newcommand{\lin}[1]{\mathrm{lin}(#1)}
-\newcommand{\dom}{\mathrm{dom}}
+\newcommand{\dom}[1]{\mathrm{dom}(#1)} 		% domain of a morphism
+\newcommand{\cod}[1]{\mathrm{cod}(#1)} 		% codomain of a morphism
 \newcommand{\supp}{\mathrm{supp}}
+\newcommand{\esssup}{\mathrm{ess\, sup}} 	% essential supremum
 
@@
 
+\newcommand{\LMeas}{\cat{LocMeas}} 	% category of localizable measurable spaces
+\newcommand{\Null}{\mathcal{N}} 	% null sets
+
+We let $\LMeas$ be the category of localizable measurable spaces~\cite{nlab:localizable_measure}.
+So objects of $\LMeas$ are triples $(X,\Sigma,\Null)$, where $(X,\Sigma)$ is a $\sigma$-algebra and $\Null\subseteq\Sigma$ is a $\sigma$-ideal, whose elements are called \emph{null sets}, such that $\Sigma/\Null$ is complete as a Boolean algebra.
+Measurable sets with null complement are also called \emph{full sets}.
+We usually just write $X$ for a localizable measurable space, and indicate its $\sigma$-algebra and $\sigma$-ideal of measurable and null sets by $\Sigma_X$ and $\Null_X$, respectively.
+The morphisms of $\LMeas$ are the measurable maps $f : X \to Y$ modulo null sets, i.e.~$f,g : X\to Y$ represent the same morphism if $\{ x\in X \: | \: f(x) \neq g(x) \} \in \Null_X$.
+
+A \emph{measure} $\mu$ on localizable measurable space $X$ is a measure $\mu : \Sigma_X \to \Rplusext$ in the usual sense, but such that every null set has zero measure.
+It hence descends to a map $\Sigma_X / \Null_X \to \Rplusext$ satisfying the inclusion-exclusion relation.
+
+\tob{Is this map $\Sigma_X/\Null_X \to \Rplusext$ necessarily Scott continuous, so that it becomes a continuous valuation?}
+
 \begin{defn}
-A \emph{measurable category} $\cat{C}$ is a category internal to measurable spaces.
-A \emph{measured family of objects} indexed by a measure space $(X,\Sigma,\mu)$ is a measurable function $A : X\to\Obj{\cat{C}}$.
+A \emph{measurable category} $\cat{C}$ is a category internal to (potentially large) measurable spaces.
 \end{defn}
+
+Given $X\in\LMeas$, we define a category $L^\infty(X,\cat{C})$ as follows.
+Its objects are measurable functions $X \to \Obj{\cat{C}}$, which we denote as indexed families $(A_x)_{x\in X}$ varying measurably in $x$.
+The morphisms $(A_x) \to (B_x)$ are bounded measurable functions $f : X \to \Mor{\cat{C}}$ such that $\dom{f_x} = A_x$ and $\cod{f_x} = B_x$ for all $x$, again modulo almost everywhere equality.
+
+\begin{prop}
+$L^\infty(X,\cat{C})$ is a W$^*$-category as well.
+\end{prop}
+
+So the categories $L^\infty(X,\Hilb)$ are our version of Yetter's \emph{categories of measurable fields of Hilbert spaces}.
+
+\todo[inline]{Can we make the comparison more precise?}
+
+\begin{proof}
+It is clear that $L^\infty(X,\cat{C})$ inherits the antilinear involution from $\cat{C}$. The norm of a morphism is defined as
+\[
+	\| (f_x) \| := \esssup_x \| f_x \|,
+\]
+which is clearly well-defined upon changing the representative.
+To show that each hom-set is also complete as a normed space, let $( (f_x)^{(n)})_{n\in\N}$ be a Cauchy sequence of (representatives of) morphisms.
+We show that the collection of all $x$ for which $(f_x^{(n)})_{n\in\N}$ is Cauchy is full.
+Since countable intersections of full sets are full, It is enough to prove that for every $\eps > 0$, the set of $x$ for which $\| f_x^{(n)} - f_x^{(m)} \| < \eps$ holds eventually is full.
+But by assumption, we know that $\esssup_x \| f_x^{(n)} - f_x^{(m)} \| < \eps$ eventually, which is enough.
+Therefore we can define a candidate representative of a limit morphism as $(f_x)$, where $f_x$ is the above limit on those $x$ where the family is Cauchy, and e.g.~zero outside of that.
+It is straightforward to show that this is the limit.
+
+The Banach algebra axiom and the C$^*$-identity follow directly from those of $\cat{C}$.
+
+\todo[inline]{How about preduals for the hom-sets?}
+\end{proof}
 
@@
 \todo[inline]{Is this the ``right'' definition? What is the relation to~\cite{measureablecats}?}
-
-\todo[inline]{Tob: I now think that the ``right'' definition may be in terms of \emph{semifinite} measures, by analogy with normal semifinite traces.
-My idea is that measurable families of objects are in some sense the objects of some ``$L^\infty(X,\cat{C})$'', which could perhaps be a W$^*$-category itself, along the lines of what Yetter~\cite{measureablecats} has proposed.
-This analogy with W$^*$-algebras makes me think that semifinite measures may be what we want}
 

### 2018-06-29 direct_integrals.tex
I've added some stuff about compact measure spaces and a few corrections.

@@
 \newcommand{\id}[1]{1_{#1}}			% identity morphism
+\newcommand{\Kay}{\mathcal{K}} % I use a convention that the name of a letter spelt phonetically signifies its mathcal version. But semantically, this is a compact class in a measure space.
+\newcommand{\powerset}{\mathcal{P}}
 
@@
 We let $\LMeas$ be the category of localizable measurable spaces~\cite{nlab:localizable_measure}.
-So objects of $\LMeas$ are triples $(X,\Sigma,\Null)$, where $(X,\Sigma)$ is a $\sigma$-algebra and $\Null\subseteq\Sigma$ is a $\sigma$-ideal, whose elements are called \emph{null sets}, such that $\Sigma/\Null$ is complete as a Boolean algebra.
+So objects of $\LMeas$ are triples $(X,\Sigma,\Null)$, where $\Sigma$ is a $\sigma$-algebra on $X$, and $\Null\subseteq\Sigma$ is a $\sigma$-ideal, whose elements are called \emph{null sets}, such that $\Sigma/\Null$ is complete as a Boolean algebra, and there exists a semifinite \cite[211F]{fremlin2} measure $\mu : \Sigma \rightarrow [0,\infty]$ such that $\Null = \mu^{-1}(0)$.
+These are what we get if we take a localizable measure space $(X,\Sigma,\mu)$  \cite[211G]{fremlin2} and only keep the information about which sets are null.
+Localizable measure spaces were introduced by Segal \cite{segal} as a characterization of the measure spaces such that $L^\infty(X)$, the C$^*$ of essentially bounded measurable functions modulo nullsets, is a W$^*$-algebra.
+
 Measurable sets with null complement are also called \emph{full sets}.
 We usually just write $X$ for a localizable measurable space, and indicate its $\sigma$-algebra and $\sigma$-ideal of measurable and null sets by $\Sigma_X$ and $\Null_X$, respectively.
-The morphisms of $\LMeas$ are the measurable maps $f : X \to Y$ modulo null sets, i.e.~$f,g : X\to Y$ represent the same morphism if $\{ x\in X \: | \: f(x) \neq g(x) \} \in \Null_X$.
+The morphisms of $\LMeas$ are equivalence classes of measurable maps $f : X \to Y$ such that for all $N \in \Null_Y$, $f^{-1}(N) \in \Null_X$\todo[inline]{We need to be using compact strictly localizable spaces to make this definition sufficient to define a normal map between W*-algebras.}, under the following equivalence relation: $f \sim g$ iff for all $T \in \Sigma_Y$ $f^{-1}(T) \triangle g^{-1}(T) \in \Null_X$.
+If $f$ and $g$ agree almost everywhere, then $f \sim g$, and the two notions are equivalent if $(Y,\Sigma_Y)$ is countably separated \cite[343F]{fremlin3}, but not in general \cite[343I]{fremlin3}.
 
 A \emph{measure} $\mu$ on localizable measurable space $X$ is a measure $\mu : \Sigma_X \to \Rplusext$ in the usual sense, but such that every null set has zero measure.
-It hence descends to a map $\Sigma_X / \Null_X \to \Rplusext$ satisfying the inclusion-exclusion relation.
+It therefore descends to a countably additive map $\Sigma_X / \Null_X \to \Rplusext$, such that every non-zero element of $\Sigma_X / \Null_X$ has non-zero measure.
+As $\Sigma_X / \Null_X$ is a complete Boolean algebra, such a map is necessarily completely additive, not just countably additive \cite[321E]{fremlin2}.
 
-\tob{Is this map $\Sigma_X/\Null_X \to \Rplusext$ necessarily Scott continuous, so that it becomes a continuous valuation?}
+Unfortunately, there exist localizable measurable spaces $X$, $Y$ such that $L^\infty(X) \cong L^\infty(Y)$ but there is no measurable map from $X$ to $Y$.
+To get out of this difficulty, we restrict ourselves to localizable measurable spaces arising from complete strictly localizable compact measure spaces.
+
+To define these notions, we first define the finite intersection property, and what a compact class is.
+Let $X$ be a set, and $\Kay \subseteq \powerset(X)$ be a family of sets.
+We say it has the \emph{finite intersection property} if for each finite subfamily $(K_i)_{i \in I}$, $K_i \in \Kay$, we have $\bigcap_{i \in I} K_i \neq 0$.
+We say that $\Kay$ is a \emph{compact class} if for any subfamily $\Kay' \subseteq \Kay$ having the finite intersection property, $\bigcap \Kay' \neq \emptyset$.
+By taking the complement of every set in the open cover definition of a compact space, the reader may recognize that the closed sets of any compact topological space form a compact class.
+Slightly less trivially, the compact subsets of any Hausdorff space form a compact class.
+
+A measure space $(X,\Sigma,\mu)$ is \emph{compact} \cite[342A (c)]{fremlin3} if there exists a compact class $\Kay$ such that for all $S \in \Sigma$
+\[
+\mu(S) = \sup \{ \mu(K) \mid K \in \Kay \cap \Sigma \text{ and } K \subseteq S \}.
+\]
+This condition is known as being \emph{inner regular with respect to} $\Kay$.
+Therefore any Radon measure on a Hausdorff space $X$ is compact, because we can take $\Kay$ to be the set of (topologically) compact subsets of $X$.
+So familiar examples, such as Lebesgue measure on $\R^n$, the Haar measure on a locally compact group, or any $\sigma$-finite measure on a standard Borel space, are compact.
 

### 2018-07-06 direct_integrals.tex
adjustments as per Robert's comments

--- 2018-06-08 direct_integrals.tex
@@
 \usepackage{todonotes}
-\newcommand{\tob}[1]{\todo[color=blue!40,inline]{#1}}
+\newcommand{\tob}[1]{\todo[color=blue!40,inline]{\textbf{T}: #1}}
 
@@
 
-\section{Direct integrals}
+\section{The universal property of direct integrals}
 
@@
 
-\begin{defn}
-A \emph{measurable category} $\cat{C}$ is a category internal to measurable spaces.
-A \emph{measured family of objects} indexed by a measure space $(X,\Sigma,\mu)$ is a measurable function $A : X\to\Obj{\cat{C}}$.
-\end{defn}
+\newcommand{\LMeas}{\cat{LocMeas}} 	% category of localizable measure spaces
+\newcommand{\Null}{\mathcal{N}} 	% null sets
+
+\tob{The precise definitions are not set in stone, and should be adjusted such that the measure theory works out. So far it's more of a sketch}
+
+\subsection{Preliminaries}
+
+We begin with some measure-theoretic preliminaries, following Fremlin~\cite{fremlin2,fremlin3}.
+
+Given a measure space $(X,\Sigma,\mu)$, the \emph{null sets} are the sets of zero measure, and the \emph{full sets} are their complements.
+The measure space is \emph{localizable} if $\Sigma$ modulo the null sets is complete as a Boolean algebra, which means that every collection of measurable sets has an \emph{essential supremum}.
+Localizable measure spaces were introduced by Segal \cite{segal} as a characterization of the measure spaces such that $L^\infty(X)$, the C$^*$-algebra of essentially bounded measurable functions modulo null sets, is a W$^*$-algebra.
+
+With $\Null$ denoting the $\sigma$-ideal of null sets, the measure descends to a countably additive map $\Sigma / \Null \to \Rplusext$, such that every non-zero element of $\Sigma / \Null$ has non-zero measure.
+As $\Sigma / \Null$ is a complete Boolean algebra, such a map is necessarily completely additive, not just countably additive \cite[321E]{fremlin2}.
+
+\tob{So we can e.g.
+take the measurable space $(S,2^S)$ for some uncountable $S$, and equip it with the measure $\mu$ which vanishes on any countable set, and is $\infty$ on all uncountable sets.
+This is a localizable measure space, right?
+I guess that it's not compact though, and clearly not strictly localizable.
+Is this a good (counter-)example for us?}
+
+Unfortunately, there exist localizable measure spaces $X$, $Y$ such that $L^\infty(X) \cong L^\infty(Y)$ but there is no measurable map from $X$ to $Y$ \todo{example?}.
+To get out of this difficulty, we restrict ourselves to localizable measure spaces arising from \emph{complete strictly localizable compact} measure spaces.
+
+To define these notions, we first define the finite intersection property, and what a compact class is.
+Let $X$ be a set, and $\Kay \subseteq \powerset(X)$ be a family of sets.
+We say it has the \emph{finite intersection property} if for each finite subfamily $(K_i)_{i \in I}$, $K_i \in \Kay$, we have $\bigcap_{i \in I} K_i \neq 0$.
+We say that $\Kay$ is a \emph{compact class} if for any subfamily $\Kay' \subseteq \Kay$ having the finite intersection property, $\bigcap \Kay' \neq \emptyset$.
+For example, the closed sets of any compact topological space form a compact class.
+Slightly less trivially, the compact subsets of any Hausdorff space form a compact class.
+
+A measure space $(X,\Sigma,\mu)$ is \emph{compact} \cite[342A(c)]{fremlin3} if there exists a compact class $\Kay$ such that for all $S \in \Sigma$
+\[
+\mu(S) = \sup \{ \mu(K) \mid K \in \Kay \cap \Sigma \text{ and } K \subseteq S \}.
+\]
+This condition is known as being \emph{inner regular with respect to} $\Kay$.
+Therefore any Radon measure on a Hausdorff space $X$ is compact, because we can take $\Kay$ to be the set of (topologically) compact subsets of $X$.
+So familiar examples, such as Lebesgue measure on $\R^n$, the Haar measure on a locally compact group, or any $\sigma$-finite measure on a standard Borel space, are compact.
+
+A measure space $(X,\Sigma,\mu)$ is \emph{strictly localizable} if it can be written as the (typically infinite) disjoint union of finite localizable measure spaces~\cite[211E]{fremlin2}.
+
+We write $\LMeas$ for the category where:
+\begin{enumerate}
+\item Objects are measure spaces which are complete, strictly localizable and compact;
+\item Morphisms are equivalence classes of measurable maps $f : X \to Y$ with $\mu_Y \ll f_* \mu_X$ \todo{or something like truly continuous?}, where for measurable $f,g : X \to Y$, we put $f \sim g$ iff for all $T \in \Sigma_Y$, we have $f^{-1}(T) \triangle g^{-1}(T) \in \Null_X$.
+\todo{Do we also want some compatibility with the compact classes?}
+\end{enumerate}
+
+If $f$ and $g$ agree almost everywhere, then $f \sim g$. The converse also holds if $(Y,\Sigma_Y)$ is countably separated \cite[343F]{fremlin3}, but not in general \cite[343I]{fremlin3}.
+
+\subsection{Measurable families of objects}
+
+Now let $\cat{C}$ be a W$^*$-category, and $X \in \LMeas$. Then we define a category $L^\infty(X,\cat{C})$ as follows:
+
+\begin{enumerate}
+\item Objects are maps $A : X \to \cat{C}, \: x\mapsto A_x$, which are measurable in the sense that there is a measurable partition of $X$ such that $A$ is constant on each part;
+\item Morphisms $f : A\to B$ are equivalence classes of bounded families $(f_x : A_x\to B_x)_{x\in X}$, which depend measurably \todo{work out what exactly this should mean such that composition makes sense} on $x$, and where $f \sim g$ if upon restriction to any measurable set on which both $A$ and $B$ are constant, we have $f^{-1}(T) \triangle g^{-1}(T) \in \Null_X$ for every measurable $T \subseteq \cat{C}(A,B)$.
+\end{enumerate}
+
+In particular, this means that two objects $A$ and $B$ which differ only on a null set are isomorphic.
+
+\begin{prop}
+$L^\infty(X,\cat{C})$ is a W$^*$-category as well.
+\end{prop}
+
+So the categories $L^\infty(X,\Hilb)$ are our version of Yetter's \emph{categories of measurable fields of Hilbert spaces}.
+
+\todo[inline]{Can we make the comparison more precise?}
+
+\begin{proof}
+It is clear that $L^\infty(X,\cat{C})$ inherits the antilinear involution from $\cat{C}$. The norm of a morphism is defined as
+\[
+	\| (f_x) \| := \esssup_x \| f_x \|,
+\]
+which is clearly well-defined upon changing the representative.
+To show that each hom-set is also complete as a normed space, let $( (f_x)^{(n)})_{n\in\N}$ be a Cauchy sequence of (representatives of) morphisms.
+We show that the collection of all $x$ for which $(f_x^{(n)})_{n\in\N}$ is Cauchy is a full set.
+Since countable intersections of full sets are full, It is enough to prove that for every $\eps > 0$, the set of $x$ for which $\| f_x^{(n)} - f_x^{(m)} \| < \eps$ holds eventually is a full set.
+But by assumption, we know that $\esssup_x \| f_x^{(n)} - f_x^{(m)} \| < \eps$ eventually, which is enough.
+Therefore we can define a candidate representative of a limit morphism as $(f_x)$, where $f_x$ is the above limit on those $x$ where the family is Cauchy, and e.g.~zero outside of that.
+It is straightforward to show that this is the limit.
+
+The submultiplicativity of the norm and the C$^*$-identity follow directly from those of $\cat{C}$.
+
+\todo[inline]{Try to construct preduals for the hom-sets in terms of spaces like $L^1(X,\cat{C}(A,B)_*)$?}
+\end{proof}
 
@@
 \[
-	\int_X \cat{C}(A_x,-) \: : \: \cat{C}\to\Set
+	\int_X \cat{C}(A_x,-) \, d\mu(x) \: : \: \cat{C}\to\Set
 \]
@@
 \[
-	\int_X \cat{C}(A_x , B) := \left\{ f : X\to\Mor{\cat{C}} \textrm{ measurable} \Biggm\vert f_x : A_x\to B,\: \int_X f_x f_x^* \,dx < \infty \right\},
+	\int_X \cat{C}(A_x , B) \, d\mu(x) := \left\{ f : X\to\Mor{\cat{C}} \textrm{ measurable} \Biggm\vert f_x : A_x\to B,\: \int_X f_x f_x^* \,dx < \infty \right\},
 \]
@@
 
+\begin{defn}
+A \emph{direct integral} of the family $(A_x)_{x\in X}$ is now an object $\int_X A_x \, d\mu(x)$ together with a natural bijection
+\[
+	\cat{C}\left( \int_X A_x \, d\mu(x), -\right) \cong \int_X \cat{C}(A_x,-) \, d\mu(x).
+\]
+\end{defn}
+
 \todo[inline]{Is this the ``right'' definition? What is the relation to~\cite{measureablecats}?}
-
-\todo[inline]{Tob: I now think that the ``right'' definition may be in terms of \emph{semifinite} measures, by analogy with normal semifinite traces.
-My idea is that measurable families of objects are in some sense the objects of some ``$L^\infty(X,\cat{C})$'', which could perhaps be a W$^*$-category itself, along the lines of what Yetter~\cite{measureablecats} has proposed.
-This analogy with W$^*$-algebras makes me think that semifinite measures may be what we want}
 
@@
 
+\begin{thm}
+Suppose that $\cat{C}$ has direct integrals. Then every $\int_X A_x \, d\mu(x)$ carries a normal representation of $L^\infty(X)$, and we have a functor
+\[
+	L^\infty(X,\cat{C}) \to \NRep{L^\infty(X),\cat{C}} , \qquad (A_x) \mapsto \int_X A_x \, d\mu(x),
+\]
+and this is an equivalence of W$^*$-categories.
+\end{thm}
+
 \bibliographystyle{plain}

### 2018-07-10 direct_integrals.tex
Fix to definition of localizable and examples.

@@
 
-Given a measure space $(X,\Sigma,\mu)$, the \emph{null sets} are the sets of zero measure, and the \emph{full sets} are their complements.
-The measure space is \emph{localizable} if $\Sigma$ modulo the null sets is complete as a Boolean algebra, which means that every collection of measurable sets has an \emph{essential supremum}.
+Given a measure space $(X,\Sigma,\mu)$, the \emph{null sets} are the sets of zero measure, forming a $\sigma$-ideal in $\Sigma$, and the \emph{full sets} are their complements.
+We say a set $S \in \Sigma$ has \emph{finite measure} if $\mu(S) < \infty$.
+If $X$ itself, and therefore all sets in $\Sigma$, have finite measure, we say that $X$ is \emph{finite}.
+A measure space is \emph{semi-finite} if for all $S \in \Sigma$ such that $\mu(S) = \infty$, there exists a set $T \in \Sigma$ such that $T \subseteq S$ and $0 < \mu(T) < \infty$.
+Applying Zorn's lemma, we see that a measurable subset of a semi-finite measure space has infinite measure \emph{if and only if} it contains subsets of arbitrarily large finite measure.
+
+A measure space is \emph{localizable} $\mu$ is semi-finite and $\Sigma$ modulo the null sets is complete as a Boolean algebra, which means that every collection of measurable sets has an \emph{essential supremum}.
 Localizable measure spaces were introduced by Segal \cite{segal} as a characterization of the measure spaces such that $L^\infty(X)$, the C$^*$-algebra of essentially bounded measurable functions modulo null sets, is a W$^*$-algebra.
@@
 As $\Sigma / \Null$ is a complete Boolean algebra, such a map is necessarily completely additive, not just countably additive \cite[321E]{fremlin2}.
+The pair $(\Sigma/\Null, \mu)$ is called the \emph{measure algebra} of $X$.
 
-\tob{So we can e.g.
-take the measurable space $(S,2^S)$ for some uncountable $S$, and equip it with the measure $\mu$ which vanishes on any countable set, and is $\infty$ on all uncountable sets.
-This is a localizable measure space, right?
-I guess that it's not compact though, and clearly not strictly localizable.
-Is this a good (counter-)example for us?}
+We can shed some light on the notion of localizability with some simple counter-examples based on the same underlying $\sigma$-algebra.
+Let $X$ be any uncountable set, and take $\Sigma$ to be the sets that are either countable or have countable complement (these sets are called \emph{cocountable}).
+This is easily seen to be a $\sigma$-algebra.
+We make $(X,\Sigma)$ into a measure space by three different choices of measure, each of which illustrates different features.
+We define the counting measure $\kappa : \Sigma \rightarrow [0,\infty]$ by
+\begin{equation}
+\label{CountingMeasDefnEqn}
+\kappa(S) = \begin{cases} |S| & \text{if } S \text{ is finite} \\
+                          \infty & \text{if } S \text{ is infinite} \end{cases}.
+\end{equation}
+This is a countably additive measure, and its only null set is $\emptyset$.
+The space $(X,\Sigma,\kappa)$ is semi-finite, but not localizable, because $\Sigma$ is not a complete Boolean algebra (otherwise it would contain all subsets of $X$).
 
-Unfortunately, there exist localizable measure spaces $X$, $Y$ such that $L^\infty(X) \cong L^\infty(Y)$ but there is no measurable map from $X$ to $Y$ \todo{example?}.
-To get out of this difficulty, we restrict ourselves to localizable measure spaces arising from \emph{complete strictly localizable compact} measure spaces.
+We define the measure $\mu_c : \Sigma \rightarrow [0,1]$ by
+\begin{equation}
+\label{CountableCocountableMeasDefnEqn}
+\mu_c(S) = \begin{cases} 0 & \text{if } S \text{ is countable} \\
+                         1 & \text{if } S \text{ is cocountable} \end{cases}.
+\end{equation}
+Then $(X,\Sigma,\mu_c)$ is a probability space, and therefore localizable. In fact, we can see that $\Sigma/ \Null$ is a 2-element Boolean algebra, equipped with its unique probability measure.
+
+We define the measure $\mu_\infty : \Sigma \rightarrow [0,\infty]$ by
+\[
+\mu_\infty(S) = \begin{cases} 0 & \text{if } S \text{ is countable} \\
+                             \infty & \text{if } S \text{ is cocountable} \end{cases}.
+\]
+
+Then $(X,\Sigma,\mu_\infty)$ is a measure space such that $\Sigma / \Null$ is a complete Boolean algebra (the 2-element one, as before), but $(X,\Sigma,\mu_\infty)$ is not localizable, because it is not semi-finite.
+
+Unfortunately, there exist localizable measure spaces $X$, $Y$ such that $L^\infty(X) \cong L^\infty(Y)$ but there is no measurable map from $X$ to $Y$ taking nullsets to nullsets under preimage.
+For example, if we consider $(1,\powerset(1),\nu)$, where $\nu$ is the unique probability measure, and the space $(X,\Sigma,\mu_c)$ from \eqref{CountableCocountableMeasDefnEqn}.
+A map $f: 1 \rightarrow X$ that preserves nullsets under preimage cannot exist, because the preimage of every singleton is required to be the empty set.
+But $1$ and $X$ both have $2$ as their measure algebra.
+This is enough to show that $L^\infty(1) \cong L^\infty(X)$, but it can also be seen directly by showing that for each measurable function $f : X \rightarrow \C$ is constant on some cocountable set, so $L^\infty(X) \cong \C \cong L^\infty(1)$.
+
+To get out of this difficulty, we restrict ourselves to \emph{complete strictly localizable compact} measure spaces.
+Doing so still allows us to use all commutative W$^*$-algebras, which would not be the case if we restricted to $\sigma$-finite measures on standard Borel spaces, for example.
 
@@
 
-A measure space $(X,\Sigma,\mu)$ is \emph{strictly localizable} if it can be written as the (typically infinite) disjoint union of finite localizable measure spaces~\cite[211E]{fremlin2}.
+We can prove directly that the space $(X,\Sigma,\mu_c)$ from \eqref{CountableCocountableMeasDefnEqn} is not compact -- if $\Kay$ were a compact class with respect to which $\mu_c$ were inner regular, we could define, for finite set $S \subseteq X$, $K_S \in \Kay$ such that $K_S \subseteq X \setminus S$ and $\mu_c(K_S) = 1$, so $K_S$ is cocountable.
+By definition $\bigcap_{S \Subset X} K_s \subseteq \bigcap X \setminus S = \emptyset$, but the family $(K_S)$ has the finite intersection property, so has nonempty intersection, a contradiction.
+
+A measure space $(X,\Sigma,\mu)$ is \emph{strictly localizable} if it can be written as the (typically infinite) coproduct of finite measure spaces~\cite[211E]{fremlin2}.
+It does not suffice for $(X,\Sigma,\mu)$ to be merely the union of finite measure spaces, or $(X,\Sigma,\kappa)$ from \eqref{CountingMeasDefnEqn} would be strictly localizable, when it isn't even localizable.
 
@@
 \item Objects are measure spaces which are complete, strictly localizable and compact;
-\item Morphisms are equivalence classes of measurable maps $f : X \to Y$ with $\mu_Y \ll f_* \mu_X$ \todo{or something like truly continuous?}, where for measurable $f,g : X \to Y$, we put $f \sim g$ iff for all $T \in \Sigma_Y$, we have $f^{-1}(T) \triangle g^{-1}(T) \in \Null_X$.
-\todo{Do we also want some compatibility with the compact classes?}
+\item Morphisms are equivalence classes of measurable maps $f : X \to Y$ with $\mu_Y \ll f_* \mu_X$ \todo{add reference to ``missing chapter'', it is true, but not trivial, that such a map defines a normal map of W$^*$-algebras.}, where for measurable $f,g : X \to Y$, we put $f \sim g$ iff for all $T \in \Sigma_Y$, we have $f^{-1}(T) \triangle g^{-1}(T) \in \Null_X$.
 \end{enumerate}

### 2018-10-21 direct_integrals.tex
minor edits and corrections

@@
 \newcommand{\Kay}{\mathcal{K}} % I use a convention that the name of a letter spelt phonetically signifies its mathcal version. But semantically, this is a compact class in a measure space.
-\newcommand{\powerset}{\mathcal{P}}
+\newcommand{\powerset}[1]{\mathcal{P}(#1)}
 
@@
 
-A measure space is \emph{localizable} $\mu$ is semi-finite and $\Sigma$ modulo the null sets is complete as a Boolean algebra, which means that every collection of measurable sets has an \emph{essential supremum}.
+A measure space is \emph{localizable} if $\mu$ is semi-finite and $\Sigma$ modulo the null sets is complete as a Boolean algebra, which means that every collection of measurable sets has an \emph{essential supremum}.
 Localizable measure spaces were introduced by Segal \cite{segal} as a characterization of the measure spaces such that $L^\infty(X)$, the C$^*$-algebra of essentially bounded measurable functions modulo null sets, is a W$^*$-algebra.
@@
 \label{CountingMeasDefnEqn}
-\kappa(S) = \begin{cases} |S| & \text{if } S \text{ is finite} \\
-                          \infty & \text{if } S \text{ is infinite} \end{cases}.
+\kappa(S) = \begin{cases} |S| & \text{if } S \text{ is finite}, \\
+                          \infty & \text{if } S \text{ is infinite}. \end{cases}
 \end{equation}
 This is a countably additive measure, and its only null set is $\emptyset$.
-The space $(X,\Sigma,\kappa)$ is semi-finite, but not localizable, because $\Sigma$ is not a complete Boolean algebra (otherwise it would contain all subsets of $X$).
+The space $(X,\Sigma,\kappa)$ is semi-finite, but not localizable, because $\Sigma$ is not a complete Boolean algebra (otherwise it would have to contain all subsets of $X$).
 
@@
 \label{CountableCocountableMeasDefnEqn}
-\mu_c(S) = \begin{cases} 0 & \text{if } S \text{ is countable} \\
-                         1 & \text{if } S \text{ is cocountable} \end{cases}.
+\mu_c(S) = \begin{cases} 0 & \text{if } S \text{ is countable}, \\
+                         1 & \text{if } S \text{ is cocountable}. \end{cases}
 \end{equation}
@@
 \[
-\mu_\infty(S) = \begin{cases} 0 & \text{if } S \text{ is countable} \\
-                             \infty & \text{if } S \text{ is cocountable} \end{cases}.
+\mu_\infty(S) = \begin{cases} 0 & \text{if } S \text{ is countable}, \\
+                             \infty & \text{if } S \text{ is cocountable}. \end{cases}
 \]
@@
 Unfortunately, there exist localizable measure spaces $X$, $Y$ such that $L^\infty(X) \cong L^\infty(Y)$ but there is no measurable map from $X$ to $Y$ taking nullsets to nullsets under preimage.
-For example, if we consider $(1,\powerset(1),\nu)$, where $\nu$ is the unique probability measure, and the space $(X,\Sigma,\mu_c)$ from \eqref{CountableCocountableMeasDefnEqn}.
+For example, if we consider $(1,\powerset{1},\nu)$, where $\nu$ is the unique probability measure, and the space $(X,\Sigma,\mu_c)$ from \eqref{CountableCocountableMeasDefnEqn}.
 A map $f: 1 \rightarrow X$ that preserves nullsets under preimage cannot exist, because the preimage of every singleton is required to be the empty set.
@@
 
-To get out of this difficulty, we restrict ourselves to \emph{complete strictly localizable compact} measure spaces.
+In order to avoid pitfalls like this, we restrict ourselves to \emph{complete strictly localizable compact} measure spaces.
 Doing so still allows us to use all commutative W$^*$-algebras, which would not be the case if we restricted to $\sigma$-finite measures on standard Borel spaces, for example.
@@
 To define these notions, we first define the finite intersection property, and what a compact class is.
-Let $X$ be a set, and $\Kay \subseteq \powerset(X)$ be a family of sets.
+Let $X$ be a set, and $\Kay \subseteq \powerset{X}$ be a family of sets.
 We say it has the \emph{finite intersection property} if for each finite subfamily $(K_i)_{i \in I}$, $K_i \in \Kay$, we have $\bigcap_{i \in I} K_i \neq 0$.

### 2018-10-21 direct_integrals.tex
preliminary stuff on measurable families

@@
 \renewcommand{\H}{\mathcal{H}}	% Hilbert space
+\newcommand{\B}[1]{\mathcal{B}(#1)}	% bounded operators
 \newcommand{\beq}{\begin{equation}}
@@
 \newcommand{\powerset}[1]{\mathcal{P}(#1)}
+
+\newcommand{\Obj}[1]{\mathrm{Obj}(#1)}	% collection of objects of a cat
+\newcommand{\Mor}[1]{\mathrm{Mor}(#1)}	% collection of morphisms of a cat
 
@@
 
-\author{Tobias Fritz, Robert Furber, Bas Westerbaan}
+\author{Tobias Fritz}
+\author{Robert Furber}
+\author{Bas Westerbaan}
 
-\address{Max Planck Institute for Mathematics in the Sciences, Leipzig, Germany}
-\email{fritz@mis.mpg.de}
+\address{Perimeter Institute for Theoeretical Physics}
+\email{tfritz@pitp.ca}
+
+\address{add your affiliation here}
 
@@
 
-\thanks{\textit{Acknowledgements.} }
+\thanks{\textit{Acknowledgements.} Part of this work was conducted while the first author was at the Max Planck Institute for Mathematics in the Sciences in Leipzig, Germany.}
 
@@
 
-\newcommand{\Obj}[1]{\mathrm{Obj}(#1)}
-\newcommand{\Mor}[1]{\mathrm{Mor}(#1)}
-
 \newcommand{\LMeas}{\cat{LocMeas}} 	% category of localizable measure spaces
@@
 
-Now let $\cat{C}$ be a W$^*$-category, and $X \in \LMeas$. Then we define a category $L^\infty(X,\cat{C})$ as follows:
+\todo[inline]{T: The following is all still tentative}
+
+Now let $X \in \LMeas$, and let $\cat{C}$ be a W$^*$-category.
+We say that $\cat{C}$ is a \emph{measurable W$^*$-category} if both objects and morphisms form (potentially large) measurable spaces, and such that all structure maps, including the involution, are measurable; we also require that the restriction to each hom-set $\cat{C}(A,B)$ recovers the Borel (or Baire?) $\sigma$-algebra.
+
+\begin{ex}
+$\NRep{\cat{N}}$ has a canonical measurable structure of this kind if we declare a collection of objects $\mathcal{S}$ to be measurable if for every Hilbert space $\H$, the set of all $\pi : \cat{N} \to \B{\H}$ which are in is $\mathcal{S}$ is measurable, in the sense of the smallest $\sigma$-algebra which makes the evaluation maps $\pi \mapsto \langle \xi,\pi(a) \xi\rangle$ for $\xi\in\H$ and $a\in\cat{N}$ measurable.
+As a degenerate special case, for $\cat{C} = \Hilb$ we obtain the discrete $\sigma$-algebra on the objects.
+
+\todo[inline]{Throughout this example, would it be better to use Borel/Baire $\sigma$-algebras instead?}
+
+On morphisms, we likewise define a collection of morphisms $\mathcal{T}$ to be measurable if for every two Hilbert spaces $\H_1,\H_2$, the set of all triples $(\pi_1,\pi_2,f)$ with $f : \H_1 \to \H_2$ intertwining $\pi_1 : \cat{N} \to \B{\H_1}$ and $\pi_2 : \cat{N} \to \B{\H_2}$ is measurable in the smallest $\sigma$-algebra which makes the three projection maps measurable, where the first two projections map to the space of normal representations on a Hilbert space, where we use the $\sigma$-algebra from the previous paragraph; the third projection maps to $\Hilb(\H_1,\H_2)$, where we similarly use the $\sigma$-algebra generated by the evaluation maps $f \mapsto \langle \zeta, f \xi\rangle$.
+
+\todo[inline]{This does not seem to make composition measurable}
+\end{ex}
+
+In general, we then define a category $L^\infty(X,\cat{C})$ as follows:
 
 \begin{enumerate}
-\item Objects are maps $A : X \to \cat{C}, \: x\mapsto A_x$, which are measurable in the sense that there is a measurable partition of $X$ such that $A$ is constant on each part;
-\item Morphisms $f : A\to B$ are equivalence classes of bounded families $(f_x : A_x\to B_x)_{x\in X}$, which depend measurably \todo{work out what exactly this should mean such that composition makes sense} on $x$, and where $f \sim g$ if upon restriction to any measurable set on which both $A$ and $B$ are constant, we have $f^{-1}(T) \triangle g^{-1}(T) \in \Null_X$ for every measurable $T \subseteq \cat{C}(A,B)$.
+\item Objects are measurable maps $A : X \to \cat{C}, \: x\mapsto A_x$;
+\item Morphisms $f : A\to B$ are equivalence classes of bounded families $(f_x : A_x\to B_x)_{x\in X}$, which depend measurably on $x$, and where equivalence $f \sim g$ means that $f^{-1}(T) \triangle g^{-1}(T) \in \Null_X$ for every measurable $T \subseteq \Mor{\cat{C}}$.
 \end{enumerate}

### 2018-10-22 direct_integrals.tex
Altered the comments with my own comments.

@@
 
-\todo[inline]{Throughout this example, would it be better to use Borel/Baire $\sigma$-algebras instead?}
+\todo[inline]{Throughout this example, would it be better to use Borel/Baire $\sigma$-algebras instead?
+
+Robert: Let's not get hung up.
+For $\H$, Borel and Baire are the same, as its a metric space, with or without separability.
+In the inseparable case, the kind of measurability we need to use is inexpressible by $\sigma$-algebras on the domain and codomain anyhow.
+}
 
@@
 
-\todo[inline]{This does not seem to make composition measurable}
+\todo[inline]{This does not seem to make composition measurable.
+
+Robert: Composition is measurable in the separable case, and the Borel and Baire $\sigma$-algebras of the ultrastrong and ultraweak topologies coincide.
+I do not expect measurability in the inseparable case for the same reason that addition of measurable functions is not measurable for inseparable-Banach-space-valued random variables.
+It's likely that the same solution works, however.}
 \end{ex}
@@
 
-\todo[inline]{Can we make the comparison more precise?}
+\todo[inline]{Can we make the comparison more precise? Yetter only considers the separable case, but in this case I expect they coincide.}
 
@@
 
-\todo[inline]{Try to construct preduals for the hom-sets in terms of spaces like $L^1(X,\cat{C}(A,B)_*)$?}
+\todo[inline]{Try to construct preduals for the hom-sets in terms of spaces like $L^1(X,\cat{C}(A,B)_*)$?
+Bochner or Gelfand-Pettis integral?
+Since we use compact measure spaces, this only amounts to the difference between integrability of $\|f(x)\|$, \emph{i.e.} absolute integrability, vs unconditional integrability.}
 \end{proof}
@@
 
-\todo[inline]{Is the restriction to $\sigma$-finite the ``right'' definition?}
+\todo[inline]{Is the restriction to $\sigma$-finite the ``right'' definition?
+
+Robert: As we use strictly localizable spaces, every measure space is a disjoint union of $\sigma$-finite spaces, or even finite spaces. But do you mean is $\sigma$-finite vs finite right?}
 

### 2018-10-22 direct_integrals.tex
some corrections, plus Example 4.2

@@
 \begin{ex}
-$\NRep{\cat{N}}$ has a canonical measurable structure of this kind if we declare a collection of objects $\mathcal{S}$ to be measurable if for every Hilbert space $\H$, the set of all $\pi : \cat{N} \to \B{\H}$ which are in is $\mathcal{S}$ is measurable, in the sense of the smallest $\sigma$-algebra which makes the evaluation maps $\pi \mapsto \langle \xi,\pi(a) \xi\rangle$ for $\xi\in\H$ and $a\in\cat{N}$ measurable.
+\label{nrep_measurable}
+$\NRep{\cat{N}}$ has a canonical measurable structure of this kind, which we describe in the following.
+First, a collection of representations $\mathcal{S}$ on a fixed Hilbert space $\H$ is measurable if it is in the smallest $\sigma$-algebra which makes the evaluation maps $\pi \mapsto \langle \xi,\pi(a) \xi\rangle$ for $\xi\in\H$ and $a\in\cat{N}$ measurable.
+Furthermore, we equip $\Obj{\Hilb}$ with the countable-cocountable $\sigma$-algebra.
+On $\Obj{\NRep{\cat{N}}}$, we hence put the $\sigma$-algebra consisting of countable or cocountable unions of measurable subsets of $\{ \pi : \cat{N} \to \B{\H} \}$ for varying $\H$.
+
 As a degenerate special case, for $\cat{C} = \Hilb$ we obtain the discrete $\sigma$-algebra on the objects.
@@
 In particular, this means that two objects $A$ and $B$ which differ only on a null set are isomorphic.
+
+\begin{ex}
+In the situation of Example~\ref{nrep_measurable}, a family of representations $x \mapsto (\H_x,\pi_x)$ is measurable if and only if the following holds: there is a countable measurable partition $X = \coprod_{n\in\N} X_n$, such that $x\mapsto \H_x$ is constant on each part $X_n$, and such that on each part $X_n$, the function $x \mapsto \pi_x$ is measurable.
+
+More concretely, let us consider $\Rep{C(S^1)} \cong \NRep{C(S^1)^{**}}$ and $X := S^1$, equipped with its usual $\sigma$-algebra and measure (the normalization is irrelevant).
+To every $x\in X$ we assign the Hilbert space $\C$, together with the unitary given by multiplication by $x$.
+This induces a representation of $C(S^1)$, or equivalently a normal representation of $C(S^1)^{**}$, as usual via functional calculus, which in this case simply corresponds to function evaluation at $x\in S^1$.
+Measurability of the function $x\mapsto \pi_x$ amounts to measurability of each map $x\mapsto \langle\xi,\pi(a)\xi\rangle$ for $\xi\in\C$ and $a\in C(S^1)^{**}$, or equivalently measurability of $x\mapsto |\xi|^2 a(x)$, which is obvious, since the evaluation at $x$ factors through the canonical homomorphism $C(S^1)^{**} \to \mathcal{L}^\infty(S^1)$, where $\mathcal{L}^\infty$ denotes the algebra of bounded measurable functions (without modding out by null sets).
+\end{ex}
 

### 2018-10-25 direct_integrals.tex
Andre Henriques stacky suggestion

@@
 
-\thanks{\textit{Acknowledgements.} Part of this work was conducted while the first author was at the Max Planck Institute for Mathematics in the Sciences in Leipzig, Germany.}
+\thanks{\textit{Acknowledgements.} Part of this work was conducted while the first author was at the Max Planck Institute for Mathematics in the Sciences in Leipzig, Germany.
+We thank Andr\'e{} Henriques for useful discussions.}
 
@@
 
+\todo[inline]{$L^\infty(X,\cat{C})$ is pseudo-functorial in $X$. Does the resulting fibration satisfy interesting descent conditions? (Suggestion due to Andr\'e{} Henriques on MO.)}
+
 \begin{proof}

### 2018-11-30 direct_integrals.tex
Direct sums draft now separate

--- 2018-10-22 direct_integrals.tex
@@
 
+\newcommand{\Calg}{\mathcal{A}}
+
 \begin{ex}
-$\NRep{\cat{N}}$ has a canonical measurable structure of this kind if we declare a collection of objects $\mathcal{S}$ to be measurable if for every Hilbert space $\H$, the set of all $\pi : \cat{N} \to \B{\H}$ which are in is $\mathcal{S}$ is measurable, in the sense of the smallest $\sigma$-algebra which makes the evaluation maps $\pi \mapsto \langle \xi,\pi(a) \xi\rangle$ for $\xi\in\H$ and $a\in\cat{N}$ measurable.
+\label{nrep_measurable}
+Let $\Calg$ be a C$^*$-algebra.
+Then $\Rep{\Calg}$ has a canonical measurable structure of this kind, which we describe in the following.
+First, a collection of representations $\mathcal{S}$ on a fixed Hilbert space $\H$ is measurable if it is in the smallest $\sigma$-algebra which makes the evaluation maps $\pi \mapsto \langle \xi,\pi(a) \xi\rangle$ for $\xi\in\H$ and $a\in\Calg$ measurable.
+Furthermore, we equip $\Obj{\Hilb}$ with the countable-cocountable $\sigma$-algebra.
+On $\Obj{\Rep{\Calg}}$, we hence put the $\sigma$-algebra consisting of countable or cocountable unions of measurable subsets of $\{ \pi : \cat{N} \to \B{\H} \}$ for varying $\H$.
+
 As a degenerate special case, for $\cat{C} = \Hilb$ we obtain the discrete $\sigma$-algebra on the objects.
@@
 In particular, this means that two objects $A$ and $B$ which differ only on a null set are isomorphic.
+
+\begin{ex}
+In the situation of Example~\ref{nrep_measurable}, a family of representations $x \mapsto (\H_x,\pi_x)$ is measurable if and only if the following holds: there is a countable measurable partition $X = \coprod_{n\in\N} X_n$, such that $x\mapsto \H_x$ is constant on each part $X_n$, and such that on each part $X_n$, the function $x \mapsto \pi_x$ is measurable.
+
+More concretely, let us consider $\Rep{C(S^1)}$ and $X := S^1$, equipped with its usual $\sigma$-algebra and measure (the normalization is irrelevant).
+To every $x\in X$ we assign the Hilbert space $\C$, together with the unitary given by multiplication by $x$.
+This induces a representation of $C(S^1)$ as usual via functional calculus, which in this case simply corresponds to function evaluation at $x\in S^1$.
+Measurability of the function $x\mapsto \pi_x$ amounts to measurability of each map $x\mapsto \langle\xi,\pi(a)\xi\rangle$ for $\xi\in\C$ and $a\in C(S^1)$, or equivalently measurability of $x\mapsto |\xi|^2 a(x)$, which is obvious.
+\end{ex}
 
@@
 \end{proof}
+
+\begin{defn}
+A \emph{normed set} $(S,\|\cdot\|)$ is a set $S$ together with a map $\|\cdot\| : S\to \R_+$.
+\end{defn}
+
+\begin{defn}
+Given a W$^*$-category $\cat{C}$, an \emph{affiliated morphism} $f : A\to B$ is a transformation
+\[
+	J \to \cat{C}(-,B),
+\]
+where $J\subseteq\cat{C}(-,A)$ is a sieve that is locally ultraweakly dense.
+\end{defn}
+
+Affiliated morphisms generalize unbounded operators.
+If $f : A \to B$ is an affiliated morphism and $g : B\to C$ a genuine morphism, then we can form the composition $gf$.
+Precomposing by a genuine morphism is not possible in general, since the pullback sieve need not be locally ultraweakly dense.
+
+\newcommand{\aff}{\mathrm{aff}}
+
+\begin{defn}
+A \emph{W$^*$-category with affiliated morphisms} is a W$^*$-category together with a set $\cat{C}_\aff(A,B)$ of affiliated morphisms for every two objects $A,B\in\cat{C}$.
+\end{defn}
+
+\tob{This terminology may not be very good, since affiliated operators still seem to have a reasonable functional calculus}
+
+\begin{ex}
+We consider $\Hilb$ as a W$^*$-category with affiliated morphisms. These correspond to the unbounded operators modulo equivalence.
+
+Let $\Calg$ be a C$^*$-algebra.
+Then $\Rep{\Calg}$ is a W$^*$-category with affiliated morphisms if we take $\Rep{\Calg}_\aff( (\H_1,\pi_1),(\H_2,\pi_2) )$ if we restrict to those equivalence classes which contain (unbounded) operators $T : \H_1 \to \H_2$ with dense domain $D(T) \subseteq \H_1$ which commute with the action of $\Calg$ in the sense that $u D(T) = D(T)$ and $u T = T u$ for all unitaries $u\in\Calg$.
+\end{ex}
 

### 2018-11-30 fromrobert/fromrobert.tex

\documentclass{article}
\usepackage{import}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{url}
\usepackage{xypic}
\usepackage[backref=page]{hyperref} % put [hidelinks] in if you want to hide the link
%\usepackage{showkeys} % comment out in final versions

% the following package defines a command
% \externaldocument[prefix]{documentname}
% to import the labels, with optional 
% prefix prefix from documentname, without
% importing the document. 
% One can then use \ref and \pageref as normal.
\usepackage{xr}

\title{}
\author{}

%Bart's macro for a hidden proof
\newif\ifignore

% flip these round to expose and hide auxproofs
\ignoretrue
%\ignorefalse

\newcommand{\TC}{\ensuremath{\mathcal{TC}}}
\newcommand{\blank}{\ensuremath{{\mbox{-}}}}
\newcommand{\NRep}{\cat{NRep}}
\newcommand{\Ell}{\mathcal{L}}
\newcommand{\tr}{\ensuremath{\mathrm{tr}}}
\renewcommand{\d}[1]{\operatorname{d}\!{#1}}
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\cat}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\Hil}{\ensuremath{\mathcal{H}}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\newenvironment{myproof}[1][\textnormal{\emph{Proof.}}]%
    { \begin{trivlist}%
        \item[\hskip \labelsep {\bfseries #1}]%
    }%
    { \end{trivlist}%
    }

\begin{document}
\maketitle

Dear Tobias,

This is my explanation of why what you wrote about measurability in $\NRep(C(S^1)^{**})$ in Example 4.2 is wrong.

First I will recap your example of a measurable family, so there is no doubt as to what I am referring to. For each $x \in X$, we define the representation $\pi_x : C(S^1) \rightarrow B(\C)$, where $\C$ has its usual Hilbert space structure, by
\[
\pi_x(a)(\psi) = a(x)\psi,
\]
where $a \in C(S^1)$ and $\psi \in \C$. I do not dispute the fact that $(\pi_x)_{x \in S^1}$ is a measurable family of representations of $C(S^1)$ (with $S^1$ being given either the Borel $\sigma$-algebra or Lebesgue measure $\sigma$-algebra when it is considered as the parameter space of the family).

We then extend each $\pi_x$ to a normal representation $\tilde{\pi}_x$ of $C(S^1)^{**}$ in the canonical way. When doing this, it helps to have the following definition. We define, for all $x \in S^1$, $\delta_x \in C(S^1)^*$:
\[
\delta_x(a) = a(x),
\]
where $a \in C(S^1)$. So $\pi_x(a)(\psi) = \delta_x(a)\psi$. The extension $\tilde{\pi}_x$ can be defined by
\[
\tilde{\pi}_x(\Phi)(\psi) = \Phi(\delta_x)\psi,
\]
where $\Phi \in C(S^1)^{**}$ and $\psi \in \C$. It is clear that this agrees with $\pi_x$ on the double-dual embedding $C(S^1) \rightarrow C(S^1)^{**}$, and it is a normal representation because for every trace-class operator $\rho \in \TC(\Hil)$, $\Phi \mapsto \tr(\Phi(\delta_x)\rho) = \Phi(\delta_x)\rho = \Phi(\rho\delta_x)$ is a normal linear functional on $C(S)^{**}$ (arising from evaluating at $\rho\delta_x \in C(S^1)^*$). Since the image of $C(S^1) \rightarrow C(S^1)^{**}$ is $\sigma(C(S^1)^{**},C(S^1)^*)$-dense in $C(S^1)^{**}$, this proves that $\tilde{\pi}_x$, as defined above, is the unique normal extension of $\pi_x$ to $C(S^1)^{**}$.

You claim that the maps $x \mapsto \langle \psi, \tilde{\pi}_x(\Phi)(\psi) \rangle$ are measurable for all $\psi \in \C$, and $\Phi \in C(S^1)^{**}$. Specializing to $\psi = 1$, this would imply that $x \mapsto \Phi(\delta_x)$ is measurable for all $\Phi \in C(S^1)^{**}$. This is what I shall refute in the next section.

\section{Unmeasurability Proof}
To show that these maps are not all measurable, we will need some background bits and pieces. Let $M(S^1)$ be the space of finite complex (Borel) measures on $S^1$, with its usual topology. Then the Riesz representation theorem states that the map $i : M(S^1) \rightarrow C(S^1)^*$
\[
i(\mu)(a) = \int_{S^1}a \d{\mu},
\]
where $a \in C(S^1)$, is a positive linear isomorphism. We use $j : C(S^1)^* \rightarrow M(S^1)$ for its inverse. We can therefore define a map $k : \Ell^\infty(S^1) \rightarrow C(S^1)^{**}$ as follows:
\[
k(a)(\phi) = \int_{S^1}a \d{j(\phi)},
\]
where $a \in C(S^1)$, $\phi \in C(S^1)^*$ and therefore $j(\phi) \in M(S^1)$.

To prove that $k$ is a *-homomorphism, we need the definition of the $\blank^*$ and multiplication in $A^{**}$, where $A$ is a C$^*$-algebra. This can be found in \cite[\S 3.1]{dauns}. For $a,b \in A$, $\phi \in A^*$ and $\Phi,\Psi \in A^{**}$, we define
\begin{align*}
(\phi \cdot a)(b) &= \phi(ab) \\
(\Phi \cdot \phi)(a) &= \Phi(\phi \cdot a) \\
(\Phi \cdot \Psi)(\phi) &= \Phi(\Psi \cdot \phi).
\end{align*}
As we work in the commutative case, it actually does not matter which side we put the multiplication by $a$ on. We have $\phi \cdot a \in A^*$, $\Phi \cdot \phi \in A^*$ and $\Phi \cdot \Psi \in A^{**}$. 
Similarly, the $\blank^*$ is defined by
\begin{align*}
\phi^*(a) &= \overline{\phi(a^*)} \\
\Phi^*(\phi) &= \overline{\Phi(\phi^*)}.
\end{align*}

Recall that given a measure $\mu \in M(S^1)$ and a function $a \in \Ell^1(\mu)$, we can define a measure $a \cdot \mu$ by
\[
(a \cdot \mu)(T) = \int_{S^1}\chi_T a \d{\mu}
\]
We will use the fact that for any finite complex measure $\mu$, $\Ell^\infty(S^1) \subseteq \Ell^1(\mu)$.

\begin{lemma}
\label{ModuleIsoLemma}
Let $a,b \in \Ell^\infty(S^1)$ and $\mu \in M(S^1)$. Then
\[
\int_{S^1} b \d{(a \cdot \mu)} = \int_{S^1} b a \d{\mu}.
\]
This implies that if $a \in C(S^1)$,
\[
i(a \cdot \mu) = i(\mu) \cdot a.
\]
Therefore for $a \in C(S^1)$ and $\phi \in C(S^1)^*$
\[
j(\phi \cdot a) = a \cdot j(\phi).
\]
\end{lemma}
\begin{proof}
We first prove that for all $a, b \in \Ell^\infty(S^1)$,
\[
\int_{S^1} b \d{(a \cdot \mu)} = \int_{S^1} b a \d{\mu}.
\]
It is true for $b = \chi_{T}$ for a Borel set $T$ by definition. By linearity it is true for simple functions. We can then take a sequence of simple functions $(b_i)_{i \in \N}$ approximating any given $b \in \Ell^\infty(S^1)$ pointwise and deduce that it is true for $b \in \Ell^\infty(S^1)$ by the dominated convergence theorem. Since continuous functions are Borel measurable, for all $a, b \in C(S^1)$ we have
\[
i(a \cdot \mu)(b) = \int_{S^1}b \d{(a \cdot \mu)} = \int_{S^1}ba\d{\mu} = i(\mu)(ba) = i(\mu)(ab) = (i(\mu) \cdot a)(b)
\]
This proves the second statement. Now, let $a \in C(S^1)$ and $\phi \in C(S^1)^*$. Then
\[
i(a \cdot j(\phi)) = i(j(\phi)) \cdot a = \phi \cdot a = i(j(\phi \cdot a)),
\]
so, as $i$ is an isomorphism, $a \cdot j(\phi) = j(\phi \cdot a)$. 
\end{proof}

\begin{lemma}
\label{KHomomorphismLemma}
$k$ is a *-homomorphism.
\end{lemma}
\begin{proof}
By linearity of integration, $k$ is linear. Let $a,b \in \Ell^\infty(S^1)$. Then for all $\phi \in C(S^1)^*$:
\[
(k(a) \cdot k(b))(\phi) = k(a)(k(b) \cdot \phi) = \int_{S^1}a \d{j(k(b) \cdot \phi)}
\]
For all $c \in C(S^1)$, we have 
\begin{align*}
(k(b) \cdot \phi)(c) &= k(b)(\phi \cdot c) \\
 &= \int_{S^1}b \d{j(\phi \cdot c)} \\
 &= \int_{S^1}b \d{(c \cdot j(\phi))} & \text{Lemma \ref{ModuleIsoLemma}}\\
 &= \int_{S^1}b c \d{j(\phi)} & \text{Lemma \ref{ModuleIsoLemma}} \\
 &= \int_{S^1}cb \d{j(\phi)} \\
 &= \int_{S^1}c \d{(b \cdot j(\phi))}.
\end{align*}
Therefore $k(b) \cdot \phi = i(b \cdot j(\phi))$, so $j(k(b) \cdot \phi) = b \cdot j(\phi)$. So
\begin{align*}
(k(a) \cdot k(b))(\phi) &= k(a)(k(b) \cdot \phi) \\
 &= \int_{S^1}a \d{j(k(b) \cdot \phi)} \\
 &= \int_{S^1}a \d{(b \cdot j(\phi))} \\
 &= \int_{S^1}ab \d{j(\phi)} & \text{Lemma \ref{ModuleIsoLemma}} \\
 &= k(ab).
\end{align*}
To prove that $k(a^*) = k(a)^*$, we use the fact that $\Phi^*$ is characterized by the fact that $\Phi^*(\phi) = \overline{\Phi(\phi)}$ for all self-adjoint $\phi$. Let $\phi$ be a self-adjoint element of $C(S^1)^*$, so $j(\phi)$ is a real signed measure. Then for all $a \in C(S^1)$,
\begin{align*}
k(a^*)(\phi) &= \int_{S^1}a^* \d{j(\phi)} \\
 &= \int_{S^1}\Re(a^*) + i\Im(a^*) \d{j(\phi)} \\
 &= \int_{S^1}\Re(a) - i\Im(a) \d{j(\phi)} \\
 &= \int_{S^1}\Re(a)\d{j(\phi)} - i\int_{S^1}\Im(a)\d{j(\phi)} \\
 &= \overline{\int_{S^1}\Re(a)\d{j(\phi)} + i\int_{S^1}\Im(a)\d{j(\phi)}} \\
 &= \overline{\int_{S^1}a\d{j(\phi)}} \\
 &= \overline{k(a)(\phi)} \\
 &= k(a)^*(\phi^*) \\
 &= k(a)^*(\phi),
\end{align*}
because $\phi$ is self-adjoint. 
\end{proof}

Using the map $k$, we can construct a family of projections in $C(S^1)^{**}$ corresponding to points in $S^1$. We define, for all $x \in S^1$,
\[
p_x = k(\chi_{\{x\}}).
\]
As $k$ is a *-homomorphism, $p_x$ is a projection in $C(S^1)^{**}$, and $(p_x)_{x \in S^1}$ is an orthogonal family of projections.

\begin{theorem}
There exist elements $\Phi \in C(S^1)^{**}$ such that $x \mapsto \Phi(\delta_x)$ is not measurable. Therefore, for such $\Phi$, $x \mapsto \langle 1, \pi_x(\Phi)(1) \rangle$ is not measurable. 
\end{theorem}
\begin{proof}
Let $T \subseteq S^1$ be an unmeasurable set. As $C(S^1)^{**}$ is a W$^*$-algebra, its projection lattice is complete, so we can define 
\[
\Phi = \bigvee_{x \in T}p_x
\]
By \cite[\S 30 Theorem 1]{halmosHilb}, sums of orthogonal families of projections on a Hilbert space converge to their joins in the strong operator topology (and \emph{a fortiori} in the weak operator topology). By taking a faithful normal representation of $C(S^1)^{**}$ on a Hilbert space, and using the fact that the weak operator topology and the ultraweak topology agree on norm-bounded sets, $\sum_{x \in T}p_x = \Phi$ in the ultraweak topology, $\sigma(C(S^1)^{**},C(S^1)^*)$. So $\Phi(\phi) = \sum_{x \in T}p_x(\phi)$ for all $\phi \in C(S^1)^*$.

For $x,y \in S^1$, we have 
\[
p_x(\delta_y) = k(\chi_{\{x\}})(\delta_y) = \int_{S^1}\chi_{\{x\}}\d{j(\delta_y)} = j(\delta_y)(\{x\}),
\]
which is $1$ if $x = y$ and $0$ if $x \neq y$. So $\Phi(\delta_x)$ is $1$ if $x \in T$ and $0$ otherwise, and so $(x \mapsto \Phi(\delta_x)) = \chi_T$, which is not measurable.
\end{proof}

Your argument refers to a ``canonical morphism'' $C(S^1)^{**} \rightarrow \Ell^\infty(S^1)$, but there is no such thing. Perhaps you are under the illusion that $\Ell^\infty(S^1)$ is a W$^*$-algebra.

To some extent the counterexample above can be viewed as arising from the map $C(S^1)^{**} \rightarrow \ell^\infty(S^1)$, the adjoint of the inclusion $\ell^1(S^1) \rightarrow C(S^1)^*$. This is because $\ell^\infty(S^1)$ is thereby expressed as a direct summand of $C(S^1)^{**}$.

Best wishes,
Robert

\bibliographystyle{unsrt}
\bibliography{fromrobert}

\end{document}

### 2018-11-30 direct_sums.tex
now including direct sums files

\documentclass[reqno,T1,11pt]{amsproc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multicol}
\usepackage[canadian]{babel}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{xifthen} % provides \isempty
\usepackage{xparse} % new commands with more than one optional argument
\usepackage{fixltx2e} % for \textsubscript
\newcommand{\unlim}{\qopname\relax m{unlim}}
\newcommand{\uwlim}{\qopname\relax m{uwlim}}

% commutative diagrams
\usepackage{tikz}
\usetikzlibrary{arrows,cd,decorations.markings,shapes.geometric,shapes}

% color of links
\definecolor{myurlcolor}{rgb}{0,0,0.4}
\definecolor{mycitecolor}{rgb}{0,0.5,0}
\definecolor{myrefcolor}{rgb}{0.5,0,0}
\usepackage[pagebackref,draft=false]{hyperref}
\hypersetup{colorlinks,
linkcolor=myrefcolor,
citecolor=mycitecolor,
urlcolor=myurlcolor}
\renewcommand*{\backref}[1]{$\uparrow$\,#1}

% general mathematical macros
\renewcommand{\H}{\mathcal{H}}	% Hilbert space
\newcommand{\B}[1]{\mathcal{B}(#1)}	% bounded operators
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Rplusext}{\overline{\mathbb{R}}_+} % extended nonnegative reals, as where measures take values
\newcommand{\C}{\mathbb{C}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\eps}{\varepsilon}
\newcommand{\lin}[1]{\mathrm{lin}(#1)}
\newcommand{\dom}[1]{\mathrm{dom}(#1)} 		% domain of a morphism
\newcommand{\cod}[1]{\mathrm{cod}(#1)} 		% codomain of a morphism
\newcommand{\supp}{\mathrm{supp}}
\newcommand{\esssup}{\mathrm{ess\, sup}} 	% essential supremum

% categories
\newcommand{\cat}[1]{\mathsf{#1}}		% generic category
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Vect}[1]{\mathsf{Vect}_{#1}}	% vector spaces over a field
\newcommand{\Ban}{\mathsf{Ban}}			% (complex) Banach spaces
\newcommand{\Hilb}{\mathsf{Hilb}} 		% Hilb, cat of Hilbert spaces
\newcommand{\Rep}[1]{\mathsf{Rep}(#1)}		% cat normal Hilbert space reps of a C$^*$-algebra
\newcommand{\NRep}[1]{\mathsf{nRep}(#1)}	% cat of normal Hilbert space reps of a W$^*$-algebra
\newcommand{\id}[1]{1_{#1}}			% identity morphism
\newcommand{\Kay}{\mathcal{K}} % I use a convention that the name of a letter spelt phonetically signifies its mathcal version. But semantically, this is a compact class in a measure space.
\newcommand{\powerset}[1]{\mathcal{P}(#1)}

\newcommand{\Obj}[1]{\mathrm{Obj}(#1)}	% collection of objects of a cat
\newcommand{\Mor}[1]{\mathrm{Mor}(#1)}	% collection of morphisms of a cat

% Theorem Environments
\swapnumbers
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{qstn}[thm]{Question}
\newtheorem{defn}[thm]{Definition}
\newtheorem{lemdefn}[thm]{Lemma + Definition}
\newtheorem{prob}[thm]{Problem}
\theoremstyle{remark}
\newtheorem{ex}[thm]{Example}
\newtheorem{rem}[thm]{Remark}
\newtheorem{note}[thm]{Note}
\numberwithin{equation}{section}

%%%%%%%%%%%% Enumeration via lowercase letters
\renewcommand{\labelenumi}{(\alph{enumi})}
\renewcommand{\theenumi}{(\alph{enumi})}
\renewcommand{\labelitemi}{$\circ$}

%%% pagebreaks allowed for align environment
\allowdisplaybreaks

% string diagram stuff
\pgfdeclarelayer{edgelayer}
\pgfdeclarelayer{nodelayer}
\pgfsetlayers{edgelayer,nodelayer,main}
\tikzstyle{none}=[inner sep=0pt]
\tikzstyle{simple}=[-,draw=black,line width=1.000]
\tikzstyle{bn}=[circle,inner sep=2pt,fill=black,draw=black,line width=0.8 pt]

%%% draft stuff
%\usepackage[draft]{showkeys}
\usepackage{todonotes}
\newcommand{\tob}[1]{\todo[color=blue!40,inline]{\textbf{T}: #1}}

%-------------------------------------------------------------------

\begin{document}
\sloppy

% vertical spacing in multiline equations
\setlength{\jot}{6pt}

%-------------------------------------------------------------------

%%%%%%%%%%%% title page stuff %%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Infinite direct sums in C$^*$-categories and W$^*$-categories]{The universal property of infinite direct sums in C$^*$-categories and W$^*$-categories}

\author{Tobias Fritz}
\author{Bas Westerbaan}

\address{Perimeter Institute for Theoeretical Physics}
\email{tfritz@pitp.ca}

\address{add your affiliation here}

\keywords{}

\subjclass[2010]{Primary: ; Secondary: }

\thanks{\textit{Acknowledgements.} We thank Robert Furber for copious help and discussion. Part of this work was conducted while the first author was at the Max Planck Institute for Mathematics in the Sciences in Leipzig, Germany.}

\begin{abstract}
When formulating universal properties for objects in a dagger category, one usually expects a universal property to characterize the universal object up to unique \emph{unitary} isomorphism. We observe that this is automatically the case in the important special case of C$^*$-categories if one uses enrichment in Banach spaces. We then formulate such a universal property for infinite direct sums in C$^*$-categories, and prove the equivalence with the existing definition due to Ghez, Lima and Roberts in the case of W$^*$-categories. This notion specializes to the standard infinite direct sums in the case of a W$^*$-category of normal representations of a W$^*$-algebra.

Finding a universal property for the more general case of direct integrals remains an open problem.
\end{abstract}

\maketitle

\section{Introduction}

Among the most central and useful concepts of category theory is the notion of universal property. Its utility relates to the fact that a universal property characterizes an object up to unique isomorphism, and to the vast number of mathematical construction which can be defined or characterized in terms of universal properties. Among these is the concept of \emph{direct sum} of two objects $A_1$ and $A_2$ in an additive category $\cat{C}$: the direct sum $A_1\oplus A_2$, which can be defined in several equivalent ways via the following standard result:

\begin{thm}[{e.g.~\cite[Proposition~1.2.4]{borceux}}]
\label{directsum_equiv_trad}
Let $\cat{C}$ be an additive category and $A_1,A_2\in\cat{C}$ two objects. Then the following pieces of structure are equivalent for an object $A\in\cat{A}$:
\begin{enumerate}
	\item\label{coprod} $A$ represents the functor
		\[
			\cat{C} \longrightarrow \cat{Ab}, \qquad B \longmapsto \cat{C}(A_1, B) \times \cat{C}(A_2, B)
		\]
	\item\label{prod} $A$ represents the functor
		\[
			\cat{C}^\op \longrightarrow \cat{Ab}, \qquad B \longmapsto \cat{C}(A_1, B) \times \cat{C}(A_2, B)
		\]
	\item\label{biprod} There are morphisms $p_i : A \to A_i$ and $\kappa_i : A_i \to A$ satisfying the \emph{biproduct equations}
	\[
		p_i \kappa_j = \delta_{ij}, \qquad \kappa_1 p_1 + \kappa_2 p_2 = 1.
	\]
\end{enumerate}
\end{thm}

Here, \ref{coprod} and~\ref{prod} state that $A$ is a coproduct and product of $A_1$ and $A_2$, respectively. \ref{biprod} is a purely equational characterization which also proves that direct sums are preserved by every additive functor.

Traditionally, universal properties have been formulated mostly for purely algebraic structures only. Treating analytical structures categorically is straightforward in some cases, but very challenging in many other cases. It usually involves equipping the categories under consideration themselves with certain analytical structures. An interesting case of where this happens is the theory of \emph{W$^*$-categories}, which model categories of Hilbert space representations of C$^*$-algebras or W$^*$-algebras, with the most paradigmatic example of a W$^*$-category being $\Hilb$, the category of Hilbert spaces with bounded linear maps.

In this paper, we study \emph{infinite} direct sums of objects in W$^*$-categories and in the more general case of C$^*$-categories. (Our definitions and results also apply to the case of finite direct sums, but reduce to Theorem~\ref{directsum_equiv_trad} in this case.) As the most basic example, the Hilbert space $\ell^2(I)$ for a (finite or) infinite set $I$ is the direct sum of $I$ many copies of $\C$. Part of the \emph{raison d'\^{e}tre} for analytic structure on an algebraic object is that it often makes the objects under consideration more well-behaved. This is indeed what happens in our case as well: while Theorem~\ref{directsum_equiv_trad} does not generalize to infinite direct sums since infinite coproducts and infinite products do no longer coincide in a general additive category (even when they exist), our main result (Theorem~\ref{directsum_equiv}) provides such a generalization for infinite direct sums in W$^*$-categories.

\subsection*{Summary} After giving some background on C$^*$-categories and W$^*$-categories in Section~\ref{background}, we start by investigating universal properties for objects in C$^*$-categories in Section~\ref{univ_props}. We find that defining universal properties in terms of representability of a functor enriched in Banach spaces provides a good notion of universal property in a C$^*$-category, since it guarantees uniqueness up to unique \emph{unitary} isomorphism of the representing object. We then introduce our definition of the particular universal property of infinite direct sums in C$^*$-categories in Section~\ref{sec_directsum_def}. In Section~\ref{wstarcatcase}, we specialize to the case of W$^*$-categories, and prove an analog of Theorem~\ref{directsum_equiv_trad} in this setting, showing that our infinite direct sums are equivalent to the direct sums in W$^*$-categories as introduced originally by Ghez, Lima and Roberts~\cite{wstarcat}.

\subsection*{Direct integrals} We have also attempted to formulate a similar universal property for \emph{direct integrals} in W$^*$-categories, but we have not been able to make this idea work. As has been argued by Simon Henry\footnote{See \href{https://mathoverflow.net/questions/313158/what-does-it-mean-for-a-category-to-admit-direct-integrals}{mathoverflow.net/questions/313158/what-does-it-mean-for-a-category-to-admit-direct-integrals}.}, finding such a universal property requires the consideration of additional structure on the W$^*$-category. Our current best bet is that doing so will involve a notion of \emph{W$^*$-category with rigged objects}, generalizing rigged Hilbert spaces~\cite{vol4}.

\section{C$^*$-categories and W$^*$-categories}
\label{background}

\todo[inline]{State the basic definitions}

\begin{ex}
\label{rep_ex}
If $A$ is a C$^*$-algebra, then the category of non-degenerate representations $\pi : A \to \B{\H}$ on Hilbert spaces $\H$ is a W$^*$-category. The morphisms $(\H_1,\pi_1) \to (\H_2,\pi_2)$ are exactly the \emph{intertwiners}, i.e.~the bounded linear maps $f : \H_1 \to \H_2$ such that the diagram
\[\begin{tikzcd}
	\H_1 \ar{r}{f} \ar[swap]{d}{\pi_1(a)} & \H_2 \ar{d}{\pi_2(a)} \\
	\H_1 \ar{r}{f} & \H_2
\end{tikzcd}\]
commutes for all $a\in A$. It is well-known that this is a W$^*$-category~\cite[Example~2.4]{wstarcat}, where the Banach space structure on each hom-set is the usual one on a space of operators using the operator norm, and the involution is given by the usual Hermitian adjoint. We denote this W$^*$-category by $\Rep{A}$. In the most basic case $A = \C$, we get $\Rep{\C} \cong \Hilb$, the category of Hilbert spaces.
\end{ex}

\begin{ex}
\label{nrep_ex}
If $N$ is a W$^*$-algebra, then the category of normal unital representations on Hilbert spaces is a W$^*$-category, with intertwiners as morphisms as above. We denote this W$^*$-category by $\NRep{N}$. For a C$^*$-algebra $A$ and $A^{**}$ its universal enveloping W$^*$-algebra, we have $\NRep{A^{**}} \cong \Rep{A}$, since non-degenerate representations of $A$ extend uniquely to normal unital representations of $A^{**}$.
\end{ex}

\section{Universal properties in C$^*$-categories}
\label{univ_props}

In this section, we write $\cat{C}$ for a C$^*$-category \cite{wstarcat}.
We start with a generalization of~\cite[Lemma 7]{westerbaan2016universal}
    to C$^*$-categories before discussing universal properties.

\begin{lem}
	Let $A,B\in\cat{C}$. Then for any~$a\colon A \to B$ with~$\|a\| \leq 1$
    and projections~$s \colon A \to A$,
        $t\colon B\to B$,
	the following are equivalent:
\begin{enumerate}
\item $a^* t a \leq 1-s$,
\item $a s a^* \leq 1 - t$,
\item $tas = 0$,
\item $sa^* t = 0$.
\end{enumerate}
\label{contrapositionlemma}
\end{lem}
\begin{proof}
We will first prove~$a^*t a \leq 1-s \:\Leftrightarrow\: tas=0$.
    Assume~$a^* t a \leq 1-s$.
    Then~$0 \leq s a^* t a s \leq s (1-s) s = 0$.
    Thus~$s a^* t a s = 0$, which is to say~$tas = 0$.
    Now, assume~$tas = 0$.
    Then~$ta = ta(1-s)$ and~$a^*t = (1-s)a^* t$.
    Hence~$a^* ta = (1-s)a^*ta(1-s) \leq 1-s$, as desired.
A similar argument proves~$as a^* \leq 1-t \Leftrightarrow sa^*t=0$.
To finish the proof, it is sufficient to
        show~$tas=0 \Leftrightarrow sa^*t=0$, which follows directly
        by applying~$(\ )^*$.
\end{proof}

\begin{lem}
Let~$a \colon A \leftrightarrows B \colon b$ be any two morphisms
    such that~$\|a \| \leq 1$, $\| b\| \leq 1$ and~$ab=1$.
    Then~$b$ is an isometry with~$a=b^*$.
\label{isometrylemma}
\end{lem}
\begin{proof}
    To start, note~$\|a ^*a\| = \|a\|^2 \leq 1$, so~$a^*a \leq 1$.
    Similarly~$b^*b \leq 1$.
    Combined~$1 = b^*a^*ab \leq b^*b \leq 1$, hence~$b^*b=1$.
    Similarly~$aa^*=1$.
    Clearly~$bb^*bb^* = b 1 b^* = bb^*$, so~$b$ is an isometry.
    Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0$.
    So~$a^* (1-0) a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
    Similarly~$bb^* \leq a^*a$. Hence~$bb^*=a^*a$.
    We are done: $a = aa^*a=abb^*=b^*$.
\end{proof}

We will need the full power of this observation only in the next section. However, the following special case is generally useful when working with universal properties in C$^*$-categories:

\begin{cor}
Let $u : A \to B$ be an invertible morphism with $\| u \| \leq 1$ and $\| u^{-1} \| \leq 1$. Then $u$ is unitary.
\label{unitaries}
\end{cor}

Now let $\Ban$ be the category of Banach spaces with bounded linear maps of norm $\leq 1$. Every C$^*$-category is $\Ban$-enriched. Our notion of universal property in a C$^*$-category is the standard one for enriched categories~\cite[Section~1.10]{kelly}, namely with enrichment in $\Ban$:

\begin{defn}
\label{def_univ}
A functor $F : \cat{C} \to \Ban$ is \emph{representable} if there is an object $A \in \cat{C}$ together with a natural isomorphism $\cat{C}(A,-)\cong F$ that is a componentwise isometric isomorphism of Banach spaces.
\end{defn}

Corollary~\ref{unitaries} and the (weak) enriched Yoneda lemma~\cite[Section~1.9]{kelly} imply:

\begin{cor}
\label{unique_up_to_unitary}
For any representable functor $\cat{C} \to \Ban$, the representing object is unique up to unique \emph{unitary} isomorphism
\end{cor}

\begin{proof}
	If both $A$ and $B$ represent $F$ via $\cat{C}(A,-)\cong F\cong \cat{C}(B,-)$, then the enriched Yoneda lemma gives us an isomorphism $u : A \to B$. Since the natural isomorphism between the hom-functors is required to be a componentwise isometry by $\Ban$-enrichment, we necessarily have $\| u \| = \| u^{-1} \| = 1$, making $u$ into a unitary by Corollary~\ref{unitaries}.
\end{proof}

This uniqueness up to unique unitary is a desideratum of a good notion of universal property for C$^*$-categories, and more generally for all dagger categories~\cite{daglims}. Corollary~\ref{unitaries} tells us that the compatibility with the dagger structure is automatic thanks to the enrichment in $\Ban$, so that we do not need to formulate our universal properties in the form of dagger limits as in~\cite{daglims}.

\section{The universal property of infinite direct sums}
\label{sec_directsum_def}

\tob{I've generalized the defn of direct sum from C$^*$-categories, where it works just the same way as for W$^*$-categories, except for Theorem~\ref{directsum_equiv}. Does this seem reasonable?}

We now turn to the study of infinite direct sums by defining them in terms of a universal property as in Definition~\ref{def_univ}, and then proving that the existing notions of direct sum in the category of Hilbert spaces and more generally all W$^*$-categories indeed satisfy this universal property. While our considerations also apply to finite direct sums as a special case, in this case our discussion provides nothing new, since then our universal property reduces to the standard one of biproducts in an additive category.

We again let $\cat{C}$ be a C$^*$-category. Let $I$ be a finite or infinite index set. For $(A_i)_{i\in I}$ a family of objects in $\cat{C}$, we consider the functor
\[
	\bigoplus_i \cat{C}(A_i,-) \: : \: \cat{C} \to \Ban
\]
given by associating to every object $B$ the set
\beq
\label{AiB}
	\bigoplus_i \cat{C}(A_i , B) \ := \ \left\{ (f_i : A_i \to B)_{i\in I} \Biggm\vert \sum_i f_i f_i^* < \infty \right\}.
\eeq
Here, $\sum_i f_i f_i^*$ can be interpreted as converging in the ultraweak topology\footnote{For positive elements~$(x_i)_{i \in I}$ of a W$^*$-algebra, convergence and value of the sum~$\sum_{i \in I} x_i$ coincides in the ultraweak and ultrastrong topology and, in fact, equals the supremum of the partial sums.} in the case where $\cat{C}(B,B)$ is a W$^*$-algebra. If this is not the case, what we mean by the condition $\sum_i f_i f_i^* < \infty$ is that there is a fixed element which upper bounds all partial sums $\sum_{i\in S} f_i f_i^*$ for finite $S\subseteq I$, or that the sum converges ultraweakly in the double dual W$^*$-algebra $\cat{C}(B,B)^{**}$.

The set~\eqref{AiB} is a vector space under componentwise addition and scalar multiplication, where closure under addition follows from $f_x g^*_x + g_x f^*_x \leq f_x f^*_x + g_x g^*_x$. We furthermore consider this vector space as a normed space under the norm
\[
	\| (f_i)_{i \in I} \| := \left\| \sum_i f_i f_i^* \right\|^{1/2},
\]
where the triangle inequality again follows from $f_x g^*_x + g_x f^*_x \leq f_x f^*_x + g_x g^*_x$, and we interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.

\begin{lem}
	Let $\cat{C}$ be a C$^*$-category, $(A_i)_I\in\cat{C}$ a family of objects and $B\in\cat{C}$. Then $\bigoplus_i \cat{C}(A_i,B)$ is complete, and therefore a Banach space.
\label{lem_complete}
\end{lem}

The proof generalizes the standard proof showing that $\ell^2(I)$ is a Hilbert space.

\begin{proof}
Suppose that we are given a Cauchy sequence of families $(f^{(n)})_{n\in\N} = ( ( f_i^{(n)} )_{i\in I} )_{n\in\N}$. For given $j$, the projection map
\[
	\bigoplus_i \cat{C}(A_i,B) \longrightarrow \cat{C}(A_j,B)	
\]
is bounded, and therefore $( f_j^{(n)} )_{n\in\N}$ is Cauchy as well; let us denote the limit by $f_j : A_j \to B$. We will now show that $f \in \bigoplus_i \cat{C}(A_i,B)$ for $f := (f_j)_{j\in I}$, which amounts to verifying the relevant square summability. The Cauchy assumption implies that the sequence $(f^{(n)})_{n\in\N}$ is uniformly bounded by some constant $C$,
\[
	\left\| \sum_i f_i^{(n)} (f_i^{(n)})^* \right\|^{1/2} \leq C.
\]
Restricting to a finite partial sum over $S\subseteq I$ and taking $n\to\infty$ shows that
\[
	\left\| \sum_{i\in S} f_i f_i^* \right\|^{1/2} \leq C
\]
as well. Since $S$ was arbitrary, we conclude that $\| f \| \leq C$.

The convergence $f_i \to f$ works similarly. By assumption, for every $\eps > 0$ we have $\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for almost all $n,m\in\N$. In particular, we know that for any finite $S \subseteq I$,
\[
	\left\| \sum_{i\in S} (f_i^{(n)} - f_i^{(m)})^* (f_i^{(n)} - f_i^{(m)}) \right\|^{1/2} \leq \eps
\]
for almost all $n,m\in\N$. Taking $m\to\infty$ gives
\[
	\left\| \sum_{i\in S} (f_i^{(n)} - f_i)^* (f_i^{(n)} - f_i) \right\|^{1/2} \leq \eps.
\]
Since $S$ was arbitrary, we conclude $\| f^{(n)} - f \| \leq \eps$ for almost all $n$.
\end{proof}

The map $B \mapsto \bigoplus_i \cat{C}(A_i,B)$ is a functorial in $B$ in the obvious way, giving a functor
\[
	\bigoplus_i \cat{C}(A_i,-) \: : \: \cat{C} \to \Ban.
\]
We now get to our main definition:

\begin{defn}
\label{directsumdef}
Let $\cat{C}$ be a C$^*$-category. An \emph{$I$-indexed direct sum} of the family of objects $(A_i)_{i\in I}$ is an object $\bigoplus_i A_i$ which represents the $\Ban$-enriched functor $\bigoplus_i \cat{C}(A_i,-)$.
\end{defn}

By Corollary~\ref{unique_up_to_unitary}, this characterizes infinite direct sums up to unique unitaries. Since $\cat{C}$ is a dagger category, the same holds dually, meaning that $\bigoplus_i A_i$ also represents the functor
\[
	\bigoplus_i \cat{C}(-,A_i) \: : \: \cat{C}^\op \to \Ban
\]
given by associating to every object $B$ the Banach space
\[
	\bigoplus_i \cat{C}(B, A_i) \ := \ \left\{\, (f_i : B \to A_i)_{i\in I} \biggm\vert \sum_i f_i^* f_i < \infty \,\right\},
\]
where the sum is, as before, taken in the ultraweak topology of $\cat{C}(B,B)^{**}$ or to be understood in terms of uniform bounds on finite partial sums, and again considered a Banach space with respect to the norm
\[
	\| (f_i)_{i \in I} \|\  := \ \Bigl\| \,\sum_i f_i^* f_i \,\Bigr\|^{1/2}.
\]
This Banach space is contravariantly functorial in $B$ in the obvious way.

\begin{rem}
For finite $I$, our direct sums are also dagger limits in the sense of~\cite{daglims}. For infinite $I$, this is no longer the case, since infinite direct sums are not even limits; compare the degeneracy result~\cite[Theorem~5.2]{daglims}, which indicates that infinite dagger products cannot be expected to exist. In this sense, our notion of universal property via Definition~\ref{directsumdef} seems better adapted to the category of Hilbert spaces or other W$^*$-categories than that of~\cite{daglims}.
\end{rem}

\begin{rem}
	We could also consider \emph{direct sums with kernel} in the following sense. Let $K = (K_{ij}) \in \C^{I\times I}$ be a positive semidefinite matrix. Then given a family of objects $(A_i)_{i\in I}$ as before, we can consider the functor
	\[
		\cat{C}\longrightarrow \Ban, \qquad B\ \longmapsto\ \bigoplus_i \cat{C}(A_i, B) \ := \ \left\{\, (f_i : A_i \to B)_{i\in I} \biggm\vert \sum_{i,j} K_{ij} f_i f_j^* < \infty \,\right\},
	\]
	where the completeness proof with respect to the norm $f\mapsto \| \sum K_{ij} f_i f_j^* \|^{1/2}$ works in the same way as the proof of Lemma~\ref{lem_complete}.

	We can thus define the \emph{direct sum with kernel $K$} to be an object representing this functor in the same way as above. For example in $\Hilb$ with $A_i = \C$ for all $i\in I$, the resulting direct sum would be the space of families $(\xi_i)_{i\in I}$ with $\xi_i\in\C$ that are $K$-square summable in that $\sum_{i,j} K_{ij} \xi_i \bar{\xi}_j < \infty$, turned into a Hilbert space with respect to the inner product with kernel $K$.
	\tob{Is it worth elaborating on this and prove a generalization of Theorem~\ref{directsum_equiv}?}
\end{rem}

\section{Direct sums in W$^*$-categories}
\label{wstarcatcase}

In a W$^*$-category, we characterize the existence of direct sums as follows:

\begin{thm}
\label{directsum_equiv}
Let $\cat{C}$ be a W$^*$-category and $(A_i)_{i\in I}$ a family of objects in $\cat{C}$. Then the following pieces of structure are equivalent for an object $A$:
\begin{enumerate}
\item\label{universalprop} $A$ is an $I$-indexed direct sum $\bigoplus_i A_i$.
\item\label{universalprop2} 
There is a family of
        morphisms~$\left(\kappa_j : A_j \to A\right)_{j\in I}$
        with~$\sum_{j \in I} \kappa_j \kappa_j^* < \infty$
        such that for any object~$B$
        and family of
        morphisms~$\left(f_j: A_j \to B\right)_{j\in I}$
        with~$\sum_{j \in I}f_j f_j^* < \infty$,
        there is a unique~$f\colon A \to B$
        such that $f \kappa_j = f_j$ for all~$j \in I$, and $\| f \|^2 = \| \sum_{j\in I} f_j f_j^* \|$.
\item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to A\right)_{j\in I}$ such that $\kappa_i^*\kappa_j = \delta_{ij}$, and
\beq
\label{complete}
        \sum_{j\in I} \kappa_j \kappa_j^* \ =\  1_A.
\eeq
\end{enumerate}
\end{thm}

As far as we know, the first definition of infinite direct sums in W$^*$-categories has been given in~\cite[p.~100]{wstarcat} in the form~\ref{biprodsum}. From our perspective, it is preferable to replace this definition by a universal property as in Definition~\ref{directsumdef}, and to regard the original definition of Ghez, Lima and Roberts as an alternative characterization instead.

\begin{proof}
\ref{universalprop2} is a simple reformulation of the universal property \ref{universalprop} via the Yoneda lemma, as follows. Given a natural isometric isomorphism $\cat{C}(A,B) \cong \bigoplus_i \cat{C}(A_i,B)$, we define the family $(\kappa_i)$ to be the image of $\id{A}$. Then for any $f : A \to B$, the naturality diagram
\[\begin{tikzcd}
	\cat{C}(A,A) \ar{r}{\cong} \ar{d}{f\circ -} & \bigoplus_i \cat{C}(A_i,A) \ar{d}{(f\circ -)} \\
	\cat{C}(A,B) \ar{r}{\cong} & \bigoplus_i \cat{C}(A_i,B)
\end{tikzcd}\]
shows that the universal property takes $f$ to the family $(f_j) := (f\kappa_j)$. By the definition of the norm on $\bigoplus_i \cat{C}(A_i,B)$ and the assumption that the universal property is an isometric isomorphism, we indeed have $\| f \|^2 = \left\| \sum_j f_j f_j^* \right\|$, which also gives $\sum_j \kappa_j \kappa_j^* < \infty$ in the case $f = \id{A}$. Conversely, it is clear that any family $(\kappa_j)$ as in~\ref{universalprop2} gives an isometric isomorphism $\cat{C}(A,B) \cong \bigoplus_i \cat{C}(A_i,B)$ that is natural in $B$.

We now derive \ref{biprodsum} from \ref{universalprop2}. Applying the assumption on norms with $(f_j) := (\kappa_j)$ gives $\| \sum_i \kappa_i \kappa_i^* \| = \| \id{A} \|$, and therefore $\| \kappa_j \| \leq 1$ for all $j\in I$. For fixed $j$, the family $(\delta_{ij} : A_i \to A_j)$ corresponds to some morphism $\pi_j : A \to A_j$ such that $\pi_j\circ\kappa_i = \delta_{ij}$. Again by the isometry property, we have $\| \pi_j \| = \| \id{A} \| \leq 1$. By Lemma~\ref{isometrylemma}, this implies $\pi_j = \kappa_j^*$. We therefore have $\kappa_j^*\kappa_i = \delta_{ij}$, as desired. This also implies that the $\kappa_j \kappa_j^*$ are mutually orthogonal projections. In order to also prove the completeness relation~\eqref{complete}, let $p : A\to A$ be any other projection that is orthogonal to each of the $\kappa_j \kappa_j^*$; we need to show $p = 0$. Again by naturality, under the isomorphism $\cat{C}(A,A) \cong \bigoplus_i \cat{C}(A_i,A)$, the new $p$ corresponds to the family $(p\kappa_i) = (0)$, where $p\kappa_i = 0$ follows from the assumed orthogonality. This implies indeed $p=0$ by the uniqueness part of the universal property.

We finally show that \ref{biprodsum} implies \ref{universalprop2}. So let~$\left(f_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
        with~$\sum_{j \in I}f_jf_j^* < \infty$.
The homset~$C(B,A)$
    is a self-dual Hilbert~$\cat{C}(B,B)$-module
    with~$\cat{C}(B,B)$-valued inner
    product~$\langle g, f\rangle \equiv g^*f$~\cite[Proposition~2.15]{wstarcat}.
The self-duality is equivalent to the fact
    that~$\cat{C}(B,A)$ is complete in the \emph{ultranorm topology};
    this is the topology
    generated by the seminorms~$\| f \|_\omega \equiv \omega(\langle f,f\rangle)^\frac{1}{2}$
    indexed by normal states~$\omega$ on~$\cat{C}(B,B)$~\cite[\S149\textsubscript{V}]{bas}, also known as the \emph{s-topology}~\cite[p.~88]{wstarcat}. We now turn to the construction of $f : A \to B$.
    Pick any normal state~$\omega$ on~$\cat{C}(B,B)$
        and finite subset~$S \subseteq I$.  We have
\begin{alignat*}{2}
    \bigl\| \sum_{j \in S} \kappa_j f_j^* \bigr\|^2_\omega
    &\ \equiv \ 
    \omega\Bigl(\bigl(\sum_{i \in S} \kappa_i f_i^*\bigr)^*
            \bigl(\sum_{j \in S} \kappa_j f_j^*\bigr)\Bigr) \\
    &\ =\ 
    \omega\Bigl( \sum_{j \in S} f_j f_j^*
    \Bigr) &\qquad&\text{as~$\kappa_j^*\kappa_i = \delta_{ij}$}.
\end{alignat*}
From this and~$\sum_{j \in I} f_jf_j^* < \infty$,
    it follows that the net~$\left(\sum_{j \in S} \kappa_j f_j^*\right)_S$
    is ultranorm Cauchy as $S$ ranges over all finite subsets of $I$.
    Define~$f := \left( \sum_{j \in I} \kappa_jf_j^* \right)^*$,
        where the sum is understood to converge ultranorm.
The assignment~$g \mapsto \kappa_j^*g$
    yields a bounded~$\cat{C}(B,B)$-linear map between
    the Hilbert~$\cat{C}(B,B)$-modules~$\cat{C}(B,A)$ and~$\cat{C}(B,A_j)$
    and is therefore ultranorm continuous~\cite[\S148]{bas}.
    Hence
        $\kappa_j^* f^*
             =  \kappa_j^* \sum_{i} \kappa_i f_i^*
             =  \sum_i \kappa_{j }^*\kappa_i f_i^*
            =  f_j^*$
            and so~$f \kappa_j = f_j$ as desired.
Next we will show that~$\|f \|^2 = \|\sum_j f_j f_j^*\|$.
    By definition of~$f^*$ and \cite[\S148\textsubscript{V}]{bas}
         we have~
 \begin{equation*}
         \langle f^*, f^*\rangle
         \ =\  \uwlim_{S\subseteq I\text{ finite}} \, \Bigl\langle \sum_{j \in S} \kappa_j f^*_j,
                        \sum_{j \in S}\kappa_j f^*_j \Bigr\rangle
         \ =\  \uwlim_{S\subseteq I\text{ finite}} \sum_{j \in S} f_j f_j^*
        \ = \ \sum_j f_j f_j^*
 \end{equation*}
and so~$\| f \|^2 = \|f^*\|^2 = \|\langle f,f \rangle \|
                = \|\sum_j f_j f_j^* \|$ as desired.
Finally, to show uniqueness,
    assume~$f'\colon A \to B$ is any (other)
    morphism with~$f' \kappa_i = f_i$.
    Note that~$\sum_j \kappa_j \kappa_j^*$
       converges ultrastrongly
       and so ultranorm as well.
With similar reasoning as before,
        the composition map~$g \mapsto f' g$ is ultranorm continuous,
        hence
\begin{equation*}
    f'
    \ =\ f' \sum_{j \in I} \kappa_j \kappa_j^*
    \ =\ \sum_{j \in I} f' \kappa_j \kappa_j^*
    \ =\ \sum_{j \in I} f_j \kappa_j^*
    \ =\ \sum_{j \in I} f \kappa_j \kappa_j^*
    \ =\ f,
\end{equation*}
    where the sums converge ultranorm.
\end{proof}

It is now straightforward to note:

\begin{cor}
Every normal $*$-functor between W$^*$-categories preserves direct sums.
\end{cor}

Rieffel has observed this for the W$^*$-categories $\Rep{A}$ from Example~\ref{rep_ex} as~\cite[Proposition~4.9]{Rieffel1974}. The following was undoubtedly already known to the authors of~\cite{wstarcat}, but we have not found it stated anywhere explicitly. We are now in a position to give a very simple proof.

\begin{prop}
Let $N$ be a W$^*$-algebra. In the W$^*$-category $\NRep{N}$, any $\ell^2$-direct sum
\beq
\label{concrete_directsum}
	\bigoplus_{i\in I} A_i := \left\{ (\xi_i)_{i\in I} \Biggm \vert \xi_i\in A_i,\: \sum_i \langle\xi_i,\xi_i\rangle < \infty \right\},
\eeq
equipped with the componentwise representation of $N$, is an \emph{$I$-indexed direct sum} in the sense of Definition~\ref{directsumdef}. Conversely, every $I$-indexed direct sum in $\NRep{N}$ is of this form.
\end{prop}

\begin{proof}
	Given a concrete direct sum as in~\eqref{concrete_directsum}, it is straightforward to see that the inclusion maps $\kappa_i : A_i \to A$ satisy the conditions of Theorem~\ref{directsum_equiv}\ref{biprodsum}. This shows that all direct sums in $\NRep{N}$ exist; the uniqueness up to unique unitaries therefore implies that every direct sum is of this form.
\end{proof}

\bibliographystyle{plain}
\bibliography{direct_integrals}

\end{document}

% vim: ft=tex.latex

### 2019-06-06 direct_sums.tex
Add some intro to C* and W*-cats and fix a typo

@@
 		\[
-			\cat{C}^\op \longrightarrow \cat{Ab}, \qquad B \longmapsto \cat{C}(A_1, B) \times \cat{C}(A_2, B)
+			\cat{C}^\op \longrightarrow \cat{Ab}, \qquad B \longmapsto \cat{C}(B, A_1) \times \cat{C}(B, A_2)
 		\]
@@
 
-\todo[inline]{State the basic definitions}
+Before we define C$^*$- and~W$^*$-category, we give some context.
+A C$^*$-category is a caterogical generalization of a unital C$^*$-algebra
+    in the sense that in a~C$^*$-category~$\cat{C}$,
+    every homset~$\cat{C}(A,A)$ will be a unital~C$^*$-algebra.
+Similarly~$\cat{C}(A,A)$ will be a~W$^*$-algebra
+    in a~W$^*$-category $\cat{C}$.
+As the definitions of~C$^*$- and~W$^*$-categories
+    will implicitly define~C$^*$- and W$^*$-algebras,
+    we'll make do with a few short comments.
+Every subalgebra of the algebra of complex~$n\times n$-matrices
+    is a C$^*$- and~W$^*$-algebra.
+More generally, every norm closed subalgebra of~$\B{\H}$,
+    the bounded operators on a Hilbert space~$\H$,
+    is a C$^*$-algebra and, in fact,
+    every C$^*$-algebra is isomorphic to such a norm-closed subalgebra
+        of~$\B{\H}$ for some Hilbert space~$\H$.
+Every weakly closed subalgebra of~$\B{\H}$ is a W$^*$-algebra
+    and every W$^*$-algebra is isomorphic to such a weakly closed subalgebra
+Every commutative unital C$^*$-algebra is isomorphic to~$C(X)$ for some
+    compact Hausdorff space~$X$
+        and every~W$^*$-algebra is isomorphic to~$L^\infty(X)$ for some
+            complete finite measure space~$X$.
+\begin{defn}
+    Let~$\cat{C}$ be a category together with a contravariant endofunctor~$*\colon \cat{C} \to \cat{C}^\op$,
+    whose action on arrows we'll denote by~$a\mapsto a^*$.
+    The following definitions are due to Ghez, Lima and Roberts \cite{wstarcat}.
+\begin{enumerate}
+\item
+    $\cat{C}$ (together with~$*$) is called a \emph{complex~$*$-category} if the following axioms hold.
+\begin{enumerate}
+    \item $\cat{C}$ is enriched over complex vector spaces ---
+            i.e.~every homset~$\cat{C}(A,B)$ is a complex vector space
+                and function composition is bilinear.
+    \item The contravariant endofunctor~$*$ is identity-on-objects,
+            involutive and conjugate-linear ---
+         i.e.~$a^*\colon B\to A$, $a^{**}=a$, $(ab)^*=b^*a^*$, $(\lambda a)^* = \overline{\lambda} a^*$ and
+            $(a+c)^* = a^* + c^*$ for arrows~$a, c\colon A \to B$, ~$b\colon C \to A$
+                and complex number~$\lambda$.
+        (This turns~$\cat{C}(A,A)$ into a~$*$-algebra.)
+    \item For any~$a\colon A \to B$
+            we have~$a=0$ whenever~$a^*a=0$.
+                Furthermore, there is a~$b\colon A \to A$
+            with~$a^*a=b^*b$ --- equivalently:
+            $a^*a$ is positive in the~$*$-algebra~$\cat{C}(A,A)$.
+\end{enumerate}
+\item
+    $\cat{C}$ (together with~$*$) is called a~\emph{normed $*$-category}
+            if it is a complex~$*$-category
+            where each~$\cat{C}(A,B)$ is a normed space and~$\| ab \| \leq \|a\|\|b\|$
+                for any~$a\colon B \to C$ and~$b\colon A \to B$.
+\item
+    $\cat{C}$ (together with~$*$) is called a~\emph{C$^*$-category}
+            if it is a normed~$*$-category
+            where every homset is norm-complete
+            and~$\|a \|^2 = \|a^*a\|$.
+        (This turns~$\cat{C}(A,A)$ into a unital~C$^*$-algebra.)
+\item
+    $\cat{C}$ is a~\emph{W$^*$-category}
+        if it is a~C$^*$-category
+        where every homset has a Banach space predual.
+        Equivalently~\cite[Prop.~2.15]{wstarcat}, a C$^*$-category~$\cat{C}$
+            is a~W$^*$-category
+            if every~$\cat{C}(A,A)$ is a W$^*$-algebra
+            and~$\cat{C}(A,B)$ is self-dual
+            as a Hilbert~$(A,A)$-module
+            with inner product~$\langle a,b\rangle \equiv a^*b$.
+\end{enumerate}
+\end{defn}
+
+\begin{ex}
+The category of Hilbert spaces with bounded linear maps between them
+    is a W$^*$-category.
+\end{ex}
 

### 2019-06-18 direct_sums.tex
minor edits

--- 2018-11-30 direct_sums.tex
@@
 \usepackage{fixltx2e} % for \textsubscript
+\usepackage{comment}
 \newcommand{\unlim}{\qopname\relax m{unlim}}
@@
 \renewcommand{\theenumi}{(\alph{enumi})}
+\renewcommand{\labelenumii}{(\roman{enumii})}
+\renewcommand{\theenumii}{(\roman{enumii})}
 \renewcommand{\labelitemi}{$\circ$}
@@
 
-\address{Perimeter Institute for Theoeretical Physics}
+\address{Perimeter Institute for Theoeretical Physics, Waterloo, Ontario, Canada}
 \email{tfritz@pitp.ca}
 
-\address{add your affiliation here}
+\address{Institute for Computing and Information Sciences, Radboud Universiteit Nijmegen, The Netherlands}
+\email{bas@westerbaan.name}
 
@@
 
-\subjclass[2010]{Primary: ; Secondary: }
+\subjclass[2010]{Primary: 46M15; Secondary: 18E05, 46L10}
 
@@
 When formulating universal properties for objects in a dagger category, one usually expects a universal property to characterize the universal object up to unique \emph{unitary} isomorphism.
-We observe that this is automatically the case in the important special case of C$^*$-categories if one uses enrichment in Banach spaces.
+We observe that this is automatically the case in the important special case of C$^*$-categories, provided that one uses enrichment in Banach spaces.
 We then formulate such a universal property for infinite direct sums in C$^*$-categories, and prove the equivalence with the existing definition due to Ghez, Lima and Roberts in the case of W$^*$-categories.
-This notion specializes to the standard infinite direct sums in the case of a W$^*$-category of normal representations of a W$^*$-algebra.
+These infinite direct sums specialize to the usual ones in the category of Hilbert spaces, and more generally in any W$^*$-category of normal representations of a W$^*$-algebra.
 
@@
 Its utility relates to the fact that a universal property characterizes an object up to unique isomorphism, and to the vast number of mathematical construction which can be defined or characterized in terms of universal properties.
-Among these is the concept of \emph{direct sum} of two objects $A_1$ and $A_2$ in an additive category $\cat{C}$: the direct sum $A_1\oplus A_2$, which can be defined in several equivalent ways via the following standard result:
+Among these is the concept of \emph{direct sum} of two objects $A_1$ and $A_2$ in an additive category $\cat{C}$.
+It can be defined in several equivalent ways via the following standard result, where $\cat{Ab}$ is the category of abelian groups.
 
@@
 \label{directsum_equiv_trad}
-Let $\cat{C}$ be an additive category and $A_1,A_2\in\cat{C}$ two objects. Then the following pieces of structure are equivalent for an object $A\in\cat{A}$:
+Let $\cat{C}$ be an additive category and $A_1,A_2\in\cat{C}$ two objects. Then the following pieces of structure are equivalent for an object $A\in\cat{C}$:
 \begin{enumerate}
@@
 It usually involves equipping the categories under consideration themselves with certain analytical structures.
-An interesting case of where this happens is the theory of \emph{W$^*$-categories}, which model categories of Hilbert space representations of C$^*$-algebras or W$^*$-algebras, with the most paradigmatic example of a W$^*$-category being $\Hilb$, the category of Hilbert spaces with bounded linear maps.
+An interesting case of where this happens is the theory of \emph{W$^*$-categories}, which model categories of Hilbert space representations of C$^*$-algebras or W$^*$-algebras.
+The most paradigmatic example of a W$^*$-category being $\Hilb$, the category of Hilbert spaces with bounded linear maps.
 
 In this paper, we study \emph{infinite} direct sums of objects in W$^*$-categories and in the more general case of C$^*$-categories.
-(Our definitions and results also apply to the case of finite direct sums, but reduce to Theorem~\ref{directsum_equiv_trad} in this case.) As the most basic example, the Hilbert space $\ell^2(I)$ for a (finite or) infinite set $I$ is the direct sum of $I$ many copies of $\C$.
+While our definitions and results also apply to the case of finite direct sums, they reduce to the standard Theorem~\ref{directsum_equiv_trad} in this case.
+As the most basic example, the Hilbert space $\ell^2(I)$ for a (finite or) infinite set $I$ is the direct sum of $I$ many copies of $\C$.
 Part of the \emph{raison d'\^{e}tre} for analytic structure on an algebraic object is that it often makes the objects under consideration more well-behaved.
-This is indeed what happens in our case as well: while Theorem~\ref{directsum_equiv_trad} does not generalize to infinite direct sums since infinite coproducts and infinite products do no longer coincide in a general additive category (even when they exist), our main result (Theorem~\ref{directsum_equiv}) provides such a generalization for infinite direct sums in W$^*$-categories.
+This is indeed what happens in our case as well: while Theorem~\ref{directsum_equiv_trad} does not generalize to infinite direct sums, since infinite coproducts and infinite products do no longer coincide in a general additive category (even when they exist), our main result (Theorem~\ref{directsum_equiv}) provides such a generalization for infinite direct sums in W$^*$-categories.
 
@@
 
-\todo[inline]{State the basic definitions}
+A category can be thought of as a many-object version of a monoid.
+This manifests itself in the fact that for any category $\cat{C}$, every endomorphism homset $\cat{C}(A,A)$ is a monoid.
+In the same way, a C$^*$-category is a many-object version of a unital C$^*$-algebra.
+Indeed in a~C$^*$-category~$\cat{C}$, every homset~$\cat{C}(A,A)$ will be a unital~C$^*$-algebra.
+Similarly, a W$^*$-category $\cat{C}$ is a many-object version of a W$^*$-algebra, and every homset~$\cat{C}(A,A)$ indeed carries the structure of a~W$^*$-algebra.
+
+\begin{comment} % T: I'm not sure if the following would be helpful, given that we assume familiarity with C*- and W*-algebras
+As the definitions of~C$^*$- and~W$^*$-categories
+    will implicitly define~C$^*$- and W$^*$-algebras,
+    we only comment briefly on the latter.
+Every subalgebra of the algebra of complex~$n\times n$-matrices
+    is a C$^*$- and~W$^*$-algebra.
+More generally, every norm closed and $*$-closed subalgebra of~$\B{\H}$,
+    the bounded operators on a Hilbert space~$\H$,
+    is a C$^*$-algebra and, in fact,
+    every C$^*$-algebra is isomorphic to such a subalgebra
+        of~$\B{\H}$ for some Hilbert space~$\H$.
+Every weakly closed subalgebra of~$\B{\H}$ is a W$^*$-algebra,
+    and every W$^*$-algebra is isomorphic to such a weakly closed subalgebra.
+Moreover, every commutative unital C$^*$-algebra is isomorphic to~$C(X)$ for some
+    compact Hausdorff space~$X$, 
+        and every commutative W$^*$-algebra is isomorphic to~$L^\infty(X,\mu)$ for some
+            complete finite measure space~$(X,\mu)$.
+\end{comment}
+
+    The following precise definitions are due to Ghez, Lima and Roberts \cite{wstarcat}.
+
+\begin{defn}
+    Let~$\cat{C}$ be a category together with an endofunctor~$*\,\colon \cat{C} \to \cat{C}^\op$,
+    whose action on arrows we denote by~$a\mapsto a^*$. The pair $(\cat{C},*)$ is called a \emph{complex~$*$-category} if the following conditions hold.
+\begin{enumerate}
+    \item $\cat{C}$ is enriched over complex vector spaces, meaning that:
+	    \begin{enumerate}
+		    \item every homset~$\cat{C}(A,B)$ is a complex vector space,
+		    \item composition of morphisms is bilinear.
+	   \end{enumerate}
+    \item The contravariant endofunctor~$*$ is
+	    \begin{enumerate}
+		    \item identity-on-objects: $a^* : B \to A$ for $a : A \to B$,
+		    \item involutive: $a^{**} = a$,
+		    \item conjugate linear:
+			    \[
+				    (\lambda a)^* = \overline{\lambda} a^*,\qquad (a+b)^* = a^* + b^*
+			    \]
+			    for morphisms $a, b : A \to B$ and $\lambda\in\C$.
+	    \end{enumerate}
+    \item For every morphism~$a : A \to B$,
+	    \begin{enumerate}
+		    \item There is $b :  A \to A$ with~$a^*a=b^*b$,
+		    \item $a^*a=0$ implies $a = 0$.
+	    \end{enumerate}
+\end{enumerate}
+\end{defn}
+
+It is easy to see that in a complex $*$-category $\cat{C}$, every endomorphism homset $\cat{C}(A,A)$ carries the structure of a complex $*$-algebra.
+
+\begin{defn}
+A complex $*$-category $(\cat{C},*)$ is called a
+\begin{enumerate}
+	\item \emph{normed $*$-category} if each homset~$\cat{C}(A,B)$ is a normed space such that
+	    \[
+		    \| ab \| \leq \|a\| \, \|b\|
+	    \]
+                for any~$a\colon B \to C$ and~$b\colon A \to B$.
+	\item\label{Cstarcat}
+    \emph{C$^*$-category} if moreover every homset is norm-complete, and the C$^*$-identity
+    \[
+	\|a \|^2 = \|a^*a\|
+     \]
+	holds for every morphism $a$.
+\item\label{Wstarcat}
+    \emph{W$^*$-category} if moreover
+        where every homset has a Banach space predual.
+
+	Equivalently~\cite[Prop.~2.15]{wstarcat}, a C$^*$-category~$\cat{C}$
+            is a~W$^*$-category
+            if every~$\cat{C}(A,A)$ is a W$^*$-algebra
+            and~$\cat{C}(A,B)$ is self-dual
+            as a Hilbert~$(A,A)$-module
+            with inner product~$\langle a,b\rangle \equiv a^*b$.
+\end{enumerate}
+\end{defn}
+
+Clearly if we have \ref{Cstarcat}, then every $\cat{C}(A,A)$ is a C$^*$-algebra. Similarly, \ref{Wstarcat} makes $\cat{C}(A,A)$ into a W$^*$-algebra.
+
+We again refer to~\cite{wstarcat} for more details on the following examples.
+
+\begin{ex}
+The category of Hilbert spaces with bounded linear maps between them
+    is a W$^*$-category in the obvious way, meaning that $*$ takes each bounded linear map to its adjoint and the norm is the operator norm.
+\end{ex}
 
@@
 
-\tob{I've generalized the defn of direct sum from C$^*$-categories, where it works just the same way as for W$^*$-categories, except for Theorem~\ref{directsum_equiv}. Does this seem reasonable?}
-
 We now turn to the study of infinite direct sums by defining them in terms of a universal property as in Definition~\ref{def_univ}, and then proving that the existing notions of direct sum in the category of Hilbert spaces and more generally all W$^*$-categories indeed satisfy this universal property.
@@
 For example in $\Hilb$ with $A_i = \C$ for all $i\in I$, the resulting direct sum would be the space of families $(\xi_i)_{i\in I}$ with $\xi_i\in\C$ that are $K$-square summable in that $\sum_{i,j} K_{ij} \xi_i \bar{\xi}_j < \infty$, turned into a Hilbert space with respect to the inner product with kernel $K$.
-	\tob{Is it worth elaborating on this and prove a generalization of Theorem~\ref{directsum_equiv}?}
 \end{rem}

### 2019-06-28 direct_sums.tex
recap of operator algs

@@
 
+\subsection*{C*-algebras and W*-algebras}  We give a very brief recap of C*-algebras and W*-algebras for the benefit of readers who have a background in category theory but not in operator algebras, while referring to the literature for the details~\cite{sakai}.
+
+\emph{C$^*$-algebras} are abstract versions of algebras of operators on a Hilbert space which are closed in the operator norm and closed under taking adjoint operators, where taking adjoints is encoded in the ``star'' operation $*$.
+Similarly, \emph{W$^*$-algebras} are abstract versions of algebras of operators on a Hilbert space which are in addition closed under limits with respect to pointwise convergence (strong operator topology).
+In fact, every abstract C$^*$-algebra or W$^*$-algebra is isomorphic to one having such a concrete representation.
+For example, every $*$-closed subalgebra of the algebra of complex~$n\times n$-matrices is a C$^*$-algebra and a~W$^*$-algebra.
+Moreover, every commutative unital C$^*$-algebra is isomorphic to~$C(X)$ for some compact Hausdorff space~$X$.
+And every commutative W$^*$-algebra is isomorphic to~$L^\infty(X,\mu)$ for some complete finite measure space~$(X,\mu)$.
+
 \section{C$^*$-categories and W$^*$-categories}
@@
 Similarly, a W$^*$-category $\cat{C}$ is a many-object version of a W$^*$-algebra, and every homset~$\cat{C}(A,A)$ indeed carries the structure of a~W$^*$-algebra.
-
-\begin{comment} % T: I'm not sure if the following would be helpful, given that we assume familiarity with C*- and W*-algebras
-As the definitions of~C$^*$- and~W$^*$-categories
-    will implicitly define~C$^*$- and W$^*$-algebras,
-    we only comment briefly on the latter.
-Every subalgebra of the algebra of complex~$n\times n$-matrices
-    is a C$^*$- and~W$^*$-algebra.
-More generally, every norm closed and $*$-closed subalgebra of~$\B{\H}$,
-    the bounded operators on a Hilbert space~$\H$,
-    is a C$^*$-algebra and, in fact,
-    every C$^*$-algebra is isomorphic to such a subalgebra
-        of~$\B{\H}$ for some Hilbert space~$\H$.
-Every weakly closed subalgebra of~$\B{\H}$ is a W$^*$-algebra,
-    and every W$^*$-algebra is isomorphic to such a weakly closed subalgebra.
-Moreover, every commutative unital C$^*$-algebra is isomorphic to~$C(X)$ for some
-    compact Hausdorff space~$X$, 
-        and every commutative W$^*$-algebra is isomorphic to~$L^\infty(X,\mu)$ for some
-            complete finite measure space~$(X,\mu)$.
-\end{comment}
 
@@
 
-% vim: ft=tex.latex
+ vim: ft=tex.latex

### 2019-06-29 direct_sums.tex
minor edits

@@
 Here, $\sum_i f_i f_i^*$ can be interpreted as converging in the ultraweak topology\footnote{For positive elements~$(x_i)_{i \in I}$ of a W$^*$-algebra, convergence and value of the sum~$\sum_{i \in I} x_i$ coincides in the ultraweak and ultrastrong topology and, in fact, equals the supremum of the partial sums.} in the case where $\cat{C}(B,B)$ is a W$^*$-algebra.
-If this is not the case, what we mean by the condition $\sum_i f_i f_i^* < \infty$ is that there is a fixed element which upper bounds all partial sums $\sum_{i\in S} f_i f_i^*$ for finite $S\subseteq I$, or that the sum converges ultraweakly in the double dual W$^*$-algebra $\cat{C}(B,B)^{**}$.
+If this is not the case, what we mean by the condition $\sum_i f_i f_i^* < \infty$ is that there is a fixed element which upper bounds all partial sums $\sum_{i\in S} f_i f_i^*$ for finite $S\subseteq I$, or equivalently that the sum converges ultraweakly in the double dual W$^*$-algebra $\cat{C}(B,B)^{**}$.
 
-The set~\eqref{AiB} is a vector space under componentwise addition and scalar multiplication, where closure under addition follows from $f_x g^*_x + g_x f^*_x \leq f_x f^*_x + g_x g^*_x$.
+The set~\eqref{AiB} is a vector space under componentwise addition and scalar multiplication, where closure under addition follows from $f_i g^*_i + g_i f^*_i \leq f_i f^*_i + g_i g^*_i$, which holds because of $(f_i - g_i) (f^*_i - g^*_i) \geq 0$.
 We furthermore consider this vector space as a normed space under the norm
@@
 \]
-where the triangle inequality again follows from $f_x g^*_x + g_x f^*_x \leq f_x f^*_x + g_x g^*_x$, and we interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.
+where the triangle inequality again follows from $f_i g^*_i + g_i f^*_i \leq f_i f^*_i + g_i g^*_i$.
+We interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.
 
 \begin{lem}
-	Let $\cat{C}$ be a C$^*$-category, $(A_i)_I\in\cat{C}$ a family of objects and $B\in\cat{C}$. Then $\bigoplus_i \cat{C}(A_i,B)$ is complete, and therefore a Banach space.
+	Let $\cat{C}$ be a C$^*$-category, $(A_i)_{i\in I}\in\cat{C}$ a family of objects and $B\in\cat{C}$. Then $\bigoplus_i \cat{C}(A_i,B)$ is complete, and therefore a Banach space.
 \label{lem_complete}

### 2019-07-02 direct_sums.tex
minor

@@
 	We can thus define the \emph{direct sum with kernel $K$} to be an object representing this functor in the same way as above.
-For example in $\Hilb$ with $A_i = \C$ for all $i\in I$, the resulting direct sum would be the space of families $(\xi_i)_{i\in I}$ with $\xi_i\in\C$ that are $K$-square summable in that $\sum_{i,j} K_{ij} \xi_i \bar{\xi}_j < \infty$, turned into a Hilbert space with respect to the inner product with kernel $K$.
+For example in $\Hilb$ with $A_i = \C$ for all $i\in I$, the resulting direct sum would be the space of families $(\xi_i)_{i\in I}$ with $\xi_i\in\C$ that are $K$-square summable in that $\sum_{i,j} K_{ij} \bar{\xi}_i \xi_j < \infty$, turned into a Hilbert space with respect to the inner product with kernel $K$.
 \end{rem}
@@
                         \sum_{j \in S}\kappa_j f^*_j \Bigr\rangle
-         \ =\  \uwlim_{S\subseteq I\text{ finite}} \sum_{j \in S} f_j f_j^*
+         \ =\  \uwlim_{S\subseteq I\text{ finite}} \, \sum_{j \in S} f_j f_j^*
         \ = \ \sum_j f_j f_j^*

### 2019-07-02 direct_sums.tex
edits in Rem 4.4

@@
 	\[
-		\cat{C}\longrightarrow \Ban, \qquad B\ \longmapsto\ \bigoplus_i \cat{C}(A_i, B) \ := \ \left\{\, (f_i : A_i \to B)_{i\in I} \biggm\vert \sum_{i,j} K_{ij} f_i f_j^* < \infty \,\right\},
+		\cat{C}\longrightarrow \Ban, \qquad B\ \longmapsto\ {\bigoplus_i}^K \cat{C}(A_i, B) \ := \ \left\{\, (f_i : A_i \to B)_{i\in I} \biggm\vert \sum_{i,j} K_{ij} f_i f_j^* < \infty \,\right\},
 	\]
@@
 
-	We can thus define the \emph{direct sum with kernel $K$} to be an object representing this functor in the same way as above.
+	We can thus try to define the \emph{direct sum with kernel $K$} to be an object representing this functor in the same way as above.
 For example in $\Hilb$ with $A_i = \C$ for all $i\in I$, the resulting direct sum would be the space of families $(\xi_i)_{i\in I}$ with $\xi_i\in\C$ that are $K$-square summable in that $\sum_{i,j} K_{ij} \bar{\xi}_i \xi_j < \infty$, turned into a Hilbert space with respect to the inner product with kernel $K$.
+We do not yet know what these direct sums with kernel will amount to in other cases.
 \end{rem}

### 2019-07-10 direct_sums.tex
acknowlegements
updates

@@
 
-\thanks{\textit{Acknowledgements.} We thank Robert Furber for copious help and discussion.
+\thanks{\textit{Acknowledgements.} We thank Robert Furber for copious help and discussion, as well as Chris Heunen and Martti Korvanen for discussion.
 Part of this work was conducted while the first author was at the Max Planck Institute for Mathematics in the Sciences in Leipzig, Germany.}
@@
 
-\subsection*{C*-algebras and W*-algebras}  We give a very brief recap of C*-algebras and W*-algebras for the benefit of readers who have a background in category theory but not in operator algebras, while referring to the literature for the details~\cite{sakai}.
+\subsection*{C*-algebras and W*-algebras}  We give a brief recap of C*-algebras and W*-algebras for the benefit of readers who have a background in category theory but not in operator algebras, while referring to the literature for the details~\cite{sakai,kr,bram}.
 
-\emph{C$^*$-algebras} are abstract versions of algebras of operators on a Hilbert space which are closed in the operator norm and closed under taking adjoint operators, where taking adjoints is encoded in the ``star'' operation $*$.
-Similarly, \emph{W$^*$-algebras} are abstract versions of algebras of operators on a Hilbert space which are in addition closed under limits with respect to pointwise convergence (strong operator topology).
-In fact, every abstract C$^*$-algebra or W$^*$-algebra is isomorphic to one having such a concrete representation.
-For example, every $*$-closed subalgebra of the algebra of complex~$n\times n$-matrices is a C$^*$-algebra and a~W$^*$-algebra.
-Moreover, every commutative unital C$^*$-algebra is isomorphic to~$C(X)$ for some compact Hausdorff space~$X$.
-And every commutative W$^*$-algebra is isomorphic to~$L^\infty(X,\mu)$ for some complete finite measure space~$(X,\mu)$.
+\emph{C$^*$-algebras} are abstract versions of algebras of operators
+on a Hilbert space which are closed in the operator norm and closed
+under taking adjoint operators, where taking adjoints is encoded
+in the ``star'' operation $*$. Similarly, \emph{W$^*$-algebras} are
+abstract versions of algebras of operators on a Hilbert space which
+are in addition closed under limits with respect to the strong
+operator topology, a sort of pointwise converge.\footnote{A W$^*$-algebra
+    of operators is usually called a von Neumann algebra.
+    However, our references \cite{bas,bram} use the terms W$^*$-algebra and
+        von Neumann algebra interchangeably.}
+In fact, every abstract C$^*$-algebra or W$^*$-algebra is isomorphic
+to one having such a concrete representation~\cite[\S30\textsubscript{XIV}, \S48\textsubscript{VII}]{bram}.
+For example, every $*$-closed subalgebra
+of the algebra of complex~$n\times n$-matrices is a C$^*$-algebra
+and a~W$^*$-algebra. Moreover, every commutative unital C$^*$-algebra
+is isomorphic to~$C(X)$ for some compact Hausdorff space~$X$~\cite[\S27]{bram}. And
+every commutative W$^*$-algebra is isomorphic to~$L^\infty(X,\mu)$
+for some complete finite measure space~$(X,\mu)$~\cite[\S53\textsubscript{XI}]{bram}.
 
@@
 A category can be thought of as a many-object version of a monoid.
-This manifests itself in the fact that for any category $\cat{C}$, every endomorphism homset $\cat{C}(A,A)$ is a monoid.
-In the same way, a C$^*$-category is a many-object version of a unital C$^*$-algebra.
-Indeed in a~C$^*$-category~$\cat{C}$, every homset~$\cat{C}(A,A)$ will be a unital~C$^*$-algebra.
-Similarly, a W$^*$-category $\cat{C}$ is a many-object version of a W$^*$-algebra, and every homset~$\cat{C}(A,A)$ indeed carries the structure of a~W$^*$-algebra.
+This manifests itself in the fact that for any category $\cat{C}$,
+every endomorphism homset $\cat{C}(A,A)$ is a monoid. In the same
+way, a C$^*$-category is a many-object version of a unital
+C$^*$-algebra. Indeed in a~C$^*$-category~$\cat{C}$, every
+homset~$\cat{C}(A,A)$ will be a unital~C$^*$-algebra. Similarly, a
+W$^*$-category $\cat{C}$ is a many-object version of a W$^*$-algebra,
+and in a W$^*$-category every homset~$\cat{C}(A,A)$ carries the structure of
+a~W$^*$-algebra.
 
@@
 			    \[
-				    (\lambda a)^* = \overline{\lambda} a^*,\qquad (a+b)^* = a^* + b^*
+				    (\lambda a)^* \ =\  \overline{\lambda} a^*,\qquad (a+b)^* \ =\  a^* + b^*
 			    \]
@@
 	    \begin{enumerate}
-		    \item There is $b :  A \to A$ with~$a^*a=b^*b$,
+		    \item there is a $b :  A \to A$ with~$a^*a=b^*b$,
 		    \item $a^*a=0$ implies $a = 0$.
@@
 	    \[
-		    \| ab \| \leq \|a\| \, \|b\|
+		    \| ab \| \ \leq\  \|a\| \, \|b\|
 	    \]
@@
     \[
-	\|a \|^2 = \|a^*a\|
+	\|a \|^2\  =\  \|a^*a\|
      \]
@@
             if every~$\cat{C}(A,A)$ is a W$^*$-algebra
-            and~$\cat{C}(A,B)$ is self-dual
-            as a Hilbert~$(A,A)$-module
+            and~$\cat{C}(A,B)$ is self dual
+            as a Hilbert~$\cat{C}(A,A)$-module
             with inner product~$\langle a,b\rangle \equiv a^*b$.
@@
 \label{rep_ex}
-If $A$ is a C$^*$-algebra, then the category of non-degenerate representations $\pi : A \to \B{\H}$ on Hilbert spaces $\H$ is a W$^*$-category.
-The morphisms $(\H_1,\pi_1) \to (\H_2,\pi_2)$ are exactly the \emph{intertwiners}, i.e.~the bounded linear maps $f : \H_1 \to \H_2$ such that the diagram
+If $A$ is a C$^*$-algebra, then the category of non-degenerate
+representations $\pi : A \to \B{\H}$ on Hilbert spaces $\H$ is a
+W$^*$-category. The morphisms~$(\H_1,\pi_1) \to (\H_2,\pi_2)$ are
+exactly the \emph{intertwiners}, i.e.~the bounded linear maps~$f :
+\H_1 \to \H_2$ such that the diagram
 \[\begin{tikzcd}
@@
 \[
-	\bigoplus_i \cat{C}(A_i,-) \: : \: \cat{C} \to \Ban
+	\bigoplus_i \cat{C}(A_i,-) \colon \  \cat{C} \to \Ban
 \]
@@
 \label{AiB}
-	\bigoplus_i \cat{C}(A_i , B) \ := \ \left\{ (f_i : A_i \to B)_{i\in I} \Biggm\vert \sum_i f_i f_i^* < \infty \right\}.
+	\bigoplus_i \cat{C}(A_i , B) \ := \ \Bigl\{ (f_i : A_i \to B)_{i\in I} \Bigm\vert\textstyle  \sum_i f_i f_i^* < \infty \Bigr\}.
 \eeq
@@
 \[
-	\| (f_i)_{i \in I} \| := \left\| \sum_i f_i f_i^* \right\|^{1/2},
+	\| (f_i)_{i \in I} \| \ :=\  \Bigl\| \textstyle \sum_i f_i f_i^* \Bigr\|^{1/2},
 \]
@@
 \[
-	\left\| \sum_i f_i^{(n)} (f_i^{(n)})^* \right\|^{1/2} \leq C.
+    \| f^{(n)} \| \ = \ 
+	\Bigl\| \textstyle \sum_i f_i^{(n)} (f_i^{(n)})^* \Bigr\|^{1/2} \ \leq\  C.
 \]
@@
 \[
-	\left\| \sum_{i\in S} f_i f_i^* \right\|^{1/2} \leq C
+	\Bigl\| \textstyle \sum_{i\in S} f_i f_i^* \Bigr\|^{1/2} \ \leq\  C
 \]
@@
 The convergence $f_i \to f$ works similarly.
-By assumption, for every $\eps > 0$ we have $\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for almost all $n,m\in\N$.
+By assumption, for every~$\eps > 0$ we have~$\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for almost all $n,m\in\N$.
 In particular, we know that for any finite $S \subseteq I$,
 \[
-	\left\| \sum_{i\in S} (f_i^{(n)} - f_i^{(m)})^* (f_i^{(n)} - f_i^{(m)}) \right\|^{1/2} \leq \eps
+	\Bigl\| \sum_{i\in S} \bigl(f_i^{(n)} - f_i^{(m)}\bigr) \bigl(f_i^{(n)} - f_i^{(m)}\bigr)^* \Bigr\|^{1/2} \ \leq\  \eps
 \]
@@
 \[
-	\left\| \sum_{i\in S} (f_i^{(n)} - f_i)^* (f_i^{(n)} - f_i) \right\|^{1/2} \leq \eps.
+	\Bigl\| \sum_{i\in S} \bigl(f_i^{(n)} - f_i\bigr) \bigl(f_i^{(n)} - f_i\bigr)^* \Bigr\|^{1/2} \ \leq \ \eps.
 \]

### 2019-07-10 direct_sums.tex
small changes

@@
 
-The proof generalizes the standard proof showing that $\ell^2(I)$ is a Hilbert space.
+%The proof generalizes the standard proof showing that $\ell^2(I)$ is a Hilbert space.
 
@@
 
-The map $B \mapsto \bigoplus_i \cat{C}(A_i,B)$ is a functorial in $B$ in the obvious way, giving a functor
+The map $B \mapsto \bigoplus_i \cat{C}(A_i,B)$ is functorial in $B$ in the obvious way, yielding a functor
 \[
-	\bigoplus_i \cat{C}(A_i,-) \: : \: \cat{C} \to \Ban.
+	\bigoplus_i \cat{C}(A_i,-)  \colon \  \cat{C} \to \Ban.
 \]
@@
 \[
-	\bigoplus_i \cat{C}(B, A_i) \ := \ \left\{\, (f_i : B \to A_i)_{i\in I} \biggm\vert \sum_i f_i^* f_i < \infty \,\right\},
+	\bigoplus_i \cat{C}(B, A_i) \ := \ 
+    \Bigl\{\, \textstyle (f_i : B \to A_i)_{i\in I} \Bigm\vert \sum_i f_i^* f_i < \infty \,\Bigr\},
 \]
@@
 	\[
-		\cat{C}\longrightarrow \Ban, \qquad B\ \longmapsto\ {\bigoplus_i}^K \cat{C}(A_i, B) \ := \ \left\{\, (f_i : A_i \to B)_{i\in I} \biggm\vert \sum_{i,j} K_{ij} f_i f_j^* < \infty \,\right\},
+		\cat{C}\longrightarrow \Ban, \qquad B\ \longmapsto\ {\bigoplus_i}^K \cat{C}(A_i, B)
+        \ := \ \Bigl\{\,  \textstyle (f_i : A_i \to B)_{i\in I} \Bigm\vert \sum_{i,j} K_{ij} f_i f_j^* < \infty \,\Bigr\},
 	\]
@@
 \end{tikzcd}\]
-shows that the universal property takes $f$ to the family $(f_j) := (f\kappa_j)$.
-By the definition of the norm on $\bigoplus_i \cat{C}(A_i,B)$ and the assumption that the universal property is an isometric isomorphism, we indeed have $\| f \|^2 = \left\| \sum_j f_j f_j^* \right\|$, which also gives $\sum_j \kappa_j \kappa_j^* < \infty$ in the case $f = \id{A}$.
-Conversely, it is clear that any family $(\kappa_j)$ as in~\ref{universalprop2} gives an isometric isomorphism $\cat{C}(A,B) \cong \bigoplus_i \cat{C}(A_i,B)$ that is natural in $B$.
+shows that the universal property takes $f$ to the family $(f_j)
+:= (f\kappa_j)$. By the definition of the norm on $\bigoplus_i
+\cat{C}(A_i,B)$ and the assumption that the universal property is
+an isometric isomorphism, we indeed have $\| f \|^2 = \| \sum_j
+f_j f_j^* \|$, which also gives $\sum_j \kappa_j \kappa_j^*
+< \infty$ in the case $f = \id{A}$. Conversely, it is clear that
+any family~$(\kappa_j)$ as in~\ref{universalprop2} gives an isometric
+isomorphism~$\cat{C}(A,B) \cong \bigoplus_i \cat{C}(A_i,B)$ that
+is natural in $B$.
 
@@
 \label{concrete_directsum}
-	\bigoplus_{i\in I} A_i := \left\{ (\xi_i)_{i\in I} \Biggm \vert \xi_i\in A_i,\: \sum_i \langle\xi_i,\xi_i\rangle < \infty \right\},
+	\bigoplus_{i\in I} A_i\  :=\  \Bigl\{ \textstyle (\xi_i)_{i\in I} \Bigm \vert \xi_i\in A_i,\: \sum_i \langle\xi_i,\xi_i\rangle < \infty \Bigr\},
 \eeq

### 2019-07-10 direct_sums.tex
fix

@@
 \[
-	\| (f_i)_{i \in I} \| \ :=\  \Bigl\| \textstyle \sum_i f_i f_i^* \Bigr\|^{1/2},
+	\| (f_i)_{i \in I} \| \ :=\  \Bigl\| \sum_i f_i f_i^* \Bigr\|^{1/2},
 \]
-where the triangle inequality again follows from $f_i g^*_i + g_i f^*_i \leq f_i f^*_i + g_i g^*_i$.
+The triangle inequality for this norm is not directly obvious.
+It follows from the following generalization of the
+    Cauchy--Schwarz inequality, see e.g.~\cite[\S142\textsubscript{III \& IV}]{bas}.
+\begin{equation*}
+    \Bigl(\sum_i f_i g_i^*\Bigr) \Bigl(\sum_j g_j f_j^*\Bigr)
+    \ \leq \ \Bigl\| \sum_i g_ig_i^* \Bigr\| \, \sum_j f_j f_j^* .
+\end{equation*}
 We interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.

### 2019-07-10 direct_sums.tex
minor

@@
 \]
+where we interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.
 The triangle inequality for this norm is not directly obvious.
-It follows from the following generalization of the
-    Cauchy--Schwarz inequality, see e.g.~\cite[\S142\textsubscript{III \& IV}]{bas}.
+t follows from the following generalization of the
+    Cauchy--Schwarz inequality, see e.g.~\cite[\S142\textsubscript{III \& IV}]{bas},
 \begin{equation*}
     \Bigl(\sum_i f_i g_i^*\Bigr) \Bigl(\sum_j g_j f_j^*\Bigr)
-    \ \leq \ \Bigl\| \sum_i g_ig_i^* \Bigr\| \, \sum_j f_j f_j^* .
+    \ \leq \ \Bigl\| \sum_i g_ig_i^* \Bigr\| \: \sum_j f_j f_j^* .
 \end{equation*}
-We interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.
 

### 2019-09-03 direct_sums.tex
comments reviewer

@@
 
-\thanks{\textit{Acknowledgements.} We thank Robert Furber for copious help and discussion, as well as Chris Heunen and Martti Korvanen for discussion.
-Part of this work was conducted while the first author was at the Max Planck Institute for Mathematics in the Sciences in Leipzig, Germany.}
+\thanks{\textit{Acknowledgements.} We thank Robert Furber for copious
+help and discussion, as well as Chris Heunen and Martti Karvonen
+for discussion. Part of this work was conducted while the first
+author was at the Max Planck Institute for Mathematics in the
+Sciences in Leipzig, Germany.}
 
@@
 
-Traditionally, universal properties have been formulated mostly for purely algebraic structures only.
-Treating analytical structures categorically is straightforward in some cases, but very challenging in many other cases.
-It usually involves equipping the categories under consideration themselves with certain analytical structures.
-An interesting case of where this happens is the theory of \emph{W$^*$-categories}, which model categories of Hilbert space representations of C$^*$-algebras or W$^*$-algebras.
-The most paradigmatic example of a W$^*$-category being $\Hilb$, the category of Hilbert spaces with bounded linear maps.
+There appear many useful universal properties for purely
+algebraic structures.  Treating analytical structures
+categorically is just as straightforward in some cases, but very challenging
+in many other cases. It usually involves equipping the categories
+under consideration themselves with certain analytical structures.
+An interesting case of where this happens is the theory of
+\emph{W$^*$-categories}, which model categories of Hilbert space
+representations of C$^*$-algebras or W$^*$-algebras. The most
+paradigmatic example of a W$^*$-category being $\Hilb$, the category
+of Hilbert spaces with bounded linear maps.
 
@@
 	    \begin{enumerate}
-		    \item there is a $b :  A \to A$ with~$a^*a=b^*b$,
+		    \item there is a morphism $b :  A \to A$ with~$a^*a=b^*b$,
 		    \item $a^*a=0$ implies $a = 0$.
@@
 	holds for every morphism $a$.
+        (This turns $\cat{C}(A,A)$ into a~C$^*$-algebra.)
 \item\label{Wstarcat}
@@
 \end{enumerate}
+As the homset~$\cat{C}(A,A)$
+    in a C$^*$-category
+    is a C$^*$-algebra,
+    we will freely use definitions from~C$^*$-algebras on endomorphisms,
+    so we say~$a\colon A \to A$ is a \emph{projection}
+        if it is a projection as an element of the~C$^*$-algebra
+        $\cat{C}(A,A)$.
+Several notions extend naturally to all morphisms:
+    a morphism~$a \colon A \to B$ is called an \emph{isometry} if~$a^*a=1$
+        (and in that case~$aa^*$ is a projection)
+        and it is a \emph{unitary} if additionally~$aa^*=1$.
 \end{defn}
-
-Clearly if we have \ref{Cstarcat}, then every $\cat{C}(A,A)$ is a C$^*$-algebra. Similarly, \ref{Wstarcat} makes $\cat{C}(A,A)$ into a W$^*$-algebra.
 
@@
     Similarly~$b^*b \leq 1$.
-    Combined~$1 = b^*a^*ab \leq b^*b \leq 1$, hence~$b^*b=1$.
+    Combined~$1 = b^*a^*ab \leq b^*b \leq 1$, hence~$b^*b=1$
+        and so~$b$ is an isometry.
     Similarly~$aa^*=1$.
-    Clearly~$bb^*bb^* = b 1 b^* = bb^*$, so~$b$ is an isometry.
-    Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0$.
+    Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0 \leq 0$.
     So~$a^* (1-0) a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
@@
 
-We again let $\cat{C}$ be a C$^*$-category. Let $I$ be a finite or infinite index set. For $(A_i)_{i\in I}$ a family of objects in $\cat{C}$, we consider the functor
+We again let $\cat{C}$ be a C$^*$-category. Let $I$ be a finite or infinite index set. For a family $(A_i)_{i\in I}$ of objects in $\cat{C}$, we consider the functor
 \[
@@
 where we interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.
-The triangle inequality for this norm is not directly obvious.
-t follows from the following generalization of the
-    Cauchy--Schwarz inequality, see e.g.~\cite[\S142\textsubscript{III \& IV}]{bas},
-\begin{equation*}
-    \Bigl(\sum_i f_i g_i^*\Bigr) \Bigl(\sum_j g_j f_j^*\Bigr)
-    \ \leq \ \Bigl\| \sum_i g_ig_i^* \Bigr\| \: \sum_j f_j f_j^* .
-\end{equation*}
+(This norm is in fact the one associated to the~$B$-valued inner
+        product~$\langle (f_i)_i, (g_i)_i \rangle  :=
+                \sum_i f_ig_i^*$,
+                i.e.~$\| (f_i)_i \|^2 = \|\langle (f_i)_i, (f_i)_i \rangle\|$.
+        The proof of the triangle inequality
+                is similar to that for a
+            norm associated to a regular inner product,
+            using a generalized Cauchy--Schwarz inequality.
+            See for instance~\cite[\S142\textsubscript{III--V}]{bas}
+                or~\cite[Prop.~2.3]{paschke}.)
 
@@
 We will now show that $f \in \bigoplus_i \cat{C}(A_i,B)$ for $f := (f_j)_{j\in I}$, which amounts to verifying the relevant square summability.
-The Cauchy assumption implies that the sequence $(f^{(n)})_{n\in\N}$ is uniformly bounded by some constant $C$,
+The Cauchy assumption implies that the sequence $(f^{(n)})_{n\in\N}$ is uniformly bounded by some constant $C$, i.e.
 \[
@@
 
-The convergence $f_i \to f$ works similarly.
+    The convergence $f^{(n)} \to f$ works similarly.
 By assumption, for every~$\eps > 0$ we have~$\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for almost all $n,m\in\N$.
@@
         there is a unique~$f\colon A \to B$
-        such that $f \kappa_j = f_j$ for all~$j \in I$, and $\| f \|^2 = \| \sum_{j\in I} f_j f_j^* \|$.
+        such that $f \kappa_j = f_j$ for all~$j \in I$.
+Moreover, such an~$f$ satisfies~$\| f \|^2 = \| \sum_{j\in I} f_j f_j^* \|$.
 \item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to A\right)_{j\in I}$ such that $\kappa_i^*\kappa_j = \delta_{ij}$, and
@@
     product~$\langle g, f\rangle \equiv g^*f$~\cite[Proposition~2.15]{wstarcat}.
-The self-duality is equivalent to the fact
+The self duality is equivalent to the fact
     that~$\cat{C}(B,A)$ is complete in the \emph{ultranorm topology};
@@
 From this and~$\sum_{j \in I} f_jf_j^* < \infty$,
-    it follows that the net~$\left(\sum_{j \in S} \kappa_j f_j^*\right)_S$
+    it follows that the net~$(\sum_{j \in S} \kappa_j f_j^*)_S$
     is ultranorm Cauchy as $S$ ranges over all finite subsets of $I$.
-    Define~$f := \left( \sum_{j \in I} \kappa_jf_j^* \right)^*$,
-        where the sum is understood to converge ultranorm.
+    Define~$f := ( \sum_{j \in I} \kappa_jf_j^* )^*$,
+        where the sum is understood to converge in the ultranorm topology.
 The assignment~$g \mapsto \kappa_j^*g$
@@
        converges ultrastrongly
-       and so ultranorm as well.
+       and so in the ultranorm topology as well.
 With similar reasoning as before,
@@
 \end{equation*}
-    where the sums converge ultranorm.
+    where the sums converge in the ultranorm topology.
 \end{proof}

### 2019-09-23 direct_sums.tex
further minor corrections

@@
 
-There appear many useful universal properties for purely
+There exist many useful universal properties throughout mathematics, especially so for purely
 algebraic structures.  Treating analytical structures
@@
 \item\label{Wstarcat}
-    \emph{W$^*$-category} if moreover
-        where every homset has a Banach space predual.
+    \emph{W$^*$-category} if moreover every homset has a Banach space predual.
 
@@
     we will freely use definitions from~C$^*$-algebras on endomorphisms,
-    so we say~$a\colon A \to A$ is a \emph{projection}
-        if it is a projection as an element of the~C$^*$-algebra
-        $\cat{C}(A,A)$.
+    so we say e.g.~that~$a\colon A \to A$ is a \emph{projection} if it is a projection as an element of the~C$^*$-algebra $\cat{C}(A,A)$.
+
 Several notions extend naturally to all morphisms:
@@
 
-In this section, we write $\cat{C}$ for a C$^*$-category \cite{wstarcat}.
+In this section, we write $\cat{C}$ for a C$^*$-category.
 We start with a generalization of~\cite[Lemma 7]{westerbaan2016universal}
@@
     Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0 \leq 0$.
-    So~$a^* (1-0) a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
+    So~$a^* 1 a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
     Similarly~$bb^* \leq a^*a$. Hence~$bb^*=a^*a$.
@@
     The convergence $f^{(n)} \to f$ works similarly.
-By assumption, for every~$\eps > 0$ we have~$\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for almost all $n,m\in\N$.
+By assumption, for every~$\eps > 0$ we have~$\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for sufficiently large $n,m\in\N$.
 In particular, we know that for any finite $S \subseteq I$,
@@
 \]
-for almost all $n,m\in\N$. Taking $m\to\infty$ gives
+for sufficiently large $n,m\in\N$. Taking $m\to\infty$ gives
 \[
@@
 \]
-Since $S$ was arbitrary, we conclude $\| f^{(n)} - f \|  \leq \eps$ for almost all $n$.
+Since $S$ was arbitrary, we conclude $\| f^{(n)} - f \|  \leq \eps$ for sufficiently large $n$.
 \end{proof}
 
-The map $B \mapsto \bigoplus_i \cat{C}(A_i,B)$ is functorial in $B$ in the obvious way, yielding a functor
+The map $B \mapsto \bigoplus_i \cat{C}(A_i,B)$ is functorial in $B$ in the obvious way: for $g : B \to C$, a family $(f_i : A_i \to B)_{i\in I}$ gets mapped to $(gf_i : A_i \to C)_{i\in I}$, which satisfies the relevant square integrability condition because of
+\[
+	\sum_i (g f_i)(g f_i)^* = g \left( \sum_i f_i f_i^* \right) g^* < \infty.
+\]
+Thus we get a functor
 \[
@@
 \begin{rem}
+	One may wonder whether our direct sums can be obtained as \emph{weighted limits} or \emph{weighted colimits}, which are generally regarded as the adequate notion of limits and colimits in an enriched context like ours.
+We believe this not to be the case: using enrichment in $\Ban$, there does not seem to be any weight which would give rise to the functor $\bigoplus_i \cat{C}(A_i,-)$ from~\eqref{AiB}.
+The main reason is that mere $\Ban$-enrichment does not have access to the involution, which however is a crucial piece of structure in~\eqref{AiB}.
+\end{rem}
+
+\begin{rem}
 	We could also consider \emph{direct sums with kernel} in the following sense.
@@
 
-	We can thus try to define the \emph{direct sum with kernel $K$} to be an object representing this functor in the same way as above.
+	We can thus define the \emph{direct sum with kernel $K$} to be an object representing this functor in the same way as above.
 For example in $\Hilb$ with $A_i = \C$ for all $i\in I$, the resulting direct sum would be the space of families $(\xi_i)_{i\in I}$ with $\xi_i\in\C$ that are $K$-square summable in that $\sum_{i,j} K_{ij} \bar{\xi}_i \xi_j < \infty$, turned into a Hilbert space with respect to the inner product with kernel $K$.
@@
 Applying the assumption on norms with $(f_j) := (\kappa_j)$ gives $\| \sum_i \kappa_i \kappa_i^* \| = \| \id{A} \|$, and therefore $\| \kappa_j \| \leq 1$ for all $j\in I$.
-For fixed $j$, the family $(\delta_{ij} : A_i \to A_j)$ corresponds to some morphism $\pi_j : A \to A_j$ such that $\pi_j\circ\kappa_i = \delta_{ij}$.
-Again by the isometry property, we have $\| \pi_j \| = \| \id{A} \| \leq 1$.
+For fixed $j$, the family $(\delta_{ij} : A_i \to A_j)_{i\in I}$ corresponds to some morphism $\pi_j : A \to A_j$ such that $\pi_j\circ\kappa_i = \delta_{ij}$.
+Again by the isometry property, we have $\| \pi_j \|^2 = \| \sum_i \delta_{ij} \delta_{ij}^* \| = \| \id{A_j} \| \leq 1$.
 By Lemma~\ref{isometrylemma}, this implies $\pi_j = \kappa_j^*$.

### 2019-09-24 direct_sums.tex
at ucl now

@@
 
-\address{Institute for Computing and Information Sciences, Radboud Universiteit Nijmegen, The Netherlands}
+\address{Department of Computer Science, University College London, United Kingdom}
 \email{bas@westerbaan.name}
@@
 
-There exist many useful universal properties throughout mathematics, especially so for purely
+There appear many useful universal properties for purely
 algebraic structures.  Treating analytical structures
@@
 \item\label{Wstarcat}
-    \emph{W$^*$-category} if moreover every homset has a Banach space predual.
+    \emph{W$^*$-category} if moreover
+        where every homset has a Banach space predual.
 
@@
     Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0 \leq 0$.
-    So~$a^* 1 a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
+    So~$a^* (1-0) a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
     Similarly~$bb^* \leq a^*a$. Hence~$bb^*=a^*a$.

### 2019-09-24 response_to_referees.txt
added old response to referee

Many thanks to the referee for their helpful and considerate suggestions.

"the relationship to weighted (co)limits should be clarified"
-> We have added Remark 4.4 on the referee's question about weighted colimits.

"It seems that the correct spelling is "Karvonen""
-> We have corrected 'Korvanen' to 'Karvonen'.

"I see the point that analytical structures aren't usually understood in terms of universal properties, but the sentence in quotes sounds a bit bizarre - how about geometric structures?"
-> We have rephrased the sentence.

"It would be good for the authors to give the definitions of a projection, isometry, unitary."
-> We have added definitions of isometry and unitary to Definition 2.2. Our proof of Lemma 3.2 was overkill, in that b*b = 1 is enough, and we have removed the extra unnecessary part.

"would read better as "there is a _morphism_ b:A->A...""
-> Fixed.

"Proof of Lemma 3.2: I don't understand how the reasoning..."
-> The reviewer tried to pattern match the wrong thing. We added "... = 1-1 = 0 ≤ 0".

"The word order in "For A_i a  family of objects" would be improved by e.g. "For a family A_i of objects in C.."" or "For a family of objects A_i in C...""
-> Fixed.

"please derive explicitly the triangle equality for the norm"
->  We reformulated the text to make more clear that this is a standard result known from the theory of Hilbert C*-modules. We also added a second citation to Paschke’s original proposition.

"might read better by adding an "i.e.", as in  "is uniformly bounded by some constant C, i.e.""
-> Fixed.

"Typo: should be "The covergence f^{(n)}->f works similarly" rather than "f_i->f""
-> Fixed.

"it might be better to say "for suffienctly large n""
-> We have followed the referee's suggestion of using "for sufficiently large n".

"However, it would make the readers life easier to write down the easy proof"
-> We have added some explanation, now on top of p.7.

"no need to try, just do"
-> Fixed, we just do.

"them 5.1(b) is poorly worded"
-> Fixed.

"proof of (b)->(c)"
-> Yes, corrected.

"Maybe this is standard, but it hurts my eyes"
-> Fixed.

"ref [4] has been published in TAC, please cite it appropriately."
-> Fixed.

### 2019-09-24 direct_sums.tex
typos

--- 2019-09-23 direct_sums.tex
@@
 
-\address{Perimeter Institute for Theoeretical Physics, Waterloo, Ontario, Canada}
+\address{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
 \email{tfritz@pitp.ca}
 
-\address{Institute for Computing and Information Sciences, Radboud Universiteit Nijmegen, The Netherlands}
+\address{Department of Computer Science, University College London, United Kingdom}
 \email{bas@westerbaan.name}
@@
 
-\thanks{\textit{Acknowledgements.} We thank Robert Furber for copious
+\thanks{\textit{Acknowledgments.} We thank Robert Furber for copious
 help and discussion, as well as Chris Heunen and Martti Karvonen
@@
 are in addition closed under limits with respect to the strong
-operator topology, a sort of pointwise converge.\footnote{A W$^*$-algebra
+operator topology, a sort of point-wise converge.\footnote{A W$^*$-algebra
     of operators is usually called a von Neumann algebra.
@@
 	We could also consider \emph{direct sums with kernel} in the following sense.
-Let $K = (K_{ij}) \in \C^{I\times I}$ be a positive semidefinite matrix.
+Let $K = (K_{ij}) \in \C^{I\times I}$ be a positive semi-definite matrix.
 Then given a family of objects $(A_i)_{i\in I}$ as before, we can consider the functor
@@
 \begin{proof}
-	Given a concrete direct sum as in~\eqref{concrete_directsum}, it is straightforward to see that the inclusion maps $\kappa_i : A_i \to A$ satisy the conditions of Theorem~\ref{directsum_equiv}\ref{biprodsum}.
+	Given a concrete direct sum as in~\eqref{concrete_directsum}, it is straightforward to see that the inclusion maps $\kappa_i : A_i \to A$ satisfy the conditions of Theorem~\ref{directsum_equiv}\ref{biprodsum}.
 This shows that all direct sums in $\NRep{N}$ exist; the uniqueness up to unique unitaries therefore implies that every direct sum is of this form.

### 2019-09-24 direct_sums.tex
typos

@@
 Among the most central and useful concepts of category theory is the notion of universal property.
-Its utility relates to the fact that a universal property characterizes an object up to unique isomorphism, and to the vast number of mathematical construction which can be defined or characterized in terms of universal properties.
+Its utility relates to the fact that a universal property characterizes an object up to unique isomorphism, and to the vast number of mathematical constructions which can be defined or characterized in terms of universal properties.
 Among these is the concept of \emph{direct sum} of two objects $A_1$ and $A_2$ in an additive category $\cat{C}$.
@@
 every commutative W$^*$-algebra is isomorphic to~$L^\infty(X,\mu)$
-for some complete finite measure space~$(X,\mu)$~\cite[\S53\textsubscript{XI}]{bram}.
+for some complete finite-measure space~$(X,\mu)$~\cite[\S53\textsubscript{XI}]{bram}.
 
@@
 
-A category can be thought of as a many-object version of a monoid.
+A category can be thought of as a many-objects version of a monoid.
 This manifests itself in the fact that for any category $\cat{C}$,
 every endomorphism homset $\cat{C}(A,A)$ is a monoid. In the same
-way, a C$^*$-category is a many-object version of a unital
+way, a C$^*$-category is a many-objects version of a unital
 C$^*$-algebra. Indeed in a~C$^*$-category~$\cat{C}$, every
 homset~$\cat{C}(A,A)$ will be a unital~C$^*$-algebra. Similarly, a
-W$^*$-category $\cat{C}$ is a many-object version of a W$^*$-algebra,
+W$^*$-category $\cat{C}$ is a many-objects version of a W$^*$-algebra,
 and in a W$^*$-category every homset~$\cat{C}(A,A)$ carries the structure of

### 2019-09-24 direct_sums.tex
typo

@@
 every commutative W$^*$-algebra is isomorphic to~$L^\infty(X,\mu)$
-for some complete finite-measure space~$(X,\mu)$~\cite[\S53\textsubscript{XI}]{bram}.
+for some complete finite measure space~$(X,\mu)$~\cite[\S53\textsubscript{XI}]{bram}.
 
